---
title: "Climate Data Wrangling"
author: "Mia Guarnieri"
date: '2022-11-02'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#attach packages
#note - Loading all of the ones Jason used, I don't think we will need all of them 

library(tidyverse)
library(here)
library(terra)
library(sf)
library(tmap)
library(beepr)
library(rgeos)
library(rmapshaper)
#library(rgdal) fyi rgdal has project function as well and will mess up terra project function

#mia and claire's file path
gdrive_data <- "/Volumes/GoogleDrive/Shared drives/Black_Bear_Aware/gdrive_data"

#sam's file path
#gdrive_data <- "G:/Shared drives/Black_Bear_Aware/gdrive_data"

#chase's file path
```


```{r}
# CA Raster example 
ca_raster <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/CARaster_Tiger_Formatted/formattedcaraster.tif"))

#check the coordinate reference system 
crs(ca_raster)


#plot CA raster to confirm
#tm_shape(ca_raster) +
 # tm_raster() +
#  tmap_mode("view")
```

# Fire 

```{r}
#read in fire layers 
ca_fire_2016 <- rast(here(gdrive_data, "/InputData/InputData_Files/fire_intensity_CA_mtbs/2016_fire_intensity_CA_mtbs/mtbs_CA_2016.tif"))

ca_fire_2017 <- rast(here(gdrive_data, "/InputData/InputData_Files/fire_intensity_CA_mtbs/2017_fire_intensity_CA_mtbs/mtbs_CA_2017.tif"))

ca_fire_2018 <- rast(here(gdrive_data, "/InputData/InputData_Files/fire_intensity_CA_mtbs/2018_fire_intensity_CA_mtbs/mtbs_CA_2018.tif"))

ca_fire_2019 <- rast(here(gdrive_data, "/InputData/InputData_Files/fire_intensity_CA_mtbs/2019_fire_intensity_CA_mtbs/mtbs_CA_2019.tif"))

ca_fire_2020 <- rast(here(gdrive_data, "/InputData/InputData_Files/fire_intensity_CA_mtbs/2020_fire_intensity_CA_mtbs/mtbs_CA_2020.tif"))

```


#Getting layers in correct extent, resolution and projection

```{r}
#reproject to the same crs (NAD 83, CA Albers (2011))
# method = near in order to preserve categorical data

ca_fire_2016_albers <- terra::project(ca_fire_2016, ca_raster, method = "near")

ca_fire_2017_albers <- terra::project(ca_fire_2017, ca_raster, method = "near")

ca_fire_2018_albers <- terra::project(ca_fire_2018, ca_raster, method = "near")

ca_fire_2019_albers <- terra::project(ca_fire_2019, ca_raster, method = "near")

ca_fire_2020_albers <- terra::project(ca_fire_2020, ca_raster, method = "near")

```

```{r}
#cropping to spatial extent of CA raster

fire_2016_mask <- mask(ca_fire_2016_albers, ca_raster)

fire_2017_mask <- mask(ca_fire_2017_albers, ca_raster)

fire_2018_mask <- mask(ca_fire_2018_albers, ca_raster)

fire_2019_mask <- mask(ca_fire_2019_albers, ca_raster)

fire_2020_mask <- mask(ca_fire_2020_albers, ca_raster)
```

#Save reclassified, reprojected, and clipped rasters to the data drive

**ONLY RUN THIS ONE TIME; IT WILL NOT RUN ONCE THE FILES ARE SAVED**

```{r}
saverast <- function(x){
  
  name <- paste0(deparse(substitute(x)), ".tif") #the name of the file will be the same as the input object, plus '.tif' at the end
  
  #write the raster into the correct folder in drive (the path will need to be updated for other files) and save it as a geotiff
  
  writeRaster(x, filename = here(gdrive_data, "/IntermediateData/IntermediateData_Files/model_ready/fire_intensity", name), filetype = "GTiff", overwrite = TRUE)
  
}

saverast(fire_2016_mask)

saverast(fire_2017_mask)

saverast(fire_2018_mask)

saverast(fire_2019_mask)

saverast(fire_2020_mask)

```

#read in refined fire layers: 
```{r}
#read in fire layers 
fire_2016_mask <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/model_ready/fire_intensity/fire_2016_mask.tif"))

fire_2017_mask <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/model_ready/fire_intensity/fire_2017_mask.tif"))

fire_2018_mask <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/model_ready/fire_intensity/fire_2018_mask.tif"))

fire_2019_mask <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/model_ready/fire_intensity/fire_2019_mask.tif"))

fire_2020_mask <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/model_ready/fire_intensity/fire_2020_mask.tif"))

```




#reclass fire intensity layers to layers that only have moderate and high fire intensity 
```{r}
#list out the from, to values in an object
m <- c(1, 0, #Unburned/Underburned/Low Burn Severity to no data
       2, 0, #Low burn severity to no data
       3, 4, #moderate burn severity to moderate/high burn severity
       4, 4, #high burn severity to moderate/high burn severity
       5, 0, #Increased Greenness/Increased Vegetation Response to no data
       6,0) #Non-Processing Area Mask to no data
       
#turn those values into a matrix
mat_fire <- matrix(m, ncol = 2, byrow = TRUE) 

#set categories (labels) for each numeric index

#index vector
index_fire <- c(0, 4)

#categories vector
cats_fire <- c("undefined","mod_high_burn")

#combine those into a dataframe
catdat_fire <- data.frame(id = index_fire, category = cats_fire)

#Reclassify

#create reclassification function

reclassify_fire <- function(x){
  fire_reclassed <- classify(x, mat_fire) #reclassify using the matrix above
  
  levels(fire_reclassed) <- catdat_fire #then reassign levels (labels) using the dataframe above
  
  return(fire_reclassed)
}

#Apply function to each layer



fire_2016_reclassed <- reclassify_fire(fire_2016_mask)

fire_2017_reclassed <- reclassify_fire(fire_2017_mask)

fire_2018_reclassed <- reclassify_fire(fire_2018_mask)

fire_2019_reclassed <- reclassify_fire(fire_2019_mask)

fire_2020_reclassed <- reclassify_fire(fire_2020_mask)






```
#Save reclassified, reprojected, and clipped rasters to the data drive

**ONLY RUN THIS ONE TIME; IT WILL NOT RUN ONCE THE FILES ARE SAVED**

```{r}
saverast <- function(x){
  
  name <- paste0(deparse(substitute(x)), ".tif") #the name of the file will be the same as the input object, plus '.tif' at the end
  
  #write the raster into the correct folder in drive (the path will need to be updated for other files) and save it as a geotiff
  
  writeRaster(x, filename = here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/med_high_fire_rasters", name), filetype = "GTiff", overwrite = TRUE)
  
}

saverast(fire_2016_reclassed)
saverast(fire_2017_reclassed)
saverast(fire_2018_reclassed)
saverast(fire_2019_reclassed)
saverast(fire_2020_reclassed)


```

#Fire distance raster were created in arc gis pro, readin g back in here and readjusting extent

```{r}
#Now we need to read back in the distance raster created in ArcGIS pro and mask these layers to match our CA extent again. Layers are already in the correct projection and resolution

dist_fire16 <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/dist_to_med_high_fire/dist_fire_2016.tif"))

dist_fire17 <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/dist_to_med_high_fire/dist_fire_2017.tif"))

dist_fire18 <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/dist_to_med_high_fire/dist_fire_2018.tif"))

dist_fire19 <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/dist_to_med_high_fire/dist_fire_2019.tif"))

dist_fire20 <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/dist_to_med_high_fire/dist_fire_2020.tif"))

# Now we need to mask these layers to just the CA extent 
fire16_ca_distance <- mask(dist_fire16, ca_raster) 



#now we will save them back to the google drive to be referenced later
```


