---
title: "Distance Raster Creation"
author: "Mia Guarnieri"
date: '2022-10-19'
output: html_document
---

##Setup and packages

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#attach packages
#note - Loading all of the ones Jason used, I don't think we will need all of them 

library(tidyverse)
library(here)
library(terra)
library(sf)
library(tmap)
library(rgrass7)

#mia and claire's file path
#gdrive_data <- "/Volumes/GoogleDrive/Shared drives/Black_Bear_Aware/gdrive_data"

#sam's file path
gdrive_data <- "G:/Shared drives/Black_Bear_Aware/gdrive_data"

#chase's file path

```

#Data Wrangling Prior to Coding

*Data Sources*: Please see ReadMe file for metadata and citations.
*Reference Paper*: Insert hagani citation here

#Load in existing data layers

##California raster

Insert info about CA raster layer here

```{r}

ca_raster <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/CARaster_Tiger_Formatted/formattedcaraster.tif"))

#check the coordinate reference system 
crs(ca_raster)

#plot CA raster to confirm
tm_shape(ca_raster) +
  tm_raster() +
  tmap_mode("view")

```

#Streams Layer 
```{r}
streams_2017 <- vect(here(gdrive_data, "/InputData/InputData_Files/2017_Streams_CDFW/California_Streams.shp"))


streams_2017_calbers <- project(streams_2017, ca_raster)

streams_2017_mask <- crop(streams_2017_calbers, ca_raster)

streams_2017_raster <- rasterize(streams_2017_mask, ca_raster)
```



#Roads Layer 

```{r}
roads_2015 <- vect(here(gdrive_data, "/InputData/InputData_Files/2015_roads_TIGER/tl_2015_06_prisecroads.shp"))

roads_2015_raster <- rasterize(roads_2015, ca_raster)

roads_2015_calbers <- project(roads_2015_raster, ca_raster)

roads_2015_mask <- mask(roads_2015_calbers, ca_raster)
```


#Create remaining data layers

```{r}

```


##Distance to road

```{r}
distance_to_roads <- distance(roads_2015_mask)

```

```{r}
#savevect <- function(x){
  
  #name <- paste0(deparse(substitute(x)), ".shp") #the name of the file will be the same as the input object, plus '.hp' at the end
  
#write the raster into the correct folder in drive (the path will need to be updated for other files) and save it as a shapefile
  
 # writeVector(x, filename = here(gdrive_data, #"/IntermediateData/IntermediateData_Files/distance_rasters/road_shapefile", #name), filetype = "ESRI Shapefile")
#}

#save the file
#savevect(roads_2015_mask)

```

```{r}
saverast <- function(x){
  
  name <- paste0(deparse(substitute(x)), ".tif") #the name of the file will be the same as the input object, plus '.tif' at the end
  
#write the raster into the correct folder in drive (the path will need to be updated for other files) and save it as a geotiff
  
  writeRaster(x, filename = here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/road_rasters", name), filetype = "GTiff", overwrite=TRUE)
}

saverast(roads_2015_mask)
```



##Distance to streams

```{r}
distance_to_streams <- distance(streams_2017_mask)

```

##Save to Google Drive
```{r}
saverast <- function(x){
  
  name <- paste0(deparse(substitute(x)), ".tif") #the name of the file will be the same as the input object, plus '.tif' at the end
  
#write the raster into the correct folder in drive (the path will need to be updated for other files) and save it as a geotiff
  
  writeRaster(x, filename = here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/stream_rasters", name), filetype = "GTiff")
}

saverast(streams_2017_raster)

```


##Distance to urban areas

```{r}
#creation of binary urban area rasters from NLCD data (urban = 1, all other cover = NA)

#read in land cover rasters - we will convert these to hold only developed areas

lc01 <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/NLCD_Rasters_Formatted/lc01_reclassed.tif"))

lc04 <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/NLCD_Rasters_Formatted/lc04_reclassed.tif"))

lc06 <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/NLCD_Rasters_Formatted/lc06_reclassed.tif"))

lc08 <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/NLCD_Rasters_Formatted/lc08_reclassed.tif"))

lc11 <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/NLCD_Rasters_Formatted/lc11_reclassed.tif"))

lc13 <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/NLCD_Rasters_Formatted/lc13_reclassed.tif"))

lc16 <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/NLCD_Rasters_Formatted/lc16_reclassed.tif"))

lc19 <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/NLCD_Rasters_Formatted/lc19_reclassed.tif"))

#list out the from, to values in an object
m <- c(0, NA,
       11, NA, 
       12, NA, 
       21, 1, #developed open space
       22, 1, #developed low intensity
       23, 1, #developed med intensity
       24, 1, #developed high intensity 
       31, NA, 
       40, NA,
       52, NA, 
       71, NA, 
       80, NA, 
       90, NA)

#turn those values into a matrix
mat <- matrix(m, ncol = 2, byrow = TRUE) 

#set categories (labels) for the numeric index

#index vector
index <- c(1)

#categories vector
cats <- c("urban")

#combine those into a dataframe
catdat <- data.frame(id = index, category = cats)

#classification function

reclassify <- function(x){
  
  x_reclassed <- classify(x, mat) #reclassify using the matrix above, set all non-forest values to NA
  
  levels(x_reclassed) <- catdat #then reassign levels (labels) using the dataframe above
  
  return(x_reclassed)
}

#reclassify lc rasters to urban areas

ua01 <- reclassify(lc01)

ua04 <- reclassify(lc04)

ua06 <- reclassify(lc06)

ua08 <- reclassify(lc08)

ua11 <- reclassify(lc11)

ua13 <- reclassify(lc13)

ua16 <- reclassify(lc16)

ua19 <- reclassify(lc19)
```

##Save reclassified, reprojected, and clipped urban rasters to the data drive for use in ArcGIS


**ONLY RUN THIS ONE TIME; IT WILL NOT RUN ONCE THE FILES ARE SAVED**

```{r}
saverast <- function(x){
  
  name <- paste0(deparse(substitute(x)), ".tif") #the name of the file will be the same as the input object, plus '.tif' at the end
  
#write the raster into the correct folder in drive (the path will need to be updated for other files) and save it as a geotiff
  
  writeRaster(x, filename = here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/urban_rasters", name), filetype = "GTiff")
}

saverast(ua01)

saverast(ua04)

saverast(ua06)

saverast(ua08)

saverast(ua11)

saverast(ua13)

saverast(ua16)

saverast(ua19)

```

##Distance to recreational areas

```{r}

```

##Distance to forest cover

```{r}
#creation of binary forest rasters from NLCD data (forest cover =1, all other cover = NA)

#read in land cover rasters - we will convert these to hold only forest cover

lc01 <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/NLCD_Rasters_Formatted/lc01_reclassed.tif"))

lc04 <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/NLCD_Rasters_Formatted/lc04_reclassed.tif"))

lc06 <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/NLCD_Rasters_Formatted/lc06_reclassed.tif"))

lc08 <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/NLCD_Rasters_Formatted/lc08_reclassed.tif"))

lc11 <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/NLCD_Rasters_Formatted/lc11_reclassed.tif"))

lc13 <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/NLCD_Rasters_Formatted/lc13_reclassed.tif"))

lc16 <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/NLCD_Rasters_Formatted/lc16_reclassed.tif"))

lc19 <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/NLCD_Rasters_Formatted/lc19_reclassed.tif"))

#list out the from, to values in an object
m <- c(0, NA,
       11, NA, 
       12, NA, 
       21, NA, 
       22, NA, 
       23, NA, 
       24, NA, 
       31, NA, 
       40, 40, #forest
       52, NA, 
       71, NA, 
       80, NA, 
       90, NA)

#turn those values into a matrix
mat <- matrix(m, ncol = 2, byrow = TRUE) 

#set categories (labels) for the numeric index

#index vector
index <- c(40)

#categories vector
cats <- c("forest")

#combine those into a dataframe
catdat <- data.frame(id = index, category = cats)

#classification function

reclassify <- function(x){
  
  x_reclassed <- classify(x, mat) #reclassify using the matrix above, set all non-forest values to NA
  
  levels(x_reclassed) <- catdat #then reassign levels (labels) using the dataframe above
  
  return(x_reclassed)
}

#reclassify lc rasters to forest cover and then resample to correct resolution

fc01 <- reclassify(lc01)

fc04 <- reclassify(lc04)

fc06 <- reclassify(lc06)

fc08 <- reclassify(lc08)

fc11 <- reclassify(lc11)

fc13 <- reclassify(lc13)

fc16 <- reclassify(lc16)

fc19 <- reclassify(lc19)

```

##Save reclassified, reprojected, and clipped forest rasters to the data drive for use in ArcGIS

**ONLY RUN THIS ONE TIME; IT WILL NOT RUN ONCE THE FILES ARE SAVED**

```{r}
saverast <- function(x){
  
  name <- paste0(deparse(substitute(x)), ".tif") #the name of the file will be the same as the input object, plus '.tif' at the end
  
#write the raster into the correct folder in drive (the path will need to be updated for other files) and save it as a geotiff
  
  writeRaster(x, filename = here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/forest_rasters", name), filetype = "GTiff")
}

saverast(fc01)

saverast(fc04)

saverast(fc06)

saverast(fc08)

saverast(fc11)

saverast(fc13)

saverast(fc16)

saverast(fc19)

```