---
title: "Distance Raster Creation"
author: "Mia Guarnieri"
date: '2022-10-19'
output: html_document
---

##Setup and packages

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#attach packages
#note - Loading all of the ones Jason used, I don't think we will need all of them 

library(tidyverse)
library(here)
library(terra)
library(sf)
library(tmap)
library(rgrass7)


#mia and claire's and chase's file path
gdrive_data <- "/Volumes/GoogleDrive/Shared drives/Black_Bear_Aware/gdrive_data"

#sam's file path

# gdrive_data <- "G:/Shared drives/Black_Bear_Aware/gdrive_data"
```

#Data Wrangling Prior to Coding

*Data Sources*: Please see ReadMe file for metadata and citations.
*Reference Paper*: Insert hagani citation here

#Load in existing data layers

##California raster

Insert info about CA raster layer here

```{r}

ca_raster <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/CARaster_Tiger_Formatted/formattedcaraster.tif"))

#check the coordinate reference system 
crs(ca_raster)

#plot CA raster to confirm
tm_shape(ca_raster) +
  tm_raster() +
  tmap_mode("view")

res(ca_raster)

```

#Streams Layer 
```{r}
streams_2017 <- vect(here(gdrive_data, "/InputData/InputData_Files/2017_Streams_CDFW/California_Streams.shp"))

#ca_extent <- ext(ca_raster)

streams_2017_calbers <- project(streams_2017, ca_raster)

ca_vector <- as.polygons(ca_raster)

streams_2017_mask <- crop(streams_2017_calbers, ca_vector)

streams_2017_raster <- rasterize(streams_2017_mask, ca_raster)
```


#Roads Layer 

```{r}
roads_2015 <- vect(here(gdrive_data, "/InputData/InputData_Files/2015_roads_TIGER/tl_2015_06_prisecroads.shp"))

roads_2015_calbers <- project(roads_2015, ca_raster)

roads_2015_mask <- crop(roads_2015_calbers, ca_raster)

roads_2015_raster <- rasterize(roads_2015_mask, ca_raster)

```



#Create remaining data layers

```{r}

```


##Distance to road


```{r}
savevect <- function(x){
  
  name <- paste0(deparse(substitute(x)), ".shp") #the name of the file will be the same as the input object, plus '.hp' at the end
  
#write the raster into the correct folder in drive (the path will need to be updated for other files) and save it as a shapefile
  
  writeVector(x, filename = here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/road_shapefile", name), filetype = "ESRI Shapefile", overwrite=TRUE)
}

#save the file
savevect(roads_2015_mask)

```

```{r}
saverast <- function(x){
  
  name <- paste0(deparse(substitute(x)), ".tif") #the name of the file will be the same as the input object, plus '.tif' at the end
  
#write the raster into the correct folder in drive (the path will need to be updated for other files) and save it as a geotiff
  
  writeRaster(x, filename = here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/road_rasters", name), filetype = "GTiff", overwrite=TRUE)
}

saverast(roads_2015_raster)
```



##Distance to streams
#Save vector to Gdrive
```{r}
savevect <- function(x){
  
  name <- paste0(deparse(substitute(x)), ".shp") #the name of the file will be the same as the input object, plus '.hp' at the end
  
#write the raster into the correct folder in drive (the path will need to be updated for other files) and save it as a shapefile
  
  writeVector(x, filename = here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/stream_shapefile", name), filetype = "ESRI Shapefile", overwrite=TRUE)
}

#save the file
savevect(streams_2017_mask)
```


##Save raster to Google Drive
```{r}
saverast <- function(x){
  
  name <- paste0(deparse(substitute(x)), ".tif") #the name of the file will be the same as the input object, plus '.tif' at the end
  
#write the raster into the correct folder in drive (the path will need to be updated for other files) and save it as a geotiff
  
  writeRaster(x, filename = here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/stream_rasters", name), filetype = "GTiff", overwrite=TRUE)
}

saverast(streams_2017_raster)

```


##Distance to urban areas

```{r}
#creation of binary urban area rasters from NLCD data (urban = 1, all other cover = NA)

#read in land cover rasters - we will convert these to hold only developed areas

lc13 <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/NLCD_Rasters_Formatted/lc13_reclassed.tif"))

lc16 <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/NLCD_Rasters_Formatted/lc16_reclassed.tif"))

lc19 <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/NLCD_Rasters_Formatted/lc19_reclassed.tif"))

#don't use these, they are outside our time frame

#lc01 <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/NLCD_Rasters_Formatted/lc01_reclassed.tif"))

#lc04 <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/NLCD_Rasters_Formatted/lc04_reclassed.tif"))

#lc06 <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/NLCD_Rasters_Formatted/lc06_reclassed.tif"))

#lc08 <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/NLCD_Rasters_Formatted/lc08_reclassed.tif"))

#lc11 <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/NLCD_Rasters_Formatted/lc11_reclassed.tif"))


#list out the from, to values in an object
m <- c(0, NA,
       11, NA, 
       12, NA, 
       21, 1, #developed open space
       22, 1, #developed low intensity
       23, 1, #developed med intensity
       24, 1, #developed high intensity 
       31, NA, 
       40, NA,
       52, NA, 
       71, NA, 
       80, NA, 
       90, NA)

#turn those values into a matrix
mat <- matrix(m, ncol = 2, byrow = TRUE) 

#set categories (labels) for the numeric index

#index vector
index <- c(1)

#categories vector
cats <- c("urban")

#combine those into a dataframe
catdat <- data.frame(id = index, category = cats)

#classification function

reclassify <- function(x){
  
  x_reclassed <- classify(x, mat) #reclassify using the matrix above, set all non-forest values to NA
  
  levels(x_reclassed) <- catdat #then reassign levels (labels) using the dataframe above
  
  return(x_reclassed)
}

#reclassify lc rasters to urban areas

ua13 <- reclassify(lc13)

ua16 <- reclassify(lc16)

ua19 <- reclassify(lc19)

#these are outside our time frame

#ua01 <- reclassify(lc01)

#ua04 <- reclassify(lc04)

#ua06 <- reclassify(lc06)

#ua08 <- reclassify(lc08)

#ua11 <- reclassify(lc11)



```

##Save reclassified, reprojected, and clipped urban rasters to the data drive for use in ArcGIS


**ONLY RUN THIS ONE TIME; IT WILL NOT RUN ONCE THE FILES ARE SAVED**

```{r}
saverast <- function(x){
  
  name <- paste0(deparse(substitute(x)), ".tif") #the name of the file will be the same as the input object, plus '.tif' at the end
  
#write the raster into the correct folder in drive (the path will need to be updated for other files) and save it as a geotiff
  
  writeRaster(x, filename = here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/urban_rasters", name), filetype = "GTiff", overwrite=TRUE)
}

saverast(ua13)

saverast(ua16)

saverast(ua19)

#outside our time frame

#saverast(ua01)

#saverast(ua04)

#saverast(ua06)

#saverast(ua08)

#saverast(ua11)

#these saved raster layers were processed in ArcGIS Pro using the Euclidean distance tool 
```


## reading urban distance rasters and masking them to CA
```{r}
#Now we need to read back in the distance raster created in ArcGIS pro and mask these layers to match our CA extent again. Layers are already in the correct projection and resolution

dist_ua19 <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/dist_to_urban/urban19_dist.tif"))

dist_ua16 <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/dist_to_urban/urban16_dist.tif"))

dist_ua13 <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/dist_to_urban/urban13_dist.tif"))

# Now we need to mask these layers to just the CA extent 
ua19_ca_distance <- mask(dist_ua19, ca_raster)

ua16_ca_distance <- mask(dist_ua16, ca_raster)

ua13_ca_distance <- mask(dist_ua13, ca_raster)

#now we will save them back to the google drive to be referenced later 

```
##Save clipped urban DISTANCE rasters to the data drive for use in model and easy access


**ONLY RUN THIS ONE TIME; IT WILL NOT RUN ONCE THE FILES ARE SAVED**

```{r}
saverast <- function(x){
  
  name <- paste0(deparse(substitute(x)), ".tif") #the name of the file will be the same as the input object, plus '.tif' at the end
  
#write the raster into the correct folder in drive (the path will need to be updated for other files) and save it as a geotiff
  
  writeRaster(x, filename = here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/dist_to_urban/model_ready", name), filetype = "GTiff", overwrite=TRUE)
}

saverast(ua19_ca_distance)

saverast(ua16_ca_distance)

saverast(ua13_ca_distance)


```



##Distance to recreational areas

###Park Entry Points
```{r}
rec_areas_PEP <- vect(here(gdrive_data, "/InputData/InputData_Files/recreational_areas/parkentrypoints/ParkEntryPoints.shp"))

rec_areas_PEP_calbers <- project(rec_areas_PEP, ca_raster)

rec_areas_PEP_mask <- crop(rec_areas_PEP_calbers, ca_raster)

rec_areas_PEP_raster <- rasterize(rec_areas_PEP_mask, ca_raster)

```



###Parking Points
```{r}
rec_areas_parking <- vect(here(gdrive_data, "/InputData/InputData_Files/recreational_areas/parkingpoints/ParkingPoints.shp"))

rec_areas_parking_calbers <- project(rec_areas_parking, ca_raster)

rec_areas_parking_mask <- crop(rec_areas_parking_calbers, ca_raster)

rec_areas_parking_raster <- rasterize(rec_areas_parking_mask, ca_raster)


```


###Campgrounds
```{r}
rec_areas_camp <- vect(here(gdrive_data, "/InputData/InputData_Files/recreational_areas/campgrounds/Campgrounds.shp"))

rec_areas_camp_calbers <- project(rec_areas_camp, ca_raster)

rec_areas_camp_mask <- crop(rec_areas_camp_calbers, ca_raster)

rec_areas_camp_raster <- rasterize(rec_areas_camp_mask, ca_raster)
```

###Picnic Grounds
```{r}
rec_areas_picnic <- vect(here(gdrive_data, "/InputData/InputData_Files/recreational_areas/picnicgrounds/PicnicGrounds.shp"))

rec_areas_picnic_calbers <- project(rec_areas_picnic, ca_raster)

rec_areas_picnic_mask <- crop(rec_areas_picnic_calbers, ca_raster)

rec_areas_picnic_raster <- rasterize(rec_areas_picnic_mask, ca_raster)
```

###Structures
```{r}
rec_areas_structures <- vect(here(gdrive_data, "/InputData/InputData_Files/recreational_areas/structures/Structures.shp"))

rec_areas_structures_calbers <- project(rec_areas_structures, ca_raster)

rec_areas_structures_mask <- crop(rec_areas_structures_calbers, ca_raster)

rec_areas_structures_raster <- rasterize(rec_areas_structures_mask, ca_raster)
```

###Buildings
```{r}
rec_areas_buildings <- vect(here(gdrive_data, "/InputData/InputData_Files/recreational_areas/buildings/Buildings.shp"))

rec_areas_buildings_calbers <- project(rec_areas_buildings, ca_raster)

rec_areas_buildings_mask <- crop(rec_areas_buildings_calbers, ca_raster)

rec_areas_buildings_raster <- rasterize(rec_areas_buildings_mask, ca_raster)
```

###Recreational Routes
```{r}
rec_areas_rr <- vect(here(gdrive_data, "/InputData/InputData_Files/recreational_areas/recreationalroutes/RecreationalRoutes.shp"))

rec_areas_rr_calbers <- project(rec_areas_rr, ca_raster)

rec_areas_rr_mask <- crop(rec_areas_rr_calbers, ca_raster)

rec_areas_rr_raster <- rasterize(rec_areas_rr_mask, ca_raster)
```

###Park Boundaries
```{r}
rec_areas_pb <- vect(here(gdrive_data, "/InputData/InputData_Files/recreational_areas/parkboundaries/ParkBoundaries.shp"))

rec_areas_pb_calbers <- project(rec_areas_pb, ca_raster)

rec_areas_pb_mask <- crop(rec_areas_pb_calbers, ca_raster)

rec_areas_pb_raster <- rasterize(rec_areas_pb_mask, ca_raster)
```

#####Saving Rec Areas to Distance Rasters
```{r}
savevect <- function(x){
  
  name <- paste0(deparse(substitute(x)), ".shp") #the name of the file will be the same as the input object, plus '.tif' at the end
  
#write the raster into the correct folder in drive (the path will need to be updated for other files) and save it as a geotiff
  
  writeVector(x, filename = here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/recreational_rasters", name), filetype = "ESRI SHAPEFILE", overwrite=TRUE)
}

savevect(rec_areas_pb_mask)
savevect(rec_areas_rr_mask)
savevect(rec_areas_buildings_mask)
savevect(rec_areas_structures_mask)
savevect(rec_areas_parking_mask)
savevect(rec_areas_PEP_mask)
savevect(rec_areas_picnic_mask)
savevect(rec_areas_camp_mask)

#merging layers 

rec_areas_ca <- merge(rec_areas_PEP_raster, rec_areas_pb_raster, rec_areas_rr_raster, rec_areas_buildings_raster, rec_areas_structures_raster, rec_areas_picnic_raster, rec_areas_camp_raster, rec_areas_parking_raster)


```

```{r}
saverast <- function(x){
  
  name <- paste0(deparse(substitute(x)), ".shp") #the name of the file will be the same as the input object, plus '.tif' at the end
  
#write the raster into the correct folder in drive (the path will need to be updated for other files) and save it as a geotiff
  
  writeRaster(x, filename = here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/recreational_rasters", name), filetype = "Gtiff", overwrite=TRUE)
}

saverast(rec_areas_ca)
```

##Distance to forest cover

```{r}
#creation of binary forest rasters from NLCD data (forest cover =1, all other cover = NA)

#read in land cover rasters - we will convert these to hold only forest cover

lc13 <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/NLCD_Rasters_Formatted/lc13_reclassed.tif"))

lc16 <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/NLCD_Rasters_Formatted/lc16_reclassed.tif"))

lc19 <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/NLCD_Rasters_Formatted/lc19_reclassed.tif"))

#these ones are outside our time frame

#lc01 <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/NLCD_Rasters_Formatted/lc01_reclassed.tif"))

#lc04 <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/NLCD_Rasters_Formatted/lc04_reclassed.tif"))

#lc06 <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/NLCD_Rasters_Formatted/lc06_reclassed.tif"))

#lc08 <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/NLCD_Rasters_Formatted/lc08_reclassed.tif"))

#lc11 <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/NLCD_Rasters_Formatted/lc11_reclassed.tif"))


#list out the from, to values in an object
m <- c(0, NA,
       11, NA, 
       12, NA, 
       21, NA, 
       22, NA, 
       23, NA, 
       24, NA, 
       31, NA, 
       40, 40, #forest
       52, NA, 
       71, NA, 
       80, NA, 
       90, NA)

#turn those values into a matrix
mat <- matrix(m, ncol = 2, byrow = TRUE) 

#set categories (labels) for the numeric index

#index vector
index <- c(40)

#categories vector
cats <- c("forest")

#combine those into a dataframe
catdat <- data.frame(id = index, category = cats)

#classification function

reclassify <- function(x){
  
  x_reclassed <- classify(x, mat) #reclassify using the matrix above, set all non-forest values to NA
  
  levels(x_reclassed) <- catdat #then reassign levels (labels) using the dataframe above
  
  return(x_reclassed)
}

#reclassify lc rasters to forest cover and then resample to correct resolution

fc13 <- reclassify(lc13)

fc16 <- reclassify(lc16)

fc19 <- reclassify(lc19)

#outside our time frame

#fc01 <- reclassify(lc01)

#fc04 <- reclassify(lc04)

#fc06 <- reclassify(lc06)

#fc08 <- reclassify(lc08)

#fc11 <- reclassify(lc11)

#Notes on distance rasters

# fc13_lowres <- terra::aggregate(fc13, 5, fun = "sum", na.rm = TRUE) - make the distance raster lower resolution
# 
# fc13_lowres_dist <- distance(fc13_lowres) #then the distance would run

#system.time({x <- distance(blah)}) to see the runtime/cpu time used

```

##Save reclassified, reprojected, and clipped forest rasters to the data drive for use in ArcGIS

**ONLY RUN THIS ONE TIME; IT WILL NOT RUN ONCE THE FILES ARE SAVED**

```{r}
saverast <- function(x){
  
  name <- paste0(deparse(substitute(x)), ".tif") #the name of the file will be the same as the input object, plus '.tif' at the end
  
#write the raster into the correct folder in drive (the path will need to be updated for other files) and save it as a geotiff
  
  writeRaster(x, filename = here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/forest_rasters", name), filetype = "GTiff", overwrite=TRUE)
}

saverast(fc13)

saverast(fc16)

saverast(fc19)

#outside our time frame

#saverast(fc01)

#saverast(fc04)

#saverast(fc06)

#saverast(fc08)

#saverast(fc11)

```


## reading forest cover distance rasters and masking them to CA
```{r}
#Now we need to read back in the distance raster created in ArcGIS pro and mask these layers to match our CA extent again. Layers are already in the correct projection and resolution

dist_fc19 <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/dist_to_forest/dist_fc19.tif"))

dist_fc16 <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/dist_to_forest/dist_fc16.tif"))

dist_fc13 <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/dist_to_forest/dist_fc13.tif"))

#added step, was having issue with distance raster from arc not matching ca_raster extent. This work around fixes the error 
dist_19_mask = resample(dist_fc19, ca_raster)

dist_16_mask = resample(dist_fc16, ca_raster)

dist_13_mask = resample(dist_fc13, ca_raster)

# Now we need to mask these layers to just the CA extent 
fc19_ca_distance <- mask(dist_19_mask, ca_raster)

fc16_ca_distance <- mask(dist_16_mask, ca_raster)

fc13_ca_distance <- mask(dist_13_mask, ca_raster)

#now we will save them back to the google drive to be referenced later 

```
##Save clipped forest cover DISTANCE rasters to the data drive for use in model and easy access


**ONLY RUN THIS ONE TIME; IT WILL NOT RUN ONCE THE FILES ARE SAVED**

```{r}
saverast <- function(x){
  
  name <- paste0(deparse(substitute(x)), ".tif") #the name of the file will be the same as the input object, plus '.tif' at the end
  
#write the raster into the correct folder in drive (the path will need to be updated for other files) and save it as a geotiff
  
  writeRaster(x, filename = here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/dist_to_forest/model_ready", name), filetype = "GTiff", overwrite=TRUE)
}

saverast(fc19_ca_distance )

saverast(fc16_ca_distance )

saverast(fc13_ca_distance )




```

