---
title: "Distance Raster Creation"
author: "Mia Guarnieri"
date: '2022-10-19'
output: html_document
---

##Setup and packages

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#attach packages
#note - Loading all of the ones Jason used, I don't think we will need all of them 

library(tidyverse)
library(here)
library(terra)
library(sf)
library(tmap)

#mia and claire's file path
#gdrive_data <- "/Volumes/GoogleDrive/Shared drives/Black_Bear_Aware/gdrive_data"

#sam's file path
gdrive_data <- "G:/Shared drives/Black_Bear_Aware/gdrive_data"

#chase's file path

```

#Data Wrangling Prior to Coding

*Data Sources*: Please see ReadMe file for metadata and citations.
*Reference Paper*: Insert hagani citation here

#Load in existing data layers

##California raster

Insert info about CA raster layer here

```{r}

ca_raster <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/CARaster_Tiger_Reclassed/CARaster_Tiger_Reclassed.img"))

#check the coordinate reference system 
crs(ca_raster)

#plot CA raster to confirm
tm_shape(ca_raster) +
  tm_raster() +
  tmap_mode("view")

```

#Streams Layer 
```{r}
streams_2017 <- vect(here(gdrive_data, "/InputData/InputData_Files/2017_Streams_CDFW/California_Streams.shp"))

streams_2017_calbers <- project(streams_2017, ca_raster)


```

#Roads Layer 

```{r}
roads_2015 <- vect(here(gdrive_data, "/InputData/InputData_Files/2015_roads_TIGER/tl_2015_06_prisecroads.shp"))

roads_2015_calbers <- project(roads_2015, ca_raster)
```


#Create remaining data layers

```{r}

```


##Distance to road

```{r}
distance_to_roads <- distance(roads_2015_calbers)

```

##Distance to streams

```{r}
distance_to_streams <- distance(streams_2017_calbers)

```

##Distance to urban areas

```{r}

```

##Distance to recreational areas

```{r}

```

##Distance to forest cover

```{r}


```
