---
title: "Fire Data Wrangling"
author: "Mia Guarnieri"
date: '2022-11-02'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#attach packages
#note - Loading all of the ones Jason used, I don't think we will need all of them 

library(tidyverse)
library(here)
library(terra)
library(sf)
library(tmap)
library(beepr)
library(rgeos)
library(rmapshaper)
#library(rgdal) fyi rgdal has project function as well and will mess up terra project function

#mia and claire's file path
gdrive_data <- "/Volumes/GoogleDrive/Shared drives/Black_Bear_Aware/gdrive_data"

#sam's file path
#gdrive_data <- "G:/Shared drives/Black_Bear_Aware/gdrive_data"

#chase's file path
```


```{r}
# CA Raster example 
ca_raster <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/CARaster_Tiger_Formatted/formattedcaraster.tif"))

#check the coordinate reference system 
crs(ca_raster)


#plot CA raster to confirm
#tm_shape(ca_raster) +
 # tm_raster() +
#  tmap_mode("view")
```

# Initial fire data layers

Rasters showing fire location and severity, from 0-4

```{r}
#read in fire layers 

ca_fire_2011 <- rast(here(gdrive_data, "/InputData/InputData_Files/fire_intensity_CA_mtbs/2011_fire_intensity_CA_mtbs/mtbs_CA_2011.tif"))

ca_fire_2012 <- rast(here(gdrive_data, "/InputData/InputData_Files/fire_intensity_CA_mtbs/2012_fire_intensity_CA_mtbs/mtbs_CA_2012.tif"))

ca_fire_2013 <- rast(here(gdrive_data, "/InputData/InputData_Files/fire_intensity_CA_mtbs/2013_fire_intensity_CA_mtbs/mtbs_CA_2013.tif"))

ca_fire_2014 <- rast(here(gdrive_data, "/InputData/InputData_Files/fire_intensity_CA_mtbs/2014_fire_intensity_CA_mtbs/mtbs_CA_2014.tif"))

ca_fire_2015 <- rast(here(gdrive_data, "/InputData/InputData_Files/fire_intensity_CA_mtbs/2015_fire_intensity_CA_mtbs/mtbs_CA_2015.tif"))

ca_fire_2016 <- rast(here(gdrive_data, "/InputData/InputData_Files/fire_intensity_CA_mtbs/2016_fire_intensity_CA_mtbs/mtbs_CA_2016.tif"))

ca_fire_2017 <- rast(here(gdrive_data, "/InputData/InputData_Files/fire_intensity_CA_mtbs/2017_fire_intensity_CA_mtbs/mtbs_CA_2017.tif"))

ca_fire_2018 <- rast(here(gdrive_data, "/InputData/InputData_Files/fire_intensity_CA_mtbs/2018_fire_intensity_CA_mtbs/mtbs_CA_2018.tif"))

ca_fire_2019 <- rast(here(gdrive_data, "/InputData/InputData_Files/fire_intensity_CA_mtbs/2019_fire_intensity_CA_mtbs/mtbs_CA_2019.tif"))

ca_fire_2020 <- rast(here(gdrive_data, "/InputData/InputData_Files/fire_intensity_CA_mtbs/2020_fire_intensity_CA_mtbs/mtbs_CA_2020.tif"))

```


#Getting layers in correct extent, resolution and projection

```{r}
#reproject to the same crs (NAD 83, CA Albers (2011))
# method = near in order to preserve categorical data

ca_fire_2011_albers <- terra::project(ca_fire_2011, ca_raster, method = "near")

ca_fire_2012_albers <- terra::project(ca_fire_2012, ca_raster, method = "near")

ca_fire_2013_albers <- terra::project(ca_fire_2013, ca_raster, method = "near")

ca_fire_2014_albers <- terra::project(ca_fire_2014, ca_raster, method = "near")

ca_fire_2015_albers <- terra::project(ca_fire_2015, ca_raster, method = "near")

ca_fire_2016_albers <- terra::project(ca_fire_2016, ca_raster, method = "near")

ca_fire_2017_albers <- terra::project(ca_fire_2017, ca_raster, method = "near")

ca_fire_2018_albers <- terra::project(ca_fire_2018, ca_raster, method = "near")

ca_fire_2019_albers <- terra::project(ca_fire_2019, ca_raster, method = "near")

ca_fire_2020_albers <- terra::project(ca_fire_2020, ca_raster, method = "near")

```

```{r}
#cropping to spatial extent of CA raster

fire_2011_mask <- mask(ca_fire_2011_albers, ca_raster)

fire_2012_mask <- mask(ca_fire_2012_albers, ca_raster)

fire_2013_mask <- mask(ca_fire_2013_albers, ca_raster)

fire_2014_mask <- mask(ca_fire_2014_albers, ca_raster)

fire_2015_mask <- mask(ca_fire_2015_albers, ca_raster)

fire_2016_mask <- mask(ca_fire_2016_albers, ca_raster)

fire_2017_mask <- mask(ca_fire_2017_albers, ca_raster)

fire_2018_mask <- mask(ca_fire_2018_albers, ca_raster)

fire_2019_mask <- mask(ca_fire_2019_albers, ca_raster)

fire_2020_mask <- mask(ca_fire_2020_albers, ca_raster)
```

#Save reprojected and clipped rasters to the data drive

```{r}
saverast <- function(x){
  
  name <- paste0(deparse(substitute(x)), ".tif") #the name of the file will be the same as the input object, plus '.tif' at the end
  
  #write the raster into the correct folder in drive (the path will need to be updated for other files) and save it as a geotiff
  
  writeRaster(x, filename = here(gdrive_data, "/IntermediateData/IntermediateData_Files/model_ready/fire_intensity", name), filetype = "GTiff", overwrite = TRUE)
  
}

saverast(fire_2011_mask)

saverast(fire_2012_mask)

saverast(fire_2013_mask)

saverast(fire_2014_mask)

saverast(fire_2015_mask)

saverast(fire_2016_mask)

saverast(fire_2017_mask)

saverast(fire_2018_mask)

saverast(fire_2019_mask)

saverast(fire_2020_mask)

```

#Read in saved, formatted layers

```{r}
#read in fire layers 

fire_2011_mask <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/model_ready/fire_intensity/fire_2011_mask.tif"))

fire_2012_mask <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/model_ready/fire_intensity/fire_2012_mask.tif"))

fire_2013_mask <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/model_ready/fire_intensity/fire_2013_mask.tif"))

fire_2014_mask <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/model_ready/fire_intensity/fire_2014_mask.tif"))

fire_2015_mask <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/model_ready/fire_intensity/fire_2015_mask.tif"))

fire_2016_mask <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/model_ready/fire_intensity/fire_2016_mask.tif"))

fire_2017_mask <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/model_ready/fire_intensity/fire_2017_mask.tif"))

fire_2018_mask <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/model_ready/fire_intensity/fire_2018_mask.tif"))

fire_2019_mask <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/model_ready/fire_intensity/fire_2019_mask.tif"))

fire_2020_mask <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/model_ready/fire_intensity/fire_2020_mask.tif"))

```


#Reclassify to keep only medium and high fire intensity

```{r}
#list out the from, to values in an object
m <- c(1, 0, #Unburned/Underburned/Low Burn Severity to no data
       2, 0, #Low burn severity to no data
       3, 4, #moderate burn severity to moderate/high burn severity
       4, 4, #high burn severity to moderate/high burn severity
       5, 0, #Increased Greenness/Increased Vegetation Response to no data
       6,0) #Non-Processing Area Mask to no data
       
#turn those values into a matrix
mat_fire <- matrix(m, ncol = 2, byrow = TRUE) 

#set categories (labels) for each numeric index

#index vector
index_fire <- c(0, 4)

#categories vector
cats_fire <- c("undefined","mod_high_burn")

#combine those into a dataframe
catdat_fire <- data.frame(id = index_fire, category = cats_fire)

#Reclassify

#create reclassification function

reclassify_fire <- function(x){
  fire_reclassed <- classify(x, mat_fire) #reclassify using the matrix above
  
  levels(fire_reclassed) <- catdat_fire #then reassign levels (labels) using the dataframe above
  
  return(fire_reclassed)
}

#Apply function to each layer

fire_2011_reclassed <- reclassify_fire(fire_2011_mask)

fire_2012_reclassed <- reclassify_fire(fire_2012_mask)

fire_2013_reclassed <- reclassify_fire(fire_2013_mask)

fire_2014_reclassed <- reclassify_fire(fire_2014_mask)

fire_2015_reclassed <- reclassify_fire(fire_2015_mask)

fire_2016_reclassed <- reclassify_fire(fire_2016_mask)

fire_2017_reclassed <- reclassify_fire(fire_2017_mask)

fire_2018_reclassed <- reclassify_fire(fire_2018_mask)

fire_2019_reclassed <- reclassify_fire(fire_2019_mask)

fire_2020_reclassed <- reclassify_fire(fire_2020_mask)

```


#Save reclassified rasters to the data drive

```{r}
saverast <- function(x){
  
  name <- paste0(deparse(substitute(x)), ".tif") #the name of the file will be the same as the input object, plus '.tif' at the end
  
  #write the raster into the correct folder in drive (the path will need to be updated for other files) and save it as a geotiff
  
  writeRaster(x, filename = here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/med_high_fire_rasters", name), filetype = "GTiff", overwrite = TRUE)
  
}

saverast(fire_2011_reclassed)
saverast(fire_2012_reclassed)
saverast(fire_2013_reclassed)
saverast(fire_2014_reclassed)
saverast(fire_2015_reclassed)
saverast(fire_2016_reclassed)
saverast(fire_2017_reclassed)
saverast(fire_2018_reclassed)
saverast(fire_2019_reclassed)
saverast(fire_2020_reclassed)

```


#Create multi-year merged layers for model data processing

```{r}
#function for saving multi-year rasters
savemultirast <- function(x){
  
  name <- paste0(deparse(substitute(x)), ".tif") #the name of the file will be the same as the input object, plus '.tif' at the end
  
  #write the raster into the correct folder in drive (the path will need to be updated for other files) and save it as a geotiff
  
  writeRaster(x, filename = here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/med_high_fire_rasters/multiyear_rasters", name), filetype = "GTiff", overwrite = TRUE)
  
}
```

##2016

```{r}
#4-5 year: combined 2011-2012

fire_2011 <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/med_high_fire_rasters/fire_2011_reclassed.tif"))

fire_2012 <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/med_high_fire_rasters/fire_2012_reclassed.tif"))

#merge those layers

fire_2011_2012 <- merge(fire_2011, fire_2012)


#2-3 year: combined 2013-2014

fire_2013 <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/med_high_fire_rasters/fire_2013_reclassed.tif"))

fire_2014 <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/med_high_fire_rasters/fire_2014_reclassed.tif"))

#merge those layers

fire_2013_2014 <- merge(fire_2013, fire_2014)

#save the layers

savemultirast(fire_2011_2012)

savemultirast(fire_2013_2014)
```

##2017

```{r}
#4-5 year: 2012-2013

fire_2012 <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/med_high_fire_rasters/fire_2012_reclassed.tif"))

fire_2013 <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/med_high_fire_rasters/fire_2013_reclassed.tif"))

#merge those layers

fire_2012_2013 <- merge(fire_2012, fire_2013)

#2-3 year: 2014-2015

fire_2014 <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/med_high_fire_rasters/fire_2014_reclassed.tif"))

fire_2015 <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/med_high_fire_rasters/fire_2015_reclassed.tif"))

#merge those layers

fire_2014_2015 <- merge(fire_2014, fire_2015)

#save the layers

savemultirast(fire_2012_2013)

savemultirast(fire_2014_2015)

```

##2018

```{r}
#4-5 year: 2013-2014 - this already exists from prior data wrangling

#2-3 year: 2015-2016

fire_2015 <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/med_high_fire_rasters/fire_2015_reclassed.tif"))

fire_2016 <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/med_high_fire_rasters/fire_2016_reclassed.tif"))

#merge those layers

fire_2015_2016 <- merge(fire_2015, fire_2016)

#save the layer

savemultirast(fire_2015_2016)
```

##2019

```{r}
#4-5 year: 2014-2015 - this already exists from prior data wrangling

#2-3 year: 2016-2017

fire_2016 <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/med_high_fire_rasters/fire_2016_reclassed.tif"))

fire_2017 <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/med_high_fire_rasters/fire_2017_reclassed.tif"))

#merge those layers

fire_2016_2017 <- merge(fire_2016, fire_2017)

#save the layer

savemultirast(fire_2016_2017)

```


##2020

```{r}
#4-5 year: 2015-2016 - this already exists from prior data wrangling

#2-3 year: 2017-2018

fire_2017 <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/med_high_fire_rasters/fire_2017_reclassed.tif"))

fire_2018 <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/med_high_fire_rasters/fire_2018_reclassed.tif"))

#merge the layers

fire_2017_2018 <- merge(fire_2017, fire_2018)

#save the layer

savemultirast(fire_2017_2018)
```

##2021

```{r}
#4-5 year: 2016-2017 - this already exists from prior data wrangling

#2-3 years: 2018-2019

fire_2018 <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/med_high_fire_rasters/fire_2018_reclassed.tif"))

fire_2019 <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/med_high_fire_rasters/fire_2019_reclassed.tif"))

#merge the layers

fire_2018_2019 <- merge(fire_2018, fire_2019)

#save the layer

savemultirast(fire_2018_2019)

```

#Fire distance rasters were created in arc gis pro, reading back in here and readjusting extent

```{r}
#Now we need to read back in the distance raster created in ArcGIS pro and mask these layers to match our CA extent again. Layers are already in the correct projection and resolution

dist_fire16 <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/dist_to_med_high_fire/dist_fire_2016.tif"))

dist_fire17 <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/dist_to_med_high_fire/dist_fire_2017.tif"))

dist_fire18 <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/dist_to_med_high_fire/dist_fire_2018.tif"))

dist_fire19 <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/dist_to_med_high_fire/dist_fire_2019.tif"))

dist_fire20 <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/dist_to_med_high_fire/dist_fire_2020.tif"))

# Now we need to mask these layers to just the CA extent 
fire16_ca_distance <- mask(dist_fire16, ca_raster) 



#now we will save them back to the google drive to be referenced later
```

#Remask the distance rasters

```{r}

```

