---
title: "Fire Data Wrangling hist first draft "
author: "Mia Guarnieri"
date: '2022-11-02'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#attach packages
#note - Loading all of the ones Jason used, I don't think we will need all of them 

library(tidyverse)
library(here)
library(terra)
library(sf)
library(tmap)
library(beepr)
library(rgeos)
library(rmapshaper)
#library(rgdal) fyi rgdal has project function as well and will mess up terra project function

#mia and claire's file path
gdrive_data <- "/Users/clairemeuter/Library/CloudStorage/GoogleDrive-clairemeuter@ucsb.edu/Shared drives/Black_Bear_Aware/gdrive_data"

#sam's file path
#gdrive_data <- "G:/Shared drives/Black_Bear_Aware/gdrive_data"

```

# Purpose: 
The purpose of this .rmd is to create distance to fire layers as variables for use in our model. 
 
See our team's Dryad page for further information on the source data. 

This .rmd resulted in the creation of these distance rasters: 
- Distance to moderate to severe burn scars from 2015
- Distance to moderate to severe burn scars from 2016
- Distance to moderate to severe burn scars from 2017
- Distance to moderate to severe burn scars from 2018
- Distance to moderate to severe burn scars from 2019
- Distance to moderate to severe burn scars from 2020
dist_fire5_6
dist_fire6_7
dist_fire7_8
dist_fire8_9
dist_fire9_10
dist_fire10_11
dist_fire11_12
- Distance to moderate to severe burn scars from 2011 to 2012
- Distance to moderate to severe burn scars from 2012 to 2013
- Distance to moderate to severe burn scars from 2013 to 2014
- Distance to moderate to severe burn scars from 2014 to 2015
- Distance to moderate to severe burn scars from 2015 to 2016
- Distance to moderate to severe burn scars from 2016 to 2017
- Distance to moderate to severe burn scars from 2017 to 2018
- Distance to moderate to severe burn scars from 2018 to 2019

## Data wrangling for fire burn scar data 
```{r}
# CA Raster example 

#reading in ca_raster, a raster formatted to the resolution, crs, and spatial extent we plan to run our model over. 

ca_raster <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/CARaster_Tiger_Formatted/formattedcaraster.tif"))

#check the coordinate reference system 
crs(ca_raster)


#plot CA raster to confirm
#tm_shape(ca_raster) +
 # tm_raster() +
#  tmap_mode("view")
```

## Wrangling initial fire data layers

Rasters showing fire location and severity, from 0-4

```{r}
#read in fire layers 
ca_fire_2005 <- rast(here(gdrive_data, "/InputData/InputData_Files/fire_intensity_CA_mtbs/2005_fire_intensity_CA_mtbs/mtbs_CA_2005.tif"))

ca_fire_2006 <- rast(here(gdrive_data, "/InputData/InputData_Files/fire_intensity_CA_mtbs/2006_fire_intensity_CA_mtbs/mtbs_CA_2006.tif"))

ca_fire_2007 <- rast(here(gdrive_data, "/InputData/InputData_Files/fire_intensity_CA_mtbs/2007_fire_intensity_CA_mtbs/mtbs_CA_2007.tif"))

ca_fire_2008 <- rast(here(gdrive_data, "/InputData/InputData_Files/fire_intensity_CA_mtbs/2008_fire_intensity_CA_mtbs/mtbs_CA_2008.tif"))

ca_fire_2009 <- rast(here(gdrive_data, "/InputData/InputData_Files/fire_intensity_CA_mtbs/2009_fire_intensity_CA_mtbs/mtbs_CA_2009.tif"))

ca_fire_2010 <- rast(here(gdrive_data, "/InputData/InputData_Files/fire_intensity_CA_mtbs/2010_fire_intensity_CA_mtbs/mtbs_CA_2010.tif"))

ca_fire_2011 <- rast(here(gdrive_data, "/InputData/InputData_Files/fire_intensity_CA_mtbs/2011_fire_intensity_CA_mtbs/mtbs_CA_2011.tif"))

ca_fire_2012 <- rast(here(gdrive_data, "/InputData/InputData_Files/fire_intensity_CA_mtbs/2012_fire_intensity_CA_mtbs/mtbs_CA_2012.tif"))

ca_fire_2013 <- rast(here(gdrive_data, "/InputData/InputData_Files/fire_intensity_CA_mtbs/2013_fire_intensity_CA_mtbs/mtbs_CA_2013.tif"))

ca_fire_2014 <- rast(here(gdrive_data, "/InputData/InputData_Files/fire_intensity_CA_mtbs/2014_fire_intensity_CA_mtbs/mtbs_CA_2014.tif"))

ca_fire_2015 <- rast(here(gdrive_data, "/InputData/InputData_Files/fire_intensity_CA_mtbs/2015_fire_intensity_CA_mtbs/mtbs_CA_2015.tif"))

ca_fire_2016 <- rast(here(gdrive_data, "/InputData/InputData_Files/fire_intensity_CA_mtbs/2016_fire_intensity_CA_mtbs/mtbs_CA_2016.tif"))

ca_fire_2017 <- rast(here(gdrive_data, "/InputData/InputData_Files/fire_intensity_CA_mtbs/2017_fire_intensity_CA_mtbs/mtbs_CA_2017.tif"))

ca_fire_2018 <- rast(here(gdrive_data, "/InputData/InputData_Files/fire_intensity_CA_mtbs/2018_fire_intensity_CA_mtbs/mtbs_CA_2018.tif"))

ca_fire_2019 <- rast(here(gdrive_data, "/InputData/InputData_Files/fire_intensity_CA_mtbs/2019_fire_intensity_CA_mtbs/mtbs_CA_2019.tif"))

ca_fire_2020 <- rast(here(gdrive_data, "/InputData/InputData_Files/fire_intensity_CA_mtbs/2020_fire_intensity_CA_mtbs/mtbs_CA_2020.tif"))

fire_list <- list(ca_fire_2005, ca_fire_2006, ca_fire_2007, ca_fire_2008, ca_fire_2009, ca_fire_2010, ca_fire_2011)

target_crs <- crs(ca_raster)
#Reproject the rasters to ca_raster
reprojected_rasters <- lapply(fire_list, function(x) project(x, target_crs,  method = "near"))

#resample rasters to avoid mask issue 
resampled_rasters <- lapply(reprojected_rasters, function(x) resample(x,ca_raster))

#mask the rasters to ca_raster

masked_rasters <- lapply(resampled_rasters, mask, ca_raster)

# now i want to unlist these files by year 
unlisted_rasters <- unlist(masked_rasters)

years <- c(2005, 2006, 2007, 2008, 2009, 2010, 2011)

names(unlisted_rasters) <- paste("fire", years, "mask", sep = "_")


#unlisted_rasters["fire_2005_mask"]

#save the masked layers to the drive 


#specify output directory 
output_dir <- "/Users/clairemeuter/Library/CloudStorage/GoogleDrive-clairemeuter@ucsb.edu/Shared drives/Black_Bear_Aware/gdrive_data/IntermediateData/IntermediateData_Files/distance_rasters/med_high_fire_hist"

for (i in seq_along(masked_rasters)) {
  filename <- paste("fire", years[i], "masked.tif", sep = "_")
  filepath <- file.path(output_dir, filename)
  writeRaster(masked_rasters[[i]], filepath, filetype = "GTiff", overwrite=TRUE)
}


```

### Reclassify to keep only medium and high fire intensity
We are only interesting in how medium and high intensity fires influences human-black bear conflict, so we will reclassify to only keep medium and hihgh intensity fires. 

```{r}
#list out the from, to values in an object
m <- c(1, 0, #Unburned/Underburned/Low Burn Severity to no data
       2, 0, #Low burn severity to no data
       3, 4, #moderate burn severity to moderate/high burn severity
       4, 4, #high burn severity to moderate/high burn severity
       5, 0, #Increased Greenness/Increased Vegetation Response to no data
       6,0) #Non-Processing Area Mask to no data
       
#turn those values into a matrix
mat_fire <- matrix(m, ncol = 2, byrow = TRUE) 

#set categories (labels) for each numeric index

#index vector
index_fire <- c(0, 4)

#categories vector
cats_fire <- c("undefined","mod_high_burn")

#combine those into a dataframe
catdat_fire <- data.frame(id = index_fire, category = cats_fire)

#Reclassify

#create reclassification function

reclassify_fire <- function(x){
  fire_reclassed <- classify(x, mat_fire) #reclassify using the matrix above
  
  levels(fire_reclassed) <- catdat_fire #then reassign levels (labels) using the dataframe above
  
  return(fire_reclassed)
}

#Apply function to each layer to reclass each layer 

reclassed_rasters <- lapply(fire_list, reclassify_fire)


```

### Save reclassified rasters to the data drive

```{r}
# Save each raster from the reclassed_rasters list

#specify output directory 
output_dir <- "/Users/clairemeuter/Library/CloudStorage/GoogleDrive-clairemeuter@ucsb.edu/Shared drives/Black_Bear_Aware/gdrive_data/IntermediateData/IntermediateData_Files/distance_rasters/med_high_fire_hist"

for (i in seq_along(reclassed_rasters)) {
  filename <- paste("fire", years[i], "reclassed.tif", sep = "_")
  filepath <- file.path(output_dir, filename)
  writeRaster(masked_rasters[[i]], filepath, filetype = "GTiff", overwrite=TRUE)
}

```








```{r}
#reproject to the same crs (NAD 83, CA Albers (2011))
# method = near in order to preserve categorical data


ca_fire_2011_albers <- terra::project(ca_fire_2011, ca_raster, method = "near")

ca_fire_2012_albers <- terra::project(ca_fire_2012, ca_raster, method = "near")

ca_fire_2013_albers <- terra::project(ca_fire_2013, ca_raster, method = "near")

ca_fire_2014_albers <- terra::project(ca_fire_2014, ca_raster, method = "near")

ca_fire_2015_albers <- terra::project(ca_fire_2015, ca_raster, method = "near")

ca_fire_2016_albers <- terra::project(ca_fire_2016, ca_raster, method = "near")

ca_fire_2017_albers <- terra::project(ca_fire_2017, ca_raster, method = "near")

ca_fire_2018_albers <- terra::project(ca_fire_2018, ca_raster, method = "near")

ca_fire_2019_albers <- terra::project(ca_fire_2019, ca_raster, method = "near")

ca_fire_2020_albers <- terra::project(ca_fire_2020, ca_raster, method = "near")

```

```{r}
#cropping to spatial extent of CA raster

fire_2011_mask <- mask(ca_fire_2011_albers, ca_raster)

fire_2012_mask <- mask(ca_fire_2012_albers, ca_raster)

fire_2013_mask <- mask(ca_fire_2013_albers, ca_raster)

fire_2014_mask <- mask(ca_fire_2014_albers, ca_raster)

fire_2015_mask <- mask(ca_fire_2015_albers, ca_raster)

fire_2016_mask <- mask(ca_fire_2016_albers, ca_raster)

fire_2017_mask <- mask(ca_fire_2017_albers, ca_raster)

fire_2018_mask <- mask(ca_fire_2018_albers, ca_raster)

fire_2019_mask <- mask(ca_fire_2019_albers, ca_raster)

fire_2020_mask <- mask(ca_fire_2020_albers, ca_raster)
```

### Save reprojected and clipped rasters to the data drive
This saves time rather than rerunning the script

```{r}
saverast <- function(x){
  
  name <- paste0(deparse(substitute(x)), ".tif") #the name of the file will be the same as the input object, plus '.tif' at the end
  
  #write the raster into the correct folder in drive (the path will need to be updated for other files) and save it as a geotiff
  
  writeRaster(x, filename = here(gdrive_data, "/IntermediateData/IntermediateData_Files/fire_intensity", name), filetype = "GTiff", overwrite = TRUE)
  
}

saverast(fire_2011_mask)

saverast(fire_2012_mask)

saverast(fire_2013_mask)

saverast(fire_2014_mask)

saverast(fire_2015_mask)

saverast(fire_2016_mask)

saverast(fire_2017_mask)

saverast(fire_2018_mask)

saverast(fire_2019_mask)

saverast(fire_2020_mask)

```

### Read in saved, formatted layers

```{r}
#read in fire layers 

fire_2005_mask <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/med_high_fire_hist/fire_2005_masked.tif"))

fire_2006_mask <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/med_high_fire_hist/fire_2006_masked.tif"))


fire_2011_mask <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/fire_intensity/fire_2011_mask.tif"))

fire_2012_mask <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/fire_intensity/fire_2012_mask.tif"))

fire_2013_mask <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/fire_intensity/fire_2013_mask.tif"))

fire_2014_mask <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/fire_intensity/fire_2014_mask.tif"))

fire_2015_mask <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/fire_intensity/fire_2015_mask.tif"))

fire_2016_mask <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/fire_intensity/fire_2016_mask.tif"))

fire_2017_mask <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/fire_intensity/fire_2017_mask.tif"))

fire_2018_mask <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/fire_intensity/fire_2018_mask.tif"))

fire_2019_mask <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/fire_intensity/fire_2019_mask.tif"))

fire_2020_mask <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/fire_intensity/fire_2020_mask.tif"))

```


### Reclassify to keep only medium and high fire intensity
We are only interesting in how medium and high intensity fires influences human-black bear conflict, so we will reclassify to only keep medium and hihgh intensity fires. 

```{r}
#list out the from, to values in an object
m <- c(1, 0, #Unburned/Underburned/Low Burn Severity to no data
       2, 0, #Low burn severity to no data
       3, 4, #moderate burn severity to moderate/high burn severity
       4, 4, #high burn severity to moderate/high burn severity
       5, 0, #Increased Greenness/Increased Vegetation Response to no data
       6,0) #Non-Processing Area Mask to no data
       
#turn those values into a matrix
mat_fire <- matrix(m, ncol = 2, byrow = TRUE) 

#set categories (labels) for each numeric index

#index vector
index_fire <- c(0, 4)

#categories vector
cats_fire <- c("undefined","mod_high_burn")

#combine those into a dataframe
catdat_fire <- data.frame(id = index_fire, category = cats_fire)

#Reclassify

#create reclassification function

reclassify_fire <- function(x){
  fire_reclassed <- classify(x, mat_fire) #reclassify using the matrix above
  
  levels(fire_reclassed) <- catdat_fire #then reassign levels (labels) using the dataframe above
  
  return(fire_reclassed)
}

#Apply function to each layer
fire_2005_reclassed <- reclassify_fire(fire_2005_mask)

fire_2011_reclassed <- reclassify_fire(fire_2011_mask)

fire_2012_reclassed <- reclassify_fire(fire_2012_mask)

fire_2013_reclassed <- reclassify_fire(fire_2013_mask)

fire_2014_reclassed <- reclassify_fire(fire_2014_mask)

fire_2015_reclassed <- reclassify_fire(fire_2015_mask)

fire_2016_reclassed <- reclassify_fire(fire_2016_mask)

fire_2017_reclassed <- reclassify_fire(fire_2017_mask)

fire_2018_reclassed <- reclassify_fire(fire_2018_mask)

fire_2019_reclassed <- reclassify_fire(fire_2019_mask)

fire_2020_reclassed <- reclassify_fire(fire_2020_mask)

```


### Save reclassified rasters to the data drive

```{r}
saverast <- function(x){
  
  name <- paste0(deparse(substitute(x)), ".tif") #the name of the file will be the same as the input object, plus '.tif' at the end
  
  #write the raster into the correct folder in drive (the path will need to be updated for other files) and save it as a geotiff
  
  writeRaster(x, filename = here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/med_high_fire_rasters", name), filetype = "GTiff", overwrite = TRUE)
  
}

saverast(fire_2005_reclassed)
saverast(fire_2011_reclassed)
saverast(fire_2012_reclassed)
saverast(fire_2013_reclassed)
saverast(fire_2014_reclassed)
saverast(fire_2015_reclassed)
saverast(fire_2016_reclassed)
saverast(fire_2017_reclassed)
saverast(fire_2018_reclassed)
saverast(fire_2019_reclassed)
saverast(fire_2020_reclassed)

```


### Create multi-year merged layers for model data processing
This step is necessary for creation of the following layers: 
dist_fire5_6
dist_fire6_7
dist_fire7_8
dist_fire8_9
dist_fire9_10
- Distance to moderate to severe burn scars from 2010 to 2011
- Distance to moderate to severe burn scars from 2011 to 2012
- Distance to moderate to severe burn scars from 2012 to 2013
- Distance to moderate to severe burn scars from 2013 to 2014
- Distance to moderate to severe burn scars from 2014 to 2015
- Distance to moderate to severe burn scars from 2015 to 2016
- Distance to moderate to severe burn scars from 2016 to 2017
- Distance to moderate to severe burn scars from 2017 to 2018
- Distance to moderate to severe burn scars from 2018 to 2019

```{r}
#function for saving multi-year rasters
savemultirast <- function(x){
  
  name <- paste0(deparse(substitute(x)), ".tif") #the name of the file will be the same as the input object, plus '.tif' at the end
  
  #write the raster into the correct folder in drive (the path will need to be updated for other files) and save it as a geotiff
  
  writeRaster(x, filename = here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/med_high_fire_rasters/multiyear_rasters", name), filetype = "GTiff", overwrite = TRUE)
  
}
```

##### 2016

```{r}
#10-11 year: combined 2005-2006
fire_2005 <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/med_high_fire_hist/fire_2005_reclassed.tif"))

fire_2006 <_ rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/med_high_fire_hist/fire_2006_reclassed.tif"))

#8-9 year: combined 2007-2008

fire_2007 <_ rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/med_high_fire_hist/fire_2007_reclassed.tif"))

fire_2008 <_ rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/med_high_fire_hist/fire_2008_reclassed.tif"))

#6 to 7 year: combined 2009_2010
fire_2009 <_ rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/med_high_fire_hist/fire_2009_reclassed.tif"))

fire_2010 <_ rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/med_high_fire_hist/fire_2010_reclassed.tif"))

#4-5 year: combined 2011-2012

fire_2011 <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/med_high_fire_rasters/fire_2011_reclassed.tif"))

fire_2012 <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/med_high_fire_rasters/fire_2012_reclassed.tif"))

#merge those layers

fire_2011_2012 <- merge(fire_2011, fire_2012)


#2-3 year: combined 2013-2014

fire_2013 <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/med_high_fire_rasters/fire_2013_reclassed.tif"))

fire_2014 <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/med_high_fire_rasters/fire_2014_reclassed.tif"))

#merge those layers

fire_2013_2014 <- merge(fire_2013, fire_2014)

#save the layers

savemultirast(fire_2011_2012)

savemultirast(fire_2013_2014)
```

#### 2017

```{r}
#10-11 year: combined 2006-2007

#8-9 year: combined 2008-2009

#6 to 7 year: combined 2010_2011

#4-5 year: 2012-2013

fire_2012 <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/med_high_fire_rasters/fire_2012_reclassed.tif"))

fire_2013 <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/med_high_fire_rasters/fire_2013_reclassed.tif"))

#merge those layers

fire_2012_2013 <- merge(fire_2012, fire_2013)

#2-3 year: 2014-2015

fire_2014 <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/med_high_fire_rasters/fire_2014_reclassed.tif"))

fire_2015 <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/med_high_fire_rasters/fire_2015_reclassed.tif"))

#merge those layers

fire_2014_2015 <- merge(fire_2014, fire_2015)

#save the layers

savemultirast(fire_2012_2013)

savemultirast(fire_2014_2015)

```

##### 2018

```{r}
#10-11 year: combined 2007-2008

#8-9 year: combined 2009-2010

#6 to 7 year: combined 2011_2012

#4-5 year: 2013-2014 - this already exists from prior data wrangling

#2-3 year: 2015-2016

fire_2015 <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/med_high_fire_rasters/fire_2015_reclassed.tif"))

fire_2016 <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/med_high_fire_rasters/fire_2016_reclassed.tif"))

#merge those layers

fire_2015_2016 <- merge(fire_2015, fire_2016)

#save the layer

savemultirast(fire_2015_2016)
```

##### 2019

```{r}
#10-11 year: combined 2008-2009

#8-9 year: combined 2010-2011

#6 to 7 year: combined 2012_2013

#4-5 year: 2014-2015 - this already exists from prior data wrangling

#2-3 year: 2016-2017

fire_2016 <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/med_high_fire_rasters/fire_2016_reclassed.tif"))

fire_2017 <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/med_high_fire_rasters/fire_2017_reclassed.tif"))

#merge those layers

fire_2016_2017 <- merge(fire_2016, fire_2017)

#save the layer

savemultirast(fire_2016_2017)

```


##### 2020

```{r}
#10-11 year: combined 2009-2010

#8-9 year: combined 2011-2012

#6 to 7 year: combined 2013_2014

#4-5 year: 2015-2016 - this already exists from prior data wrangling

#2-3 year: 2017-2018

fire_2017 <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/med_high_fire_rasters/fire_2017_reclassed.tif"))

fire_2018 <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/med_high_fire_rasters/fire_2018_reclassed.tif"))

#merge the layers

fire_2017_2018 <- merge(fire_2017, fire_2018)

#save the layer

savemultirast(fire_2017_2018)
```

##### 2021

```{r}
#10-11 year: combined 2010-2011

#8-9 year: combined 2012-2013

#6 to 7 year: combined 2014_2015

#4-5 year: 2016-2017 - this already exists from prior data wrangling

#2-3 years: 2018-2019

fire_2018 <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/med_high_fire_rasters/fire_2018_reclassed.tif"))

fire_2019 <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/med_high_fire_rasters/fire_2019_reclassed.tif"))

#merge the layers

fire_2018_2019 <- merge(fire_2018, fire_2019)

#save the layer

savemultirast(fire_2018_2019)

```

### Distance to Fire Arc Layers
Due to the size of our rasters, using R to create distance layers was not feasible, and instead our group used ArcGIS to create these distance layers. We then reformatted these layers and saved them to our folder "model_ready" indicating they were ready for use in our final model code


```{r}
#Now we need to read back in the distance raster created in ArcGIS pro and mask these layers to match our CA extent again. Layers are already in the correct projection and resolution
#reading in the single years
dist_fire15 <-rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/dist_to_med_high_fire/dist_fire_2015.tif"))

dist_fire16 <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/dist_to_med_high_fire/dist_fire_2016.tif"))

dist_fire17 <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/dist_to_med_high_fire/dist_fire_2017.tif"))

dist_fire18 <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/dist_to_med_high_fire/dist_fire_2018.tif"))

dist_fire19 <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/dist_to_med_high_fire/dist_fire_2019.tif"))

dist_fire20 <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/dist_to_med_high_fire/dist_fire_2020.tif"))

#reading in the multiyears: 

#2011-2012 merged mod/high distance layer
dist_fire11_12 <-rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/dist_to_med_high_fire/multiyear_rasters/dist_fire_2011_2012.tif"))

#2012-2013 merged  mod/high distance layer
dist_fire12_13 <-rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/dist_to_med_high_fire/multiyear_rasters/dist_fire_2012_2013.tif"))

#2013-2014 merged mod/high distance layer
dist_fire13_14 <-rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/dist_to_med_high_fire/multiyear_rasters/dist_fire_2013_2014.tif"))

#2014-2015 merged  mod/high distance layer
dist_fire14_15 <-rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/dist_to_med_high_fire/multiyear_rasters/dist_fire_2014_2015.tif"))

#2015-2016 merged mod/high distance layer
dist_fire15_16 <-rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/dist_to_med_high_fire/multiyear_rasters/dist_fire_2015_2016.tif"))

#2016-2017 merged mod/high distance layer
dist_fire16_17 <-rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/dist_to_med_high_fire/multiyear_rasters/dist_fire_2016_2017.tif"))

#2017-2018 merged mod/high distance layer
dist_fire17_18 <-rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/dist_to_med_high_fire/multiyear_rasters/dist_fire_2017_2018.tif"))

#2018-2019 merged mod/high distance layer

dist_fire18_19 <-rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/dist_to_med_high_fire/multiyear_rasters/dist_fire_2018_2019.tif"))


```

#Remask the distance rasters

```{r}
# Now we need to mask these layers to just the CA extent 

##all the single years:
#added step, was having issue with distance raster from arc not matching ca_raster extent. This work around fixes the error 
dist_fire15 = resample(dist_fire15, ca_raster)
#only ran this for 2015, because 2015 was only one throwing error code 

fire15_ca_distance <- mask(dist_fire15, ca_raster) 

fire16_ca_distance <- mask(dist_fire16, ca_raster) 

fire17_ca_distance <- mask(dist_fire17, ca_raster) 

fire18_ca_distance <- mask(dist_fire18, ca_raster) 

fire19_ca_distance <- mask(dist_fire19, ca_raster) 

fire20_ca_distance <- mask(dist_fire20, ca_raster) 

# all the merged years 

#2011-2012 merged mod/high distance layer

fire11_12_ca_distance <- mask(dist_fire11_12, ca_raster) 

#2012-2013 merged  mod/high distance layer

fire12_13_ca_distance <- mask(dist_fire12_13, ca_raster) 

#2013-2014 merged mod/high distance layer

fire13_14_ca_distance <- mask(dist_fire13_14, ca_raster) 

#2014-2015 merged  mod/high distance layer

fire14_15_ca_distance <- mask(dist_fire14_15, ca_raster) 

#2015-2016 merged mod/high distance layer
#had to add trouble shoot work around here: 
dist_fire15_16 = resample(dist_fire15_16, ca_raster)
fire15_16_ca_distance <- mask(dist_fire15_16, ca_raster) 

#2016-2017 merged mod/high distance layer
dist_fire16_17 = resample(dist_fire16_17, ca_raster)

fire16_17_ca_distance <- mask(dist_fire16_17, ca_raster) 

#2017-2018 merged mod/high distance layer
dist_fire17_18 = resample(dist_fire17_18, ca_raster)

fire17_18_ca_distance <- mask(dist_fire17_18, ca_raster) 

#2018-2019 merged mod/high distance layer
dist_fire18_19 = resample(dist_fire18_19, ca_raster)

fire18_19_ca_distance <- mask(dist_fire18_19, ca_raster) 
```

### Save clipped fire DISTANCE rasters to the data drive for use in model and easy access

```{r}
saverast <- function(x){
  
  name <- paste0(deparse(substitute(x)), ".tif") #the name of the file will be the same as the input object, plus '.tif' at the end
  
  #write the raster into the correct folder in drive (the path will need to be updated for other files) and save it as a geotiff
  
  writeRaster(x, filename = here(gdrive_data, "/IntermediateData/IntermediateData_Files/model_ready/dist_fire", name), filetype = "GTiff", overwrite = TRUE)
  
}

saverast(fire15_ca_distance)

saverast(fire16_ca_distance)

saverast(fire17_ca_distance)

saverast(fire18_ca_distance)

saverast(fire19_ca_distance)

saverast(fire20_ca_distance)

saverast(fire11_12_ca_distance)

saverast(fire12_13_ca_distance)

saverast(fire13_14_ca_distance)

saverast(fire14_15_ca_distance)

saverast(fire15_16_ca_distance)

saverast(fire16_17_ca_distance)

saverast(fire17_18_ca_distance)

saverast(fire18_19_ca_distance)


```

