---
title: "Human Population Density Data Wrangling"
author: "Mia Guarnieri"
date: '2022-10-19'
output: html_document
---

##Setup and packages

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#attach packages
#note - Loading all of the ones Jason used, I don't think we will need all of them 

library(tidyverse)
library(here)
library(terra)
library(sf)
library(tmap)

#mia and claire's file path
#gdrive_data <- "/Volumes/GoogleDrive/Shared drives/Black_Bear_Aware/gdrive_data"

#sam's file path
gdrive_data <- "G:/Shared drives/Black_Bear_Aware/gdrive_data"

#chase's file path

```

#Data Wrangling Prior to Coding

*Data Sources*: Please see ReadMe file for metadata and citations.
*Reference Paper*: Insert hagani citation here

#Load in existing data layers

##California raster

Insert info about CA raster layer here

```{r}

ca_raster <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/CARaster_Tiger_Reclassed/CARaster_Tiger_Reclassed.img"))

#check the coordinate reference system 
crs(ca_raster)

#plot CA raster to confirm
tm_shape(ca_raster) +
  tm_raster() +
  tmap_mode("view")

```


##Human population density

Insert info about human pop density layers here

```{r}
#data are by year, from 2016 - 2020 (1 file per year)


human_pop_den_2016 <- rast(here(gdrive_data,"InputData/InputData_Files/HumanPopulationDensity_LandScan/2016_HumanPopulationDensity_LandScan/2016_conus_night_colorized.tif"))

human_pop_den_2017 <- rast(here(gdrive_data,"InputData/InputData_Files/HumanPopulationDensity_LandScan/2017_HumanPopulationDensity_LandScan/2017_conus_night_colorized.tif"))

human_pop_den_2018 <- rast(here(gdrive_data,"InputData/InputData_Files/HumanPopulationDensity_LandScan/2018_HumanPopulationDensity_LandScan/2018_conus_night_colorized.tif"))

human_pop_den_2019 <- rast(here(gdrive_data, "InputData/InputData_Files/HumanPopulationDensity_LandScan/2019_HumanPopulationDensity_LandScan/2019_conus_night_colorized.tif"))

human_pop_den_2020 <- rast(here(gdrive_data, "InputData/InputData_Files/HumanPopulationDensity_LandScan/2020_HumanPopulationDensity_LandScan/2020_conus_night_colorized.tif"))

human_pop_den_2021 <- rast(here(gdrive_data,"InputData/InputData_Files/HumanPopulationDensity_LandScan/2021_HumanPopulationDensity_LandScan/2021_conus_night_colorized.tif"))

```



#Adjust the data to the same geographic extent and projection


```{r}


```

