---
title: "Main Model Code"
author: "Mia Guarnieri"
date: '2022-09-05'
output: html_document
---

Test

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#attach packages
#note - Loading all of the ones Jason used, I don't think we will need all of them 

library(rgdal) #for loading vector data 
library(ggplot2)
library(terra) #or loading raster data
library(dismo)
library(rJava)
library(spThin)
library(bbmle)
library(Rlab)
library(ResourceSelection)
library(mctest)
library(dplyr)
library(randomForest)
library(corrplot)
library(ENMeval)
library(biomod2)
library(jtools)
library(car)
library(mgcv)
library(lubridate)
library(ROCR)
library(Metrics)
library(MuMIn)
library(boot)
library(splines)
library(tidyverse)


```

#Data Wrangling Prior to Coding

*Data Sources*: Please see ReadMe file for metadata and citations.
*Reference Paper*: Insert hagani citation here

#Load in the data

```{r}
ca_raster <- raster("G:/Data/IntermediateData/IntermediateData_Files/CARaster_Tiger_Reclassed/CARaster_Tiger_Reclassed.img")

plot(ca_raster)

ca_raster_df <- as.data.frame(ca_raster)

#check the coordinate reference system 
crs(ca_raster)
#NAD

```

#Reformatting Data

Step 1: Adjust the data to the same geographic extent and projection
```{r}
 human_pop_2020 <- raster("G:/Data/InputData/InputData_Files/HumanPopulationDensity_LandScan/2020_HumanPopulationDensity_LandScan/2020_HumanPopulationDensity_LandScan_colorized.tif")

plot(human_pop_2020)


```
```{r}
land_cover_2016 <- raster("G:/Data/InputData/InputData_Files/LandCover_NLCD/2016_LandCover_NLCD/nlcd_2016_land_cover_l48_20210604.img")
land_cover_2016
```


Step 2: Buffering conflict points to create conflict raster
```{r}

```

Step 3: Create distance layers from roads, streams, etc...
```{r}

```


#Model Creation

Step 1: replicating the model from Hagani et al.
```{r}

```

Step 2
```{r}

```

