---
title: "model_ready_tests"
author: "Claire Meuter and Mia Guarnieri"
date: "2022-10-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(terra)
library(sf)
library(tmap)
```

## In this rmarkdown we will double check that model inputs are in correct and ready to be fed into model stack 

#CA raster layer (not in model but our ideal projection, extent and resolution)
```{r}
#read in layer 
ca_raster <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/CARaster_Tiger_Formatted/formattedcaraster.tif"))
#check the coordinate reference system, resolution, extent 
crs(ca_raster)
res(ca_raster)
ext(ca_raster)

# SpatExtent : -380101.0666, 540047.238885548, -605327.054, 450346.970898149 (xmin, xmax, ymin, ymax)
```


# Elevation
```{r}
#read in layer 
elevation_test <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/Elevation_Formatted/dem_formatted.tif"))
#check the coordinate reference system, resolution, extent 
crs(elevation_test)
res(elevation_test)
ext(elevation_test)

# SpatExtent : -380101.0666, 540047.238885548, -605327.054, 450346.970898149 (xmin, xmax, ymin, ymax)

#elevation is model ready 
#plot elevation raster to confirm
tm_shape(elevation) +
  tm_raster() +
  tmap_mode("view")

```

# Aspect
```{r}
#read in aspect layer from google drive 
#read in layer 
aspect_test <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/aspect/aspect.tif"))
#check the coordinate reference system, resolution, extent 
crs(aspect_test)
res(aspect_test)
ext(aspect_test)

# SpatExtent : -380101.0666, 540047.238885548, -605327.054, 450346.970898149 (xmin, xmax, ymin, ymax)
#Aspect is model ready
```

#Terrian ruggedness
```{r}
#read in layer 
terrian_test <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/terrain_ruggedness/terrain_ruggedness.tif"))
#check the coordinate reference system, resolution, extent 
crs(terrian_test)
res(terrian_test)
ext(terrian_test)

# SpatExtent : -380101.0666, 540047.238885548, -605327.054, 450346.970898149 (xmin, xmax, ymin, ymax)
#terrian is model ready 
```

# Human population density 
```{r}
#read in layer 
pop_den_2016_test <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/Pop_Den_Rasters_Formatted/pop_den_2016_mask.tif"))

pop_den_2017_test <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/Pop_Den_Rasters_Formatted/pop_den_2017_mask.tif"))

pop_den_2018_test <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/Pop_Den_Rasters_Formatted/pop_den_2018_mask.tif"))

pop_den_2019_test <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/Pop_Den_Rasters_Formatted/pop_den_2019_mask.tif"))

pop_den_2020_test <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/Pop_Den_Rasters_Formatted/pop_den_2020_mask.tif"))

pop_den_2021_test <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/Pop_Den_Rasters_Formatted/pop_den_2021_mask.tif"))

#check the coordinate reference system, resolution, extent 
crs(pop_den_2018_test)
res(pop_den_2018_test)
ext(pop_den_2018_test)

# SpatExtent : -380101.0666, 540047.238885548, -605327.054, 450346.970898149 (xmin, xmax, ymin, ymax)
#terrian is model ready 
```

#Land Cover
```{r}

#read in layer 
lc_13_test <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/NLCD_Rasters_Formatted/lc13_reclassed.tif"))
#check the coordinate reference system, resolution, extent 
crs(lc_13_test)
res(lc_13_test)
ext(lc_13_test)

lc_16_test <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/NLCD_Rasters_Formatted/lc16_reclassed.tif"))
#check the coordinate reference system, resolution, extent 
crs(lc_16_test)
res(lc_16_test)
ext(lc_16_test)

lc_19_test <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/NLCD_Rasters_Formatted/lc19_reclassed.tif"))
#check the coordinate reference system, resolution, extent 
crs(lc_19_test)
res(lc_19_test)
ext(lc_19_test)

# SpatExtent : -380101.0666, 540047.238885548, -605327.054, 450346.970898149 (xmin, xmax, ymin, ymax)
#land cover is model ready 

```

#Road Density
```{r}
#read in layer 
road_density_test <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/road_density_formatted/road_density_ca.tif"))
#check the coordinate reference system, resolution, extent 
crs(road_density_test)
res(road_density_test)
ext(road_density_test)

# SpatExtent : -380101.0666, 540047.238885548, -605327.054, 450346.970898149 (xmin, xmax, ymin, ymax)
#road density is model ready 
```

#Distance to roads
```{r}
#read in layer 
road_dist_test <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/dist_to_road/model_ready/roads15_ca_distance.tif"))
#check the coordinate reference system, resolution, extent 
crs(road_dist_test)
res(road_dist_test)
ext(road_dist_test)

# SpatExtent : -380101.0666, 540047.238885548, -605327.054, 450346.970898149 (xmin, xmax, ymin, ymax)
#distance to roads is model ready 
```

#Distance to streams 
```{r}
#read in layer 
stream_dist_test <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/dist_to_stream/model_ready/stream17_ca_distance.tif"))
#check the coordinate reference system, resolution, extent 
crs(stream_dist_test)
res(stream_dist_test)
ext(stream_dist_test)

# SpatExtent : -380101.0666, 540047.238885548, -605327.054, 450346.970898149 (xmin, xmax, ymin, ymax)
#distance to streams is model ready 
```

#Distance to urban areas 
```{r}
#read in layer 
#urban13
urban13_dist_test <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/dist_to_urban/model_ready/ua13_ca_distance.tif"))
#check the coordinate reference system, resolution, extent 
crs(urban13_dist_test)
res(urban13_dist_test)
ext(urban13_dist_test)

#urban16
urban16_dist_test <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/dist_to_urban/model_ready/ua16_ca_distance.tif"))
#check the coordinate reference system, resolution, extent 
crs(urban16_dist_test)
res(urban16_dist_test)
ext(urban16_dist_test)

#urban19
urban19_dist_test <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/dist_to_urban/model_ready/ua19_ca_distance.tif"))
#check the coordinate reference system, resolution, extent 
crs(urban19_dist_test)
res(urban19_dist_test)
ext(urban19_dist_test)

# SpatExtent : -380101.0666, 540047.238885548, -605327.054, 450346.970898149 (xmin, xmax, ymin, ymax)
#distance to urban is model ready 
```

#Distance to reactional areas 
```{r}
#read in layer 
rec_dist_test <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/dist_to_stream/model_ready/stream17_ca_distance.tif"))
#check the coordinate reference system, resolution, extent 
crs(stream_dist_test)
res(stream_dist_test)
ext(stream_dist_test)

# SpatExtent : -380101.0666, 540047.238885548, -605327.054, 450346.970898149 (xmin, xmax, ymin, ymax)
#distance to roads is model ready 
```

#Distance to forest cover 
```{r}

```

#Forest density (canopy cover) 
```{r}

```


