---
title: "model_ready_tests"
author: "Claire Meuter and Mia Guarnieri"
date: "2022-10-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(terra)
library(sf)
library(tmap)
```

## In this rmarkdown we will double check that model inputs are in correct and ready to be fed into model stack 

#CA raster layer (not in model but our ideal projection, extent and resolution)
```{r}
#read in layer 
ca_raster <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/CARaster_Tiger_Formatted/formattedcaraster.tif"))
#check the coordinate reference system, resolution, extent 
crs(ca_raster)
res(ca_raster)
ext(ca_raster)

# SpatExtent : -380101.0666, 540047.238885548, -605327.054, 450346.970898149 (xmin, xmax, ymin, ymax)
```


# Elevation
```{r}
#read in layer 
elevation <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/Elevation_Formatted/dem_formatted.tif"))
#check the coordinate reference system, resolution, extent 
crs(elevation)
res(elevation)
ext(elevation)

# SpatExtent : -380101.0666, 540047.238885548, -605327.054, 450346.970898149 (xmin, xmax, ymin, ymax)

#elevation is model ready 
#plot elevation raster to confirm
tm_shape(elevation) +
  tm_raster() +
  tmap_mode("view")

```

# Aspect
```{r}
#read in aspect layer from google drive 
#read in layer 
aspect <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/aspect/aspect.tif"))
#check the coordinate reference system, resolution, extent 
crs(aspect)
res(aspect)
ext(aspect)

# SpatExtent : -380101.0666, 540047.238885548, -605327.054, 450346.970898149 (xmin, xmax, ymin, ymax)
#Aspect is model ready
```

#Terrian ruggedness
```{r}
#read in layer 
terrian <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/terrain_ruggedness/terrain_ruggedness.tif"))
#check the coordinate reference system, resolution, extent 
crs(terrian)
res(terrian)
ext(terrian)

# SpatExtent : -380101.0666, 540047.238885548, -605327.054, 450346.970898149 (xmin, xmax, ymin, ymax)
#terrian is model ready 
```

# Human population density 
```{r}
#read in layer 
pop_den_2016 <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/Pop_Den_Rasters_Formatted/pop_den_2016_mask.tif"))

pop_den_2017 <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/Pop_Den_Rasters_Formatted/pop_den_2017_mask.tif"))

pop_den_2018 <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/Pop_Den_Rasters_Formatted/pop_den_2018_mask.tif"))

pop_den_2019 <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/Pop_Den_Rasters_Formatted/pop_den_2019_mask.tif"))

pop_den_2020 <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/Pop_Den_Rasters_Formatted/pop_den_2020_mask.tif"))

pop_den_2021 <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/Pop_Den_Rasters_Formatted/pop_den_2021_mask.tif"))

#check the coordinate reference system, resolution, extent 
crs(pop_den_2018)
res(pop_den_2018)
ext(pop_den_2018)

# SpatExtent : -380101.0666, 540047.238885548, -605327.054, 450346.970898149 (xmin, xmax, ymin, ymax)
#terrian is model ready 
```

#Land Cover
```{r}

#read in layer 
lc_13 <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/NLCD_Rasters_Formatted/lc13_reclassed.tif"))
#check the coordinate reference system, resolution, extent 
crs(lc_13)
res(lc_13)
ext(lc_13)

lc_16 <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/NLCD_Rasters_Formatted/lc16_reclassed.tif"))
#check the coordinate reference system, resolution, extent 
crs(lc_16)
res(lc_16)
ext(lc_16)

lc_19 <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/NLCD_Rasters_Formatted/lc19_reclassed.tif"))
#check the coordinate reference system, resolution, extent 
crs(lc_19)
res(lc_19)
ext(lc_19)

# SpatExtent : -380101.0666, 540047.238885548, -605327.054, 450346.970898149 (xmin, xmax, ymin, ymax)
#land cover is model ready 

```

#Road Density
```{r}

```

#Distance to roads
```{r}

```

#Distance to streams 
```{r}

```

#Distance to urban areas 
```{r}

```

#Distance to reactional areas 
```{r}

```

#Distance to forest cover 
```{r}

```

#Forest density (canopy cover) 
```{r}

```


