---
title: "Model Creation"
author: "Mia Guarnieri"
date: '2022-10-19'
output: html_document
---

##Setup and packages

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#attach packages
#note - Loading all of the ones Jason used, I don't think we will need all of them 

library(tidyverse)
library(here)
library(terra)
library(sf)
library(tmap)
library(lubridate)

#mia, chase, and claire's file path
#gdrive_data <- "/Volumes/GoogleDrive/Shared drives/Black_Bear_Aware/gdrive_data"

#sam's file path
gdrive_data <- "G:/Shared drives/Black_Bear_Aware/gdrive_data"

```

#Data Wrangling Prior to Coding

*Data Sources*: Please see ReadMe file for metadata and citations.
*Reference Paper*: Insert hagani citation here

#Load in existing data layers

##California raster

Insert info about CA raster layer here

```{r}

ca_raster <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/CARaster_Tiger_Reclassed/CARaster_Tiger_Reclassed.img")) #I think this is the wrong CA raster?

#check the coordinate reference system 
crs(ca_raster)

#plot CA raster to confirm
tm_shape(ca_raster) +
  tm_raster() +
  tmap_mode("view")

```

##Conflict data

Insert info about WIR data here

```{r}
conflict_points <- vect(here(gdrive_data, "/IntermediateData/IntermediateData_Files/conflict_formatted/conflict_points/conflict_points.shp"))

conflict_buffer <- vect(here(gdrive_data, "/IntermediateData/IntermediateData_Files/conflict_formatted/conflict_buffered/conflict_buffered.shp"))

```

#Stack environmental rasters for easy extraction

```{r}

```


#Extract raster values by year for conflict points (test)

```{r}
#add a year column to confict point data

con_year <- conflict_points_albers %>% 
  separate(incident_date, into = c("inc_date", "inc_time"), sep = " ") %>% 
  separate(inc_date, into = c("month", "day", "year"), sep = "/") %>% 
  mutate(year = as.character.Date(year))

#filter for 2016
con_16 <- con_year %>% 
  filter(year == "2016") %>% 
  filter(wir_id != "WIR-2016-001081") #removing an erroneous point (was showing up in Israel)

con_16_sv <- vect(con_16)

con16_back <- st_as_sf(con_16_sv)

#read in data for that year

dem <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/Elevation_Formatted/dem_formatted.tif"))

lc <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/NLCD_Rasters_Formatted/lc16_reclassed.tif"))


#extract data points from these rasters and add them to the shapefile dataframe

dem_ex <- extract(dem, con_16_sv, method = "simple", bind = TRUE, ID = TRUE)

```

#Sam is playing around with the code here:
```{r}
#try to stack rasters

raster_stack_test <- c(roads15_ca_distance,stream17_ca_distance)

tm_shape(raster_stack_test) +
  tm_raster() +
  tmap_mode("view")

raster_test_df <- as.data.frame(raster_stack_test)
```

