---
title: "model_ready_tests"
author: "Claire Meuter and Mia Guarnieri"
date: "2022-10-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## In this rmarkdown we will double check that model inputs are in correct and ready to be fed into model stack 

#CA raster layer (not in model but our ideal projection, extent and resolution)
```{r}
#read in layer 
ca_raster <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/CARaster_Tiger_Formatted/formattedcaraster.tif"))
#check the coordinate reference system, resolution, extent 
crs(ca_raster)
res(ca_raster)
ext(ca_raster)

# SpatExtent : -380101.0666, 540047.238885548, -605327.054, 450346.970898149 (xmin, xmax, ymin, ymax)
```


# Elevation
```{r}
#just skeleton code- not up to date actually 
#read in layer 
elevation <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/CARaster_Tiger_Formatted/formattedcaraster.tif"))
#check the coordinate reference system, resolution, extent 
crs(elevation)
res(elevation)
ext(elevation)

# SpatExtent : -380101.0666, 540047.238885548, -605327.054, 450346.970898149 (xmin, xmax, ymin, ymax)
```

# Aspect
```{r}
#read in aspect layer from google drive 
#read in layer 
aspect <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/aspect/aspect.tif"))
#check the coordinate reference system, resolution, extent 
crs(aspect)
res(aspect)
ext(aspect)

# SpatExtent : -380101.0666, 540047.238885548, -605327.054, 450346.970898149 (xmin, xmax, ymin, ymax)
```

#Terrian ruggedness
```{r}
#read in layer 
terrian <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/terrain_ruggedness/terrain_ruggedness.tif"))
#check the coordinate reference system, resolution, extent 
crs(terrian)
res(terrian)
ext(terrian)

# SpatExtent : -380101.0666, 540047.238885548, -605327.054, 450346.970898149 (xmin, xmax, ymin, ymax)
```

# Human population density 
```{r}

```

#Land Cover
```{r}

```

#Road Density
```{r}

```

#Distance to roads
```{r}

```

#Distance to streams 
```{r}

```

#Distance to urban areas 
```{r}

```

#Distance to reactional areas 
```{r}

```

#Distance to forest cover 
```{r}

```

#Forest density (canopy cover) 
```{r}

```


