---
title: "Land Cover Data Wrangling"
author: "Mia Guarnieri"
date: '2022-10-19'
output: html_document
---

##Setup and packages

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#attach packages

library(tidyverse)
library(here)
library(terra)
library(sf)
library(tmap)
library(beepr)
#mia and claire's file path
gdrive_data <- "/Volumes/GoogleDrive/Shared drives/Black_Bear_Aware/gdrive_data"

#sam's file path
#gdrive_data <- "G:/Shared drives/Black_Bear_Aware/gdrive_data"

#chase's file path

```

#Data Wrangling Prior to Coding

*Data Sources*: Please see ReadMe file for metadata and citations.
*Reference Paper*: Insert hagani citation here

#Load in relevant data layers

##California raster

Insert info about CA raster layer here

```{r}

ca_raster <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/CARaster_Tiger_Formatted/formattedcaraster.tif"))

#check the coordinate reference system and resolution
crs(ca_raster)

res(ca_raster)


#plot CA raster to confirm
tm_shape(ca_raster) +
  tm_raster() +
  tmap_mode("view")

res(ca_raster)

```

##Land cover

Land cover data from the NLCD from 2001-2019 (~3 year intervals).

Land cover class codes and descriptions for contiguous US:

0
Unclassified
Producer defined

11
Open Water - All areas of open water, generally with less than 25% cover or vegetation or soil
NLCD Legend Land Cover Class Descriptions

12
Perennial Ice/Snow - All areas characterized by a perennial cover of ice and/or snow, generally greater than 25% of total cover.
NLCD Legend Land Cover Class Descriptions

21
Developed, Open Space - Includes areas with a mixture of some constructed materials, but mostly vegetation in the form of lawn grasses.  Impervious surfaces account for less than 20 percent of total cover.  These areas most commonly include large-lot single-family housing units, parks, golf courses, and vegetation planted in developed settings for recreation, erosion control, or aesthetic purposes.
NLCD Legend Land Cover Class Descriptions

22
Developed, Low Intensity -Includes areas with a mixture of constructed materials and vegetation.  Impervious surfaces account for 20-49 percent of total cover.  These areas most commonly include single-family housing units.
NLCD Legend Land Cover Class Descriptions

23
Developed, Medium Intensity - Includes areas with a mixture of constructed materials and vegetation. Impervious surfaces account for 50-79 percent of the total cover.  These areas most commonly include single-family housing units.
NLCD Legend Land Cover Class Descriptions

24
Developed, High Intensity - Includes highly developed areas where people reside or work in high numbers. Examples include apartment complexes, row houses and commercial/industrial.  Impervious surfaces account for 80 to 100 percent of the total cover.
NLCD Legend Land Cover Class Descriptions

31
Barren Land (Rock/Sand/Clay) - Barren areas of bedrock, desert pavement, scarps, talus, slides, volcanic material, glacial debris, sand dunes, strip mines, gravel pits and other accumulations of earthen material. Generally, vegetation accounts for less than 15% of total cover.
NLCD Legend Land Cover Class Descriptions

41
Deciduous Forest - Areas dominated by trees generally greater than 5 meters tall, and greater than 20% of total vegetation cover. More than 75 percent of the tree species shed foliage simultaneously in response to seasonal change.
NLCD Legend Land Cover Class Descriptions

42
Evergreen Forest - Areas dominated by trees generally greater than 5 meters tall, and greater than 20% of total vegetation cover. More than 75 percent of the tree species maintain their leaves all year. Canopy is never without green foliage.
NLCD Legend Land Cover Class Descriptions

43
Mixed Forest - Areas dominated by trees generally greater than 5 meters tall, and greater than 20% of total vegetation cover. Neither deciduous nor evergreen species are greater than 75 percent of total tree cover.
NLCD Legend Land Cover Class Descriptions

52
Shrub/Scrub - Areas dominated by shrubs; less than 5 meters tall with shrub canopy typically greater than 20% of total vegetation. This class includes true shrubs, young trees in an early successional stage or trees stunted from environmental conditions.
NLCD Legend Land Cover Class Descriptions

71
Grassland/Herbaceous - Areas dominated by grammanoid or herbaceous vegetation, generally greater than 80% of total vegetation.  These areas are not subject to intensive management such as tilling, but can be utilized for grazing.
NLCD Legend Land Cover Class Descriptions

81
Pasture/Hay - Areas of grasses, legumes, or grass-legume mixtures planted for livestock grazing or the production of seed or hay crops, typically on a perennial cycle. Pasture/hay vegetation accounts for greater than 20 percent of total vegetation.
NLCD Legend Land Cover Class Descriptions

82
Cultivated Crops - Areas used for the production of annual crops, such as corn, soybeans, vegetables, tobacco, and cotton, and also perennial woody crops such as orchards and vineyards. Crop vegetation accounts for greater than 20 percent of total vegetation. This class also includes all land being actively tilled.
NLCD Legend Land Cover Class Descriptions

90
Woody Wetlands - Areas where forest or shrub land vegetation accounts for greater than 20 percent of vegetative cover and the soil or substrate is periodically saturated with or covered with water.
NLCD Legend Land Cover Class Descriptions

95
Emergent Herbaceous Wetlands - Areas where perennial herbaceous vegetation accounts for greater than 80 percent of vegetative cover and the soil or substrate is periodically saturated with or covered with water.
NLCD Legend Land Cover Class Descriptions

```{r}
#data are by year, from 2001 - 2016 (1 file per year)


#we will save these just in case, but we do not need them as they are outside of our temporal extent

# land_cover_2001 <- rast(here(gdrive_data, "/InputData/InputData_Files/LandCover_NLCD/2001_LandCover_NLCD/2001_LandCover_NLCD.img"))
# 
# land_cover_2004 <- rast(here(gdrive_data, "/InputData/InputData_Files/LandCover_NLCD/2004_LandCover_NLCD/nlcd_2004_land_cover_l48_20210604.img"))
# 
# land_cover_2006 <- rast(here(gdrive_data, "/InputData/InputData_Files/LandCover_NLCD/2006_LandCover_NLCD/nlcd_2006_land_cover_l48_20210604.img"))
# 
# land_cover_2008 <- rast(here(gdrive_data, "/InputData/InputData_Files/LandCover_NLCD/2008_LandCover_NLCD/nlcd_2008_land_cover_l48_20210604.img"))
# 
# land_cover_2011 <- rast(here(gdrive_data, "/InputData/InputData_Files/LandCover_NLCD/2011_LandCover_NLCD/nlcd_2011_land_cover_l48_20210604.img"))

#run these


land_cover_2013 <- rast(here(gdrive_data, "/InputData/InputData_Files/LandCover_NLCD/2013_LandCover_NLCD/nlcd_2013_land_cover_l48_20210604.img"))

land_cover_2016 <- rast(here(gdrive_data, "/InputData/InputData_Files/LandCover_NLCD/2016_LandCover_NLCD/nlcd_2016_land_cover_l48_20210604.img"))

land_cover_2019 <- rast(here(gdrive_data, "/InputData/InputData_Files/LandCover_NLCD/2019_LandCover_NLCD/nlcd_2019_land_cover_l48_20210604.img"))

beep(3)
```

#Adjust the data to the same geographic extent and projection

We will additionally reclassify land cover rasters after they are clipped, to avoid exceeding memory capacity with large files

```{r}
#reproject to the same crs (NAD 83, CA Albers (2011))
<<<<<<< HEAD
# lc_2001_calbers <- project(land_cover_2001, ca_raster)

# lc_2004_calbers <- project(land_cover_2004, ca_raster)

# lc_2006_calbers <- project(land_cover_2006, ca_raster)

# lc_2008_calbers <- project(land_cover_2008, ca_raster)

# lc_2011_calbers <- project(land_cover_2011, ca_raster)
=======
>>>>>>> 72cb868fc8372eed2825ce4762e60bd6e0826167

lc_2013_calbers <- project(land_cover_2013, ca_raster)

lc_2016_calbers <- project(land_cover_2016, ca_raster)

lc_2019_calbers <- project(land_cover_2019, ca_raster)


<<<<<<< HEAD
#cropping the files to the spatial extent of CA
# lc_2001_mask <- mask(lc_2001_calbers, ca_raster)

# lc_2004_mask <- mask(lc_2004_calbers, ca_raster)

# lc_2006_mask <- mask(lc_2004_calbers, ca_raster)

# lc_2008_mask <- mask(lc_2004_calbers, ca_raster)

# lc_2011_mask <- mask(lc_2004_calbers, ca_raster)
=======
#old files - do not need them anymore as they are outside our temporal extent

# lc_2001_calbers <- project(land_cover_2001, ca_raster)
# 
# lc_2004_calbers <- project(land_cover_2004, ca_raster)
# 
# lc_2006_calbers <- project(land_cover_2006, ca_raster)
# 
# lc_2008_calbers <- project(land_cover_2008, ca_raster)
# 
# lc_2011_calbers <- project(land_cover_2011, ca_raster)


#cropping the files to the spatial extent of CA

lc_2013_mask <- mask(lc_2013_calbers, ca_raster)

lc_2016_mask <- mask(lc_2016_calbers, ca_raster)
>>>>>>> 72cb868fc8372eed2825ce4762e60bd6e0826167

lc_2019_mask <- mask(lc_2019_calbers, ca_raster)

#old files - do not need them anymore as they are outside our temporal extent

# lc_2001_mask <- mask(lc_2001_calbers, ca_raster)
# 
# lc_2004_mask <- mask(lc_2004_calbers, ca_raster)
# 
# lc_2006_mask <- mask(lc_2006_calbers, ca_raster)
# 
# lc_2008_mask <- mask(lc_2008_calbers, ca_raster)
# 
# lc_2011_mask <- mask(lc_2011_calbers, ca_raster)

#Create reclassification matrix for land cover values (from, to)

#list out the from, to values in an object
m <- c(11, 11, #open water 
       12, 12, #perennial ice/snow
       21, 21, #developed open space
       22, 22, #developed low intensity
       23, 23, #developed med intensity
       24, 24, #developed high intensity
       31, 31, #barren land (rock/sand/clay)
       41, 40, #deciduous forest --> forest
       42, 40, #evergreen forest --> forest
       43, 40, #mixed forest --> forest
       52, 52, #shrub/scrub
       71, 71, #grassland/herbaceous
       81, 80, # pasture/hay --> planted_cultivated
       82, 80, #cultivated crops --> planted_cultivated
       90, 90, #woody wetlands --> wetlands
       95, 90) #emergent herbaceous wetlands --> wetlands

#turn those values into a matrix
mat <- matrix(m, ncol = 2, byrow = TRUE) 

#set categories (labels) for each numeric index

#index vector
index <- c(0, 11, 12, 21, 22, 23, 24, 31, 40, 52, 71, 80, 90)

#categories vector
cats <- c("undefined","open_water", "perrenial_ice_snow", "dev_open_space","dev_low_int", "dev_med_int", "dev_high_int", "barren_land", "forest", "shrub_scrub", "grassland_herbaceous", "planted_cultivated", "wetlands")

#combine those into a dataframe
catdat <- data.frame(id = index, category = cats)

#Reclassify

#create reclassification function

reclassify <- function(x){
  x_reclassed <- classify(x, mat) #reclassify using the matrix above
  
  levels(x_reclassed) <- catdat #then reassign levels (labels) using the dataframe above
  
  return(x_reclassed)
}

#Apply function to each layer

<<<<<<< HEAD
#lc01_reclassed <- reclassify(lc_2001_mask)

#lc04_reclassed <- reclassify(lc_2004_mask)

#lc06_reclassed <- reclassify(lc_2006_mask)

#lc08_reclassed <- reclassify(lc_2008_mask)

#lc11_reclassed <- reclassify(lc_2011_mask)

=======
>>>>>>> 72cb868fc8372eed2825ce4762e60bd6e0826167
lc13_reclassed <- reclassify(lc_2013_mask)

lc16_reclassed <- reclassify(lc_2016_mask)

lc19_reclassed <- reclassify(lc_2019_mask)

<<<<<<< HEAD
beep(3)
=======
#don't need these; they are outside our temporal extent

# lc01_reclassed <- reclassify(lc_2001_mask)
# 
# lc04_reclassed <- reclassify(lc_2004_mask)
# 
# lc06_reclassed <- reclassify(lc_2006_mask)
# 
# lc08_reclassed <- reclassify(lc_2008_mask)
# 
# lc11_reclassed <- reclassify(lc_2011_mask)
>>>>>>> 72cb868fc8372eed2825ce4762e60bd6e0826167

```

#Save reclassified, reprojected, and clipped rasters to the data drive

**ONLY RUN THIS ONE TIME; IT WILL NOT RUN ONCE THE FILES ARE SAVED**

```{r}
saverast <- function(x){
  
  name <- paste0(deparse(substitute(x)), ".tif") #the name of the file will be the same as the input object, plus '.tif' at the end
  
  #write the raster into the correct folder in drive (the path will need to be updated for other files) and save it as a geotiff
  
  writeRaster(x, filename = here(gdrive_data, "/IntermediateData/IntermediateData_Files/NLCD_Rasters_Formatted", name), filetype = "GTiff", overwrite = TRUE)
  
}

saverast(lc13_reclassed)

saverast(lc16_reclassed)

saverast(lc19_reclassed)

#outside our temporal extent - don't run these

# saverast(lc01_reclassed)
# 
# saverast(lc04_reclassed)
# 
# saverast(lc06_reclassed)
# 
# saverast(lc08_reclassed)
# 
# saverast(lc11_reclassed)

```

