---
title: "Other Data Wrangling"
author: "Mia Guarnieri"
date: '2022-10-19'
output: html_document
---

##Setup and packages

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#attach packages
#note - Loading all of the ones Jason used, I don't think we will need all of them 

library(tidyverse)
library(here)
library(terra)
library(sf)
library(tmap)

#mia and claire's file path
gdrive_data <- "/Volumes/GoogleDrive/Shared drives/Black_Bear_Aware/gdrive_data"

#sam's file path
#gdrive_data <- "G:/Shared drives/Black_Bear_Aware/gdrive_data"

#chase's file path

```

#Data Wrangling Prior to Coding

*Data Sources*: Please see ReadMe file for metadata and citations.
*Reference Paper*: Insert hagani citation here

#Load in existing data layers

##California raster

Insert info about CA raster layer here

```{r}

ca_raster <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/CARaster_Tiger_Reclassed/CARaster_Tiger_Reclassed.img"))

#check the coordinate reference system 
crs(ca_raster)

#plot CA raster to confirm
tm_shape(ca_raster) +
  tm_raster() +
  tmap_mode("view")

```

##Conflict data

Insert info about WIR data here

```{r}

#Buffer conflict points to create conflict raster
# first loading in conflict data 
conflict_points <- read_csv(here(gdrive_data,"/IntermediateData/IntermediateData_Files/2022_07_18_WIRClean_CDFW.csv"))

# Turning conflict_points into a sf object from data frame 

#st_as_sf(conflict_points, coords = "latitude","longitude")
conflict_points_sf <- st_as_sf(conflict_points,coords = c("latitude","longitude"), crs = 6414)

# look into buffering with SF SF user st_buffer
#need to set conflict crs 
#need to set extent 
#Need to buffer 

#test plot
#how to turn data frame with lat and long columns into an sf object 




tm_shape(conflict_points_sf) +
  tm_dots()+
  tmap_mode("view")
class(conflict_points_sf)
```

##Elevation (DEM, mean elevation)

Two separate DEM rasters were available at 7.5-arc second resolution from USGS EROS. These will need to be merged togther.

```{r}

#northern California raster

dem_norcal <- rast(here(gdrive_data, "/InputData/InputData_Files/2010_DEM_USGSEROS_Norcal/30n150w_20101117_gmted_mea075.tif"))

#southern California raster

dem_socal <- rast(here(gdrive_data, "/InputData/InputData_Files/2010_DEM_USGSEROS_Socal/30n120w_20101117_gmted_mea075.tif"))

#merging the two rasters

dem_ca <- merge(dem_norcal, dem_socal)

#plot DEM raster
tm_shape(dem_ca) +
  tm_raster() +
  tmap_mode("view")

```

##Road density

```{r}
road_density_2014 <- rast(here(gdrive_data,"/InputData/InputData_Files/2014_UsaRoadDensity_ESRI/2014_UsaRoadDensity_ESRI.tif"))
```


#Adjust the data to the same geographic extent and projection

##Conflict

```{r}

```

##Elevation
```{r}
#reproject to the same crs (NAD 83, CA Albers (2011))
dem_calbers <- project(dem_ca, ca_raster)

#cropping to the spatial extent of CA
dem_mask <- mask(dem_calbers, ca_raster)

```

##Road density

```{r}

```
