---
title: "Other Data Wrangling"
author: "Mia Guarnieri"
date: '2022-10-19'
output: html_document
---

##Setup and packages

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#attach packages
#note - Loading all of the ones Jason used, I don't think we will need all of them 

library(tidyverse)
library(here)
library(terra)
library(sf)
library(tmap)
library(beepr)
library(rgeos)
library(rmapshaper)

#mia and claire's file path
gdrive_data <- "/Volumes/GoogleDrive/Shared drives/Black_Bear_Aware/gdrive_data"

#sam's file path
#gdrive_data <- "G:/Shared drives/Black_Bear_Aware/gdrive_data"

#chase's file path

```

#Data Wrangling Prior to Coding

*Data Sources*: Please see ReadMe file for metadata and citations.
*Reference Paper*: Insert hagani citation here

#Load in existing data layers

##California raster

Insert info about CA raster layer here

```{r}
ca_raster <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/CARaster_Tiger_Formatted/formattedcaraster.tif"))

#check the coordinate reference system 
crs(ca_raster)


#plot CA raster to confirm
tm_shape(ca_raster) +
  tm_raster() +
  tmap_mode("view")

```

##Conflict data

Insert info about WIR data here

```{r}
#Buffer conflict points to create conflict raster

# load in conflict data 
conflict_points <- read_csv(here(gdrive_data,"/IntermediateData/IntermediateData_Files/2022_07_18_WIRClean_CDFW.csv"))

# turn conflict_points into a sf object from data frame 

#st_as_sf(conflict_points, coords = "latitude","longitude")
conflict_points_sf <- st_as_sf(conflict_points,coords = c("longitude","latitude"), crs = 4326)

#4326 is code for WGS 84- map does not render properly otherwise, so we need to reproject  to 6414 (NAD83(2011) / California Albers) (but also I think this issue might go away when I remove the error points that are outside CA)

#reprojecting to NAD83(2011) / California Albers
conflict_points_albers <- st_transform(conflict_points_sf,crs = 6414)

#plot to check for erroneous points

tm_shape(conflict_points_albers) +
  tm_dots() +
  tmap_mode("view")

#wir_id of erroneous points: WIR-2016-001081, WIR-2017-002420

#filter erroneous points out of the WIR data

conflict <- conflict_points_albers %>% 
  filter(wir_id != "WIR-2016-001081",
         wir_id != "WIR-2017-002420")

#plot to double check
tm_shape(conflict) +
  tm_dots() +
  tmap_mode("view")

#saving the file to intermediate data in the gdrive

st_write(conflict, dsn = here(gdrive_data, "/IntermediateData/IntermediateData_Files/conflict_formatted/conflict_points", "conflict_points.shp"))

#test: reading from the gdrive to make sure it works and plotting it

con_points <- read_sf(here(gdrive_data, "/IntermediateData/IntermediateData_Files/conflict_formatted/conflict_points/conflict_points.shp"))

tm_shape(con_points) +
  tm_dots() +
  tmap_mode("view")

#buffering points using SF fuction st_buffer. Units are in meters
conflict_points_buffered <- st_buffer(con_points, dist = 5000)

#plot to confirm

 tm_shape(conflict_points_buffered) +
   tm_polygons() +
  tm_shape(con_points) +
   tm_dots(col  = "blue") +
   tmap_mode("view")
 
#saving the file to intermediate data in the gdrive

 st_write(conflict_points_buffered, dsn = here(gdrive_data, "/IntermediateData/IntermediateData_Files/conflict_formatted/conflict_buffered", "conflict_buffered.shp"))

#test: reading from the gdrive to make sure it works and plotting it

con_buffer <- read_sf(here(gdrive_data, "/IntermediateData/IntermediateData_Files/conflict_formatted/conflict_buffered/conflict_buffered.shp"))

tm_shape(con_buffer) +
  tm_polygons() +
tm_shape(con_points) +
   tm_dots(col  = "blue") +
tmap_mode("view")

#comparing arc vs r generated buffers

arc_buffers <- read_sf(here(gdrive_data, "/IntermediateData/IntermediateData_Files/conflict_formatted/conflict_buffer_test/conflict_points_buffer_arctest.shp"))

conflict_buffer <- read_sf(here(gdrive_data, "/IntermediateData/IntermediateData_Files/conflict_formatted/conflict_buffered/conflict_buffered.shp"))

tm_shape(conflict_buffer) +
  tm_polygons() +
tm_shape(arc_buffers) +
  tm_polygons(col = "blue") +
tmap_mode("view")

#refining buffer data; only keeping columns we need and sorting the dates to be readable by R

conflict_buffer_refined <- read_sf(here(gdrive_data, "/IntermediateData/IntermediateData_Files/conflict_formatted/conflict_buffered/conflict_buffered.shp")) %>% 
  separate(incdnt_d, into = c("inc_date", "inc_time"), sep = " ") %>% 
  separate(inc_date, into = c("month", "day", "year"), sep = "/") %>% 
  mutate(year = as.character.Date(year),
         "type" = cnfrmd_,
         "county" = cnty_nm,
         "conflict" = 1) %>% 
  select(conflict, wir_id:year, type, county, geometry)

#saving the file to intermediate data in the gdrive

 st_write(conflict_buffer_refined, dsn = here(gdrive_data, "/IntermediateData/IntermediateData_Files/conflict_formatted/conflict_buffered_refined", "conflict_buffered_refined.shp"))

```

```{r}
#creation of zero (non-conflict) points

#generate random points within CA and buffer

#make CA raster a shapefile

ca_raster <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/CARaster_Tiger_Formatted/formattedcaraster.tif"))

ca_pol <- as.polygons(ca_raster)

ca_sf <- st_as_sf(ca_pol)

sample_points <- st_sample(ca_sf, 4663) %>% 
  st_as_sf()

sample_buff <- st_buffer(sample_points, dist = 5000)

#buffer sample points
sample_points_buff <- st_buffer(sample_points_claire, dist = 5000)

#load in conflict buffer

con_buff <- read_sf(here(gdrive_data, "/IntermediateData/IntermediateData_Files/conflict_formatted/conflict_buffered_refined/conflict_buffered_refined.shp"))


#select for the ones that are NOT intersecting

 zeros <- ms_erase(sample_points,con_buff)

```

#claire's way of getting zeros
```{r}
#reading in CA raster 
ca_raster <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/CARaster_Tiger_Formatted/formattedcaraster.tif"))
#reading in conflict buffer shp file
con_buff <- read_sf(here(gdrive_data, "/IntermediateData/IntermediateData_Files/conflict_formatted/conflict_buffered_refined/conflict_buffered_refined.shp"))

#Making sure con_buf is in the same projection
con_buff <- st_transform(con_buff,crs(ca_raster))

#Doing an inverse mask- Anything withing the conflict buffer shapefiles will be a 0 value. So we get a california raster with all the conflict buffer locations zeroed out
con_buff_inverse <- mask(ca_raster, con_buff, inverse = TRUE)

# Mapping just to be sure it lookes right:
tm_shape(con_buff_inverse) +
  tm_raster() +
  tmap_mode("view")

#making con_buff_inverse a shapefile 
ca_buff_pol <- as.polygons(con_buff_inverse)

ca_buff_sf <- st_as_sf(ca_buff_pol)

#mapping to make sure all looks okay 

tm_shape(ca_buff_sf) +
   tm_polygons()  +
   tmap_mode("view")

#sampling random points from this misshapen california shpfile
sample_points_claire <- st_sample(ca_buff_sf, 4663)

#mappping to see what it looks like 
tm_shape(sample_points_claire) +
  tm_dots() +
  tmap_mode("view")

#Buffering random points out. Because we're doing this, we will need to clip the geometry one more time later on if possible (?) bc these point buffered out might mess up where are conflict points are

sample_points_buffered <- st_buffer(sample_points_claire, dist = 5000)

#Testing to make sure the buffer worked
 tm_shape(sample_points_buffered) +
   tm_polygons() +
  tm_shape(sample_points_claire) +
   tm_dots(col  = "blue") +
   tmap_mode("view")+ 
   tm_shape(con_buff)+
   tm_polygons(col  = "red")
 
 
 #cropping sample_points_buffered by conflict_points_buffered
 #This is where I can't get the code to run 
 
 #conflict_points <- st_difference(sample_points_buffered,con_buff)
 
 #so instead I'm going to try it with the function ms_erase package rmapshaper which might be faster 
 
 zeros_claire <- ms_erase(sample_points_buffered,con_buff)
 #map to see what it looks like 
 
 tm_shape(zeros_claire) +
   tm_polygons(col  = "green") +
   tmap_mode("view")
 
  tm_shape(con_buff) +
  tm_polygons(alpha = 0.5) +
tm_shape(zeros_claire) +
   tm_polygons(col = "blue")+
tmap_mode("view")
 
 #mapping it, I can see that worked! We could probably go back to Mia's way of doing zeros now that we know this function runs faster. The nice thing about mine is we should have a lot of zeros still 
 
 #Now we need to join conflict and non conflict points into one df
#I think the first step will be to make a conflict column in the zeros_claire dataframe that has conflict = 0 
 
zeros_df <- as(zeros_claire, "Spatial" )
 
zeros_df1 <-as(zeros_df, "SpatialPolygonsDataFrame")

zeros_df2 <- as(zeros_df1, "data.frame")
```






##Elevation (DEM, mean elevation)

Two separate DEM rasters were available at 7.5-arc second resolution from USGS EROS. These will need to be merged togther.

```{r}

#northern California raster

dem_norcal <- rast(here(gdrive_data, "/InputData/InputData_Files/2010_DEM_USGSEROS_Norcal/30n150w_20101117_gmted_mea075.tif"))

#southern California raster

dem_socal <- rast(here(gdrive_data, "/InputData/InputData_Files/2010_DEM_USGSEROS_Socal/30n120w_20101117_gmted_mea075.tif"))

#merging the two rasters

dem_ca <- merge(dem_norcal, dem_socal)

#plot DEM raster
tm_shape(dem_ca) +
  tm_raster() +
  tmap_mode("view")

```

##Road density

```{r}
#road_density_2014 <- rast(here(gdrive_data,"/InputData/InputData_Files/2014_UsaRoadDensity_ESRI/2014_UsaRoadDensity_ESRI.tif"))

#We're no longer using above road_density_2014 layer. Instead we'll read in the roads layer, reproject it, and export it to ArcGIS to use tool line density

#reading in roads polyline: 
roads_2015_tiger <- vect(here(gdrive_data,"/InputData/InputData_Files/2015_roads_TIGER/tl_2015_06_prisecroads.shp"))

#Reprojecting roads layer
albers_roads_2015_tiger <- project(roads_2015_tiger, ca_raster)

#Check coordinate system
crs(albers_roads_2015_tiger)

#export roads density - ONLY NEED TO DO ONCE, HAS BEEN DONE
# writeVector(albers_roads_2015_tiger,filename = here(gdrive_data, "/IntermediateData/IntermediateData_Files/2015_roads_TIGER_Formatted/albers_roads_2015_tiger"))


#read in newly created road density. Road density was created in ArcGIS Pro from Roads 2015 Tiger layer. It was created by Claire on 10/23/2022 using the ArcGIS Pro tool line density 

road_density <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/road_density_arc/road_density_arc.tif"))

#need to fix extents do not match error code
road_density_resample  = resample(road_density, ca_raster)

#masking the layer to CA
road_density_ca <- mask(road_density_resample, ca_raster)

#This road_density_ca is now ready to be read into the main model. Saving it back to gdrive as road density formatted 

writeRaster(road_density_ca, filename = here(gdrive_data, "/IntermediateData/IntermediateData_Files/road_density_formatted/road_density_ca.tif"), filetype = "GTiff", overwrite = TRUE)


```


#Adjust the data to the same geographic extent and projection



##Conflict

```{r}

```

##Elevation
```{r}
#reproject to the same crs (NAD 83, CA Albers (2011)) and resolution (~230 m) as all other data
dem_calbers <- project(dem_ca, ca_raster)

#cropping to the spatial extent of CA
dem_formatted <- mask(dem_calbers, ca_raster)

#save the new file in intermediate data

#function to save rasters, updated with correct filepath

saverast <- function(x){
  
  name <- paste0(deparse(substitute(x)), ".tif") #the name of the file will be the same as the input object, plus '.tif' at the end
  
  #write the raster into the correct folder in drive (the path will need to be updated for other files) and save it as a geotiff
  
  writeRaster(x, filename = here(gdrive_data, "/IntermediateData/IntermediateData_Files/Elevation_Formatted", name), filetype = "GTiff", overwrite = TRUE)
  
}

saverast(dem_formatted)

```



## Creating aspect and terrian 
```{r}
#aspect
# reading in elevation data 
elev <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/Elevation_Formatted/dem_formatted.tif"))

aspect <- terrain(elev,v = "aspect",neighbors = 8)
#I think this worked? 

#save the raster

aspect_name <- paste0(deparse(substitute(aspect)), ".tif") #the name of the file will be the same as the input object, plus '.tif' at the end
  
#write the raster into the correct folder in drive (the path will need to be updated for other files) and save it as a geotiff

writeRaster(aspect, filename = here(gdrive_data, "/IntermediateData/IntermediateData_Files/aspect", aspect_name), filetype = "GTiff", overwrite = TRUE)


# terrian ruggedness
terrain_ruggedness <- terrain(elev,v = "TRI",neighbors = 8)

#save the raster

tr_name <- paste0(deparse(substitute(terrain_ruggedness)), ".tif") #the name of the file will be the same as the input object, plus '.tif' at the end
  
#write the raster into the correct folder in drive (the path will need to be updated for other files) and save it as a geotiff

writeRaster(terrain_ruggedness, filename = here(gdrive_data, "/IntermediateData/IntermediateData_Files/terrain_ruggedness", tr_name), filetype = "GTiff", overwrite = TRUE)
```

#Messing around with Drought Data

#2022
```{r}
drought_2022 <- vect(here(gdrive_data, "/InputData/InputData_Files/Drought/us_drought_monitor/2022/USDM_20221025.shp"))

drought_2022_calbers <- project(drought_2022, ca_raster)

drought_2022_mask <- crop(drought_2022_calbers, ca_raster)

drought_2022_raster <- rasterize(drought_2022_mask, ca_raster)
```

```{r}
saverast <- function(x){
  
 name <- paste0(deparse(substitute(x)), ".tif") #the name of the file will be the same as the input object, plus '.tif' at the end
  
#write the raster into the correct folder in drive (the path will need to be updated for other files) and save it as a geotiff
  
writeRaster(x, filename = here(gdrive_data, "/IntermediateData/IntermediateData_Files/drought/us_drought_monitor/2022", name), filetype = "Gtiff", overwrite=TRUE)
}

saverast(drought_2022_raster)
```

# 2016
```{r}
drought_2016 <- vect(here(gdrive_data, "/InputData/InputData_Files/Drought/us_drought_monitor/2016/USDM_20161227.shp"))

drought_2016_calbers <- project(drought_2016, ca_raster)

drought_2016_mask <- crop(drought_2016_calbers, ca_raster)

drought_2016_raster <- rasterize(drought_2016_mask, ca_raster)
```

```{r}
saverast <- function(x){
  
 name <- paste0(deparse(substitute(x)), ".tif") #the name of the file will be the same as the input object, plus '.tif' at the end
  
#write the raster into the correct folder in drive (the path will need to be updated for other files) and save it as a geotiff
  
writeRaster(x, filename = here(gdrive_data, "/IntermediateData/IntermediateData_Files/drought/us_drought_monitor/2016", name), filetype = "Gtiff", overwrite=TRUE)
}

saverast(drought_2016_raster)
```

#2017
```{r}
drought_2017 <- vect(here(gdrive_data, "/InputData/InputData_Files/Drought/us_drought_monitor/2017/USDM_20170905.shp"))

drought_2017_calbers <- project(drought_2017, ca_raster)

drought_2017_mask <- crop(drought_2017_calbers, ca_raster)

drought_2017_raster <- rasterize(drought_2017_mask, ca_raster)
```

```{r}
saverast <- function(x){
  
 name <- paste0(deparse(substitute(x)), ".tif") #the name of the file will be the same as the input object, plus '.tif' at the end
  
#write the raster into the correct folder in drive (the path will need to be updated for other files) and save it as a geotiff
  
writeRaster(x, filename = here(gdrive_data, "/IntermediateData/IntermediateData_Files/drought/us_drought_monitor/2017", name), filetype = "Gtiff", overwrite=TRUE)
}

saverast(drought_2017_raster)
```

#2018
```{r}
drought_2018 <- vect(here(gdrive_data, "/InputData/InputData_Files/Drought/us_drought_monitor/2018/USDM_20180904.shp"))

drought_2018_calbers <- project(drought_2018, ca_raster)

drought_2018_mask <- crop(drought_2018_calbers, ca_raster)

drought_2018_raster <- rasterize(drought_2018_mask, ca_raster)
```

```{r}
saverast <- function(x){
  
 name <- paste0(deparse(substitute(x)), ".tif") #the name of the file will be the same as the input object, plus '.tif' at the end
  
#write the raster into the correct folder in drive (the path will need to be updated for other files) and save it as a geotiff
  
writeRaster(x, filename = here(gdrive_data, "/IntermediateData/IntermediateData_Files/drought/us_drought_monitor/2018", name), filetype = "Gtiff", overwrite=TRUE)
}

saverast(drought_2018_raster)
```

#2019
```{r}
drought_2019 <- vect(here(gdrive_data, "/InputData/InputData_Files/Drought/us_drought_monitor/2019/USDM_20190903.shp"))

drought_2019_calbers <- project(drought_2019, ca_raster)

drought_2019_mask <- crop(drought_2019_calbers, ca_raster)

drought_2019_raster <- rasterize(drought_2019_mask, ca_raster)
```

```{r}
saverast <- function(x){
  
 name <- paste0(deparse(substitute(x)), ".tif") #the name of the file will be the same as the input object, plus '.tif' at the end
  
#write the raster into the correct folder in drive (the path will need to be updated for other files) and save it as a geotiff
  
writeRaster(x, filename = here(gdrive_data, "/IntermediateData/IntermediateData_Files/drought/us_drought_monitor/2019", name), filetype = "Gtiff", overwrite=TRUE)
}

saverast(drought_2019_raster)
```

#2020
```{r}
drought_2020 <- vect(here(gdrive_data, "/InputData/InputData_Files/Drought/us_drought_monitor/2020/USDM_20200908.shp"))

drought_2020_calbers <- project(drought_2020, ca_raster)

drought_2020_mask <- crop(drought_2020_calbers, ca_raster)

drought_2020_raster <- rasterize(drought_2020_mask, ca_raster)
```

```{r}
saverast <- function(x){
  
 name <- paste0(deparse(substitute(x)), ".tif") #the name of the file will be the same as the input object, plus '.tif' at the end
  
#write the raster into the correct folder in drive (the path will need to be updated for other files) and save it as a geotiff
  
writeRaster(x, filename = here(gdrive_data, "/IntermediateData/IntermediateData_Files/drought/us_drought_monitor/2020", name), filetype = "Gtiff", overwrite=TRUE)
}

saverast(drought_2020_raster)
```

#2021
```{r}
drought_2021 <- vect(here(gdrive_data, "/InputData/InputData_Files/Drought/us_drought_monitor/2021/USDM_20210907.shp"))

drought_2021_calbers <- project(drought_2021, ca_raster)

drought_2021_mask <- crop(drought_2021_calbers, ca_raster)

drought_2021_raster <- rasterize(drought_2021_mask, ca_raster)
```

```{r}
saverast <- function(x){
  
 name <- paste0(deparse(substitute(x)), ".tif") #the name of the file will be the same as the input object, plus '.tif' at the end
  
#write the raster into the correct folder in drive (the path will need to be updated for other files) and save it as a geotiff
  
writeRaster(x, filename = here(gdrive_data, "/IntermediateData/IntermediateData_Files/drought/us_drought_monitor/2021", name), filetype = "Gtiff", overwrite=TRUE)
}

saverast(drought_2021_raster)
```