---
title: "Data Exploration on Public vs Private Lands in California"
author: "Claire Meuter"
date: "2023-03-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(terra)
library(sf)
library(stars)
library(tmap)

#claire's file path
gdrive_data <- "/Users/clairemeuter/Library/CloudStorage/GoogleDrive-clairemeuter@ucsb.edu/Shared drives/Black_Bear_Aware/gdrive_data"
```

## **Introduction:** Here, we explore the relationship between ownership of land and locations of human-black bear conflict in across California. 

## Data wrangling
My data wrangling to-do list:
- Read in my .gdl (geodatabase)
- explore maps
- Add shp files for private land (this source leaves all private land unlabeled, with areas without an ownership feature are ASSUMED to be private (but not included in the dataset as such)).
- intersect my maps 

Reading in land ownership .gdl
```{r}
land_own <- sf::st_read(here(gdrive_data, "/InputData/InputData_Files/2022_CA_ownershipFrap.gdb"), layer = "ownership22_1")

# Checking the class of land_own
class(land_own)
# sf and data.frame

```
Mapping land_own to see what it looks like 
```{r}


#tm_shape(land_own) +
 # tm_polygons() +
  #tmap_options(check.and.fix = TRUE) +
 # tmap_mode("view")
# currently, this map is massive with 52056 obs. I want to combine all of my polygons by the ownership level (Own_Level)
#Using unique(), I see that Own_Level has 7 distinctions, so hopefully at the end of this I'll have 7 multipolygons 
unique(land_own$Own_Level)
# City, County, Federal, Non Profit, Special District, State, Tribal          
```
Simplifying land_own
```{r}
#simplifying land_own 
 land_own_sim <- land_own %>% 
  group_by(Own_Level) %>% 
  summarise(Shape = sf::st_union(Shape)) %>% 
  ungroup()




# This works, but I don't really have an attribute table.. 
plot(land_own_sim)

#looking into recalculating geometrys (like in Arc) 


#saving the new file so I don't have to deal with the larger 
st_write(land_own_sim, dsn = here(gdrive_data, "/IntermediateData/IntermediateData_Files/land_own", "land_own_sim.shp"))

# I made this shp into a raster in arc because it was easier (having trouble w/ coding. If I could fix later that'd be cool )

#read in the raster I made in arc 
land_own_raster <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/land_own/own_level_raster/own_level_raster.tif"))

#exploring what the NAs look like 
plot(is.na(land_own_raster))



```

# reclassify land_own_raster with NAs as private 
```{r}
# first a little exploration 
unique(land_own_raster$Own_Level)


levels(land_own_raster)


# value   own level 
#1	City			
#2	County			
#3	Federal			
#4	Non Profit			
#5	`Special District`			
#6	State			
#7	Tribal





#Create reclassification matrix for land own values (from, to)

#list out the from, to values in an object
m <- c(1, 1, #city
       2, 2, #county
       3, 3, #federal
       4, 4, #nonprofit
       5, 5, #special district
       6, 6, #state
       7, 7, #tribal
       NA, 8) #private 

#turn those values into a matrix
mat <- matrix(m, ncol = 2, byrow = TRUE) 

#set categories (labels) for each numeric index

#index vector
index <- c(1, 2, 3, 4, 5, 6, 7, 8)

#categories vector
cats <- c("City", "County", "Federal","Non Profit", "Special District", "State", "Tribal", "Private")

#combine those into a dataframe
catdat <- data.frame(id = index, category = cats)

#Reclassify

#create reclassification function

reclassify <- function(x){
  x_reclassed <- classify(x, mat, others = NA) #reclassify using the matrix above
  
  levels(x_reclassed) <- catdat #then reassign levels (labels) using the dataframe above
  
  return(x_reclassed)
}

#Apply function to each layer



land_own_reclassed <- reclassify(land_own_raster)
#plot(land_own_reclassed)
#levels(land_own_reclassed)
#[[1]]
#  id         category
#1  1             City
#2  2           County
#3  3          Federal
#4  4       Non Profit
#5  5 Special District
#6  6            State
#7  7           Tribal
#8  8          Private

# now land_own_reclassed needs to be reprojected and clipped to the ca_raster 

#read in holy raster layer 
ca_raster <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/CARaster_Tiger_Formatted/formattedcaraster.tif"))

#make sure raster is in the same projection
land_own_reclass_reproject <- project(land_own_reclassed,ca_raster)

#mask to clear out NAs around california. land_own_mask is the raster i need!
land_own_mask <- mask(land_own_reclass_reproject, ca_raster)

plot(land_own_mask)
#looks beautiful
```
# clip the CA raster by the high conflict vect 
```{r}
#reading in present day high conflict shp layer 
high_con_sf <- vect(here(gdrive_data, "/AnalysisData/model_outputs/shapefiles/high_conflict/mod3sq_current_highconf.shp"))


#masking to high conflict areas

masked_own <- mask(land_own_mask, high_con_sf)

#finding the area of each category from masked_own 
area_by_category <- as.data.frame(masked_own) %>% 
  group_by(category) %>% 
  tally() %>% 
  mutate(area = n * res(masked_own)[1] * res(masked_own)[2]) 



```

