---
title: "Projection Data"
author: "Chase Tarr"
date: "2023-01-27"
output: html_document
---

Drought Projections uses data sourced from CalAdapt in the form of Precipitation values and Potential Evaporation. These values are then cropped to the state of California, and masked to the extent of our general California Raster in order to keep resolution and extent the same across the project (230 meters x 230 meters). The Coordinate Reference System (CRS) is coded below as well for a visualization of what we are using. Once cropped and masked, the values are combined into one dataset and loaded into the Palmer Drought Severity Index function (pdsi) where the function gives values of drought severity based on these values and other constants such as available water content (awc), etc. These outputs are then combined once again with the previous dataset in order to maintain x and y values, so then it can be rasterized for a visual output. This is then cropped and masked and saved as a tif file for our model to use. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#attach packages
#note - Loading all of the ones Jason used, I don't think we will need all of them 

library(tidyverse)
library(here)
library(terra)
library(sf)
library(tmap)
library(beepr)
library(rgeos)
library(rmapshaper)
library(raster)
library(stars)
library(purrr)
library(remotes)
library(dplyr)
library(data.table)
library(rgdal)
library(sp)
library(raster)

#mia and claire's and Chase's file path
gdrive_data <- "/Volumes/GoogleDrive/Shared drives/Black_Bear_Aware/gdrive_data"

#sam's file path
#gdrive_data <- "G:/Shared drives/Black_Bear_Aware/gdrive_data"
```

#Installing the necessary palmer drought severity index package from the original github upload
```{r}
install_github("Sibada/scPDSI")

```

#Uploading the scPDSI package the pdsi function is located within
```{r}
library(scPDSI)
```


#Uploading the california raster for reference
```{r}
ca_raster <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/CARaster_Tiger_Formatted/formattedcaraster.tif"))

```

#Reading in Potential Evapotranspiration Data
```{r}
CalAdapt_PE <- rast(here(gdrive_data, "/InputData/InputData_Files/Drought/CalAdapt/data-5/et_month_HadGEM2-CC_rcp45_2030-01.v0.CA_NV.tif")) %>%
  terra::project(ca_raster)

plot(CalAdapt_PE)

```
#Reading in Precipitation Data
```{r}
CalAdapt_P <- rast(here(gdrive_data, "/InputData/InputData_Files/Drought/CalAdapt/data-5/rainfall_month_CCSM4_rcp45_2030-01.v0.CA_NV.tif"))%>% terra::project(ca_raster)

plot(CalAdapt_P)
```

#Projecting potential evapotranspiration data to necessary CRS
```{r}
caladapt_pe_calbers <- terra::project(CalAdapt_PE, ca_raster)

#masking raster to California extent
caladapt_pe_masked <- mask(caladapt_pe_calbers, ca_raster)

#test plot
#plot(caladapt_pe_masked)
```

#Projecting precipitation data to necessary CRS
```{r}
caladapt_P_calbers <- terra::project(CalAdapt_P, ca_raster)

#masking raster to California extent
caladapt_P_masked <- mask(caladapt_P_calbers, ca_raster)

#test plot
#plot(caladapt_P_masked)
```

#Merging the rasters into one dataframe / running the PDSI function
```{r}
#merging/stacking the rasters 
caladapt_merged <- c(caladapt_pe_masked,caladapt_P_masked)

##testing to see if merged raster output is correct
#plot(caladapt_merged)

#converting to a dataframe
caladapt_merged_df <- as.data.frame(caladapt_merged, xy=TRUE) %>% 
  rename(P = "rainfall_month_CCSM4_rcp45_2030-01.v0.CA_NV",
    PE = "et_month_HadGEM2-CC_rcp45_2030-01.v0.CA_NV")

#setting names to variables
P <- caladapt_merged_df$P
PE <- caladapt_merged_df$PE

#Running PDSI function for year 2030
pdsi_df<-(pdsi(P,PE, start = 2030))

#pdsi dataframe making sure column length is the same
cal_adapt_pdsi_Df <- dplyr::mutate(caladapt_merged_df,
              pdsi_measure = (pdsi_df$X[1:7901930]))

#turning back into matrix if necessary
#cal_adapt_pdsi_df <- as.matrix(cal_adapt_pdsi_Df)

#subsetting dataframe to include precipitation and potential evaporation values
df <- subset(cal_adapt_pdsi_Df, select = -c(P,PE))

#double checking to see what values are in previous dataframe compared to new dataframe
head(cal_adapt_pdsi_Df)

head(df)

#df2 <- subset(df, select = -c(x,y))

#write.table(cal_adapt_pdsi_Df, file = here(gdrive_data, "/InputData/InputData_Files/Drought/CalAdapt/pdsi_df.csv"), row.names = FALSE)

#df_csv_write <- apply(df,2,as.character)

```
#Rasterizing drought data with raster package if necessary
#```{r}
#pdsi_raster <- rasterFromXYZ(df)

#plot(pdsi_raster)
#```


#Rasterizing Drought Data in Terra Package
```{r}
pdsi_rast_2030 <- terra::rast(df)

#plotting to see output
plot(pdsi_rast_2030)

#setting extent of drought raster to match ca_raster
crs(pdsi_rast_2030) <- "PROJCRS[\"NAD_1983_2011_California_Teale_Albers\",\n    BASEGEOGCRS[\"NAD83(2011)\",\n        DATUM[\"NAD83 (National Spatial Reference System 2011)\",\n            ELLIPSOID[\"GRS 1980\",6378137,298.257222101004,\n                LENGTHUNIT[\"metre\",1]]],\n        PRIMEM[\"Greenwich\",0,\n            ANGLEUNIT[\"degree\",0.0174532925199433]],\n        ID[\"EPSG\",6318]],\n    CONVERSION[\"Albers Equal Area\",\n        METHOD[\"Albers Equal Area\",\n            ID[\"EPSG\",9822]],\n        PARAMETER[\"Latitude of false origin\",0,\n            ANGLEUNIT[\"degree\",0.0174532925199433],\n            ID[\"EPSG\",8821]],\n        PARAMETER[\"Longitude of false origin\",-120,\n            ANGLEUNIT[\"degree\",0.0174532925199433],\n            ID[\"EPSG\",8822]],\n        PARAMETER[\"Latitude of 1st standard parallel\",34,\n            ANGLEUNIT[\"degree\",0.0174532925199433],\n            ID[\"EPSG\",8823]],\n        PARAMETER[\"Latitude of 2nd standard parallel\",40.5,\n            ANGLEUNIT[\"degree\",0.0174532925199433],\n            ID[\"EPSG\",8824]],\n        PARAMETER[\"Easting at false origin\",0,\n            LENGTHUNIT[\"metre\",1],\n            ID[\"EPSG\",8826]],\n        PARAMETER[\"Northing at false origin\",-4000000,\n            LENGTHUNIT[\"metre\",1],\n            ID[\"EPSG\",8827]]],\n    CS[Cartesian,2],\n        AXIS[\"easting\",east,\n            ORDER[1],\n            LENGTHUNIT[\"metre\",1]],\n        AXIS[\"northing\",north,\n            ORDER[2],\n            LENGTHUNIT[\"metre\",1]],\n    ID[\"EPSG\",6414]]" 

#reprojecting to ca_raster if crs does not work
caladapt_Pdsi_calbers <- terra::project(pdsi_rast_2030, ca_raster)

#mask to CA
caladapt_pdsi_masked <- mask(caladapt_Pdsi_calbers, ca_raster)

plot(pdsi_rast_2030)

pdsi_rast_2030_mr <- pdsi_rast_2030

pdsi_calbers <- terra::project(pdsi_rast_2030_mr, ca_raster)

pdsi_masked <- mask(pdsi_calbers, ca_raster)
```

#Saving raster into model ready folder for output
```{r}
saverast <- function(x){
  
 name <- paste0(deparse(substitute(x)), ".tif") #the name of the file will be the same as the input object, plus '.tif' at the end
  
#write the raster into the correct folder in drive (the path will need to be updated for other files) and save it as a geotiff
  
writeRaster(x, filename = here(gdrive_data, "/IntermediateData/IntermediateData_Files/model_ready/drought", name), filetype = "Gtiff", overwrite=TRUE)
}

saverast(pdsi_rast_2030)

saverast(pdsi_rast_2030_mr)

saverast(pdsi_masked)
plot(pdsi_masked)
```



