---
title: "Drought aggregation"
author: "Mia Guarnieri"
date: "2023-01-30"
output: html_document
---

#Summary

This Rmd contains the instructions and code for calculating the mean Palmer Drought Severity Index (PDSI) value from weekly data taken from US Drought Monitor. We analysed data from 2016 to 2021 here. Prior to using this analysis technique, all shapefiles for a given year were moved into one folder (one folder for 2016 shapefiles, one folder for 2017 shapefiles, etc).

# Load in necessary packages (and file path)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#reading in the required packages

library(tidyverse)
library(here)
library(terra)
library(sf)
library(tmap)
library(rgdal)

#setting up a file path - useful if your data is being read in from a Google Drive, your desktop, or somewhere other than your designated working directory
gdrive_data <- "/Users/mia/Library/CloudStorage/GoogleDrive-mguarnieri@ucsb.edu/Shared drives/Black_Bear_Aware/gdrive_data"
```


# Drought data wrangling for 2016

```{r}
#create a list of the shapefile names within the folder
d16_list <- list.files(path = here(gdrive_data, "/InputData/InputData_Files/Drought/USDM_shapefiles/2016_USDM_M"), pattern = '.shp')

#For each file in that list, convert it to a raster and save it to an intermediate data folder
for (i in d16_list) {

#load in already formatted raster for a rasterization template   
caraster <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/CARaster_Tiger_Formatted/formattedcaraster.tif"))

#read in the drought data shapefile  
drought <- read_sf(here(gdrive_data, "/InputData/InputData_Files/Drought/USDM_shapefiles/2016_USDM_M", i))

#turn that file into a spatvector  
drought_vect <- vect(drought) %>% 
  terra::project(., caraster)

#rasterize the spatvector using the caraster as a template, designate which field to rasterize (PDSI)  
drought_rast <- rasterize(drought_vect, caraster, field = "DM")

#crop the data to be only within the extent of the caraster file (the state of California)  
drought_rast_ca <- mask(drought_rast, caraster)

#the name of the file will be the same as the input file, plus '.tif' at the end
name <- paste0(i, ".tif") 

#save the raster to the designated folder  
  writeRaster(drought_rast_ca, filename = here(gdrive_data, "/IntermediateData/IntermediateData_Files/drought/drought_rasters/UDM_2016", name), filetype = "GTiff", overwrite = TRUE)
  
}

#create a list of raster file names from within the folder designated inside of the for loop
d16_rast_list <- list.files(path = here(gdrive_data, "/IntermediateData/IntermediateData_Files/drought/drought_rasters/UDM_2016"))

#stack all those raster files
stack16 <- c(rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/drought/drought_rasters/UDM_2016/", d16_rast_list)))

#create a new raster with the mean of all of those raster files
mean_16 <- terra::mean(stack16, na.rm = TRUE)

#save the mean raster - this is the annual mean PDSI value within each pixel
writeRaster(mean_16, filename = here(gdrive_data, "/IntermediateData/IntermediateData_Files/drought/drought_rasters/UDM_2016/mean_drought_16.tif"), filetype = "GTiff", overwrite = TRUE)
  
```

# Drought data wrangling for 2017

```{r}
#create a list of the shapefile names
d17_list <- list.files(path = here(gdrive_data, "/InputData/InputData_Files/Drought/USDM_shapefiles/2017_USDM_M"), pattern = '.shp')

#read each file in that list, convert it to a raster, and save it to an intermediate data folder (same process as above)
for (i in d17_list) {
  
caraster <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/CARaster_Tiger_Formatted/formattedcaraster.tif"))
  
drought <- read_sf(here(gdrive_data, "/InputData/InputData_Files/Drought/USDM_shapefiles/2017_USDM_M", i))
  
drought_vect <- vect(drought) %>% 
  terra::project(., caraster)
  
drought_rast <- rasterize(drought_vect, caraster, field = "DM")
  
drought_rast_ca <- mask(drought_rast, caraster)
  
name <- paste0(i, ".tif") #the name of the file will be the same as the input file, plus '.tif' at the end
  
  writeRaster(drought_rast_ca, filename = here(gdrive_data, "/IntermediateData/IntermediateData_Files/drought/drought_rasters/UDM_2017", name), filetype = "GTiff", overwrite = TRUE)
  
}

#create a list of raster names within the folder
d17_rast_list <- list.files(path = here(gdrive_data, "/IntermediateData/IntermediateData_Files/drought/drought_rasters/UDM_2017"))

#stack all those raster files
stack17 <- c(rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/drought/drought_rasters/UDM_2017/", d17_rast_list)))

#create a new raster with annual mean PSDI values
mean_17 <- terra::mean(stack17, na.rm = TRUE)

#save the mean raster
writeRaster(mean_17, filename = here(gdrive_data, "/IntermediateData/IntermediateData_Files/drought/drought_rasters/UDM_2017/mean_drought_17.tif"), filetype = "GTiff", overwrite = TRUE)

```

# Drought data wrangling for 2018

```{r}
#create a list of the shapefile names
d18_list <- list.files(path = here(gdrive_data, "/InputData/InputData_Files/Drought/USDM_shapefiles/2018_USDM_M"), pattern = '.shp')

#read each file in that list, convert it to a raster, and save it to an intermediate data folder (same process as above)
for (i in d18_list) {
  
caraster <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/CARaster_Tiger_Formatted/formattedcaraster.tif"))
  
drought <- read_sf(here(gdrive_data, "/InputData/InputData_Files/Drought/USDM_shapefiles/2018_USDM_M", i))
  
drought_vect <- vect(drought) %>% 
  terra::project(., caraster)
  
drought_rast <- rasterize(drought_vect, caraster, field = "DM")
  
drought_rast_ca <- mask(drought_rast, caraster)
  
name <- paste0(i, ".tif") #the name of the file will be the same as the input file, plus '.tif' at the end
  
  writeRaster(drought_rast_ca, filename = here(gdrive_data, "/IntermediateData/IntermediateData_Files/drought/drought_rasters/UDM_2018", name), filetype = "GTiff", overwrite = TRUE)
  
}

#create a list of raster names within the folder
d18_rast_list <- list.files(path = here(gdrive_data, "/IntermediateData/IntermediateData_Files/drought/drought_rasters/UDM_2018"))

#stack all those raster files
stack18 <- c(rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/drought/drought_rasters/UDM_2018/", d18_rast_list)))

#create a new raster with annual mean PSDI values
mean_18 <- terra::mean(stack18, na.rm = TRUE)

#save the mean raster
writeRaster(mean_18, filename = here(gdrive_data, "/IntermediateData/IntermediateData_Files/drought/drought_rasters/UDM_2018/mean_drought_18.tif"), filetype = "GTiff", overwrite = TRUE)
```

# Drought data wrangling for 2019

```{r}
#create a list of the shapefile names
d19_list <- list.files(path = here(gdrive_data, "/InputData/InputData_Files/Drought/USDM_shapefiles/2019_USDM_M"), pattern = '.shp')

#read each file in that list, convert it to a raster, and save it to an intermediate data folder (same process as above)
for (i in d19_list) {
  
caraster <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/CARaster_Tiger_Formatted/formattedcaraster.tif"))
  
drought <- read_sf(here(gdrive_data, "/InputData/InputData_Files/Drought/USDM_shapefiles/2019_USDM_M", i))
  
drought_vect <- vect(drought) %>% 
  terra::project(., caraster)
  
drought_rast <- rasterize(drought_vect, caraster, field = "DM")
  
drought_rast_ca <- mask(drought_rast, caraster)
  
name <- paste0(i, ".tif") #the name of the file will be the same as the input file, plus '.tif' at the end
  
  writeRaster(drought_rast_ca, filename = here(gdrive_data, "/IntermediateData/IntermediateData_Files/drought/drought_rasters/UDM_2019", name), filetype = "GTiff", overwrite = TRUE)
  
}

#create a list of raster names within the folder
d19_rast_list <- list.files(path = here(gdrive_data, "/IntermediateData/IntermediateData_Files/drought/drought_rasters/UDM_2019"))

#stack all those raster files
stack19 <- c(rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/drought/drought_rasters/UDM_2019/", d19_rast_list)))

#create a new raster with annual mean PSDI values
mean_19 <- terra::mean(stack19, na.rm = TRUE)

#save the mean raster
writeRaster(mean_19, filename = here(gdrive_data, "/IntermediateData/IntermediateData_Files/drought/drought_rasters/UDM_2019/mean_drought_19.tif"), filetype = "GTiff", overwrite = TRUE)
```

# Drought data wrangling for 2020

```{r}
#create a list of the shapefile names
d20_list <- list.files(path = here(gdrive_data, "/InputData/InputData_Files/Drought/USDM_shapefiles/2020_USDM_M"), pattern = '.shp')

#read each file in that list, convert it to a raster, and save it to an intermediate data folder (same process as above)
for (i in d20_list) {
  
caraster <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/CARaster_Tiger_Formatted/formattedcaraster.tif"))
  
drought <- read_sf(here(gdrive_data, "/InputData/InputData_Files/Drought/USDM_shapefiles/2020_USDM_M", i))
  
drought_vect <- vect(drought) %>% 
  terra::project(., caraster)
  
drought_rast <- rasterize(drought_vect, caraster, field = "DM")
  
drought_rast_ca <- mask(drought_rast, caraster)
  
name <- paste0(i, ".tif") #the name of the file will be the same as the input file, plus '.tif' at the end
  
  writeRaster(drought_rast_ca, filename = here(gdrive_data, "/IntermediateData/IntermediateData_Files/drought/drought_rasters/UDM_2020", name), filetype = "GTiff", overwrite = TRUE)
  
}

#create a list of raster names within the folder
d20_rast_list <- list.files(path = here(gdrive_data, "/IntermediateData/IntermediateData_Files/drought/drought_rasters/UDM_2020"))

#stack all those raster files
stack20 <- c(rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/drought/drought_rasters/UDM_2020/", d20_rast_list)))

#create a new raster with annual mean PSDI values
mean_20 <- terra::mean(stack20, na.rm = TRUE)

#save the mean raster
writeRaster(mean_20, filename = here(gdrive_data, "/IntermediateData/IntermediateData_Files/drought/drought_rasters/UDM_2020/mean_drought_20.tif"), filetype = "GTiff", overwrite = TRUE)

```

# Drought data wrangling for 2021

```{r}
#create a list of the shapefile names
d21_list <- list.files(path = here(gdrive_data, "/InputData/InputData_Files/Drought/USDM_shapefiles/2021_USDM_M"), pattern = '.shp')

#read each file in that list, convert it to a raster, and save it to an intermediate data folder (same process as above)
for (i in d21_list) {
  
caraster <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/CARaster_Tiger_Formatted/formattedcaraster.tif"))
  
drought <- read_sf(here(gdrive_data, "/InputData/InputData_Files/Drought/USDM_shapefiles/2021_USDM_M", i))
  
drought_vect <- vect(drought) %>% 
  terra::project(., caraster)
  
drought_rast <- rasterize(drought_vect, caraster, field = "DM")
  
drought_rast_ca <- mask(drought_rast, caraster)
  
name <- paste0(i, ".tif") #the name of the file will be the same as the input file, plus '.tif' at the end
  
  writeRaster(drought_rast_ca, filename = here(gdrive_data, "/IntermediateData/IntermediateData_Files/drought/drought_rasters/UDM_2021", name), filetype = "GTiff", overwrite = TRUE)
  
}

#create a list of raster names within the folder
d21_rast_list <- list.files(path = here(gdrive_data, "/IntermediateData/IntermediateData_Files/drought/drought_rasters/UDM_2021"))

#stack all those raster files
stack21 <- c(rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/drought/drought_rasters/UDM_2021/", d21_rast_list)))

#create a new raster with annual mean PSDI values
mean_21 <- terra::mean(stack21, na.rm = TRUE)

#save the mean raster
writeRaster(mean_21, filename = here(gdrive_data, "/IntermediateData/IntermediateData_Files/drought/drought_rasters/UDM_2021/mean_drought_21.tif"), filetype = "GTiff", overwrite = TRUE)

```

