---
title: "fire_projection_wrangling"
author: "Claire Meuter"
date: "2023-04-26"
output: html_document
---

# Setup and Packages
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(here)
library(terra)
library(sf)
library(tmap)
library(beepr)
library(rgeos)
library(rmapshaper)

#claire's file path
gdrive_data <- "/Users/clairemeuter/Library/CloudStorage/GoogleDrive-clairemeuter@ucsb.edu/Shared drives/Black_Bear_Aware/gdrive_data"

```
# Purpose: 
The purpose of this .rmd is to create distance to fire layers for use in projecting our model on the year 2030. 
These data layers are sourced from Isaac Park's work on modeling fire likelihood across California from 2024  to 2029.  https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0254723#sec015
We processed two sets of fire projections- one for a climate scenario of RCP 4.5 and one for a climate scenario at an RCP 8.5. 
See our team's Dryad page for further information on the source data.  

This .rmd resulted in the creation of 6 distance rasters: 
- Distance to projected fire (under RCP 4.5) from 2025 to 2026
- Distance to projected fire (under RCP 4.5) from 2027 to 2028
- Distance to projected fire (under RCP 4.5) from 2029
- Distance to projected fire (under RCP 8.5) from 2025 to 2026
- Distance to projected fire (under RCP 8.5) from 2027 to 2028
- Distance to projected fire (under RCP 8.5) from 2029

## Data wrangling for RCP 4.5 
```{r}
# CA Raster example 

#reading in ca_raster, a raster formatted to the resolution, crs, and spatial extent we plan to run our model over. 

ca_raster <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/CARaster_Tiger_Formatted/formattedcaraster.tif"))

#check the coordinate reference system 
crs(ca_raster)
```
Fire layers from Isaac Park from his publication https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0254723#sec015
```{r}
#read in fire layers 

#2024 to 2029 sequence 
fire_pred4_2024 <- rast(here(gdrive_data, "/InputData/InputData_Files/rcp4.5_AnnualFireProjections_Isaac/PredRisk_2024_2024LogGam_Class.tif"))

fire_pred4_2025 <- rast(here(gdrive_data, "/InputData/InputData_Files/rcp4.5_AnnualFireProjections_Isaac/PredRisk_2025_2025LogGam_Class.tif"))

fire_pred4_2026 <- rast(here(gdrive_data, "/InputData/InputData_Files/rcp4.5_AnnualFireProjections_Isaac/PredRisk_2026_2026LogGam_Class.tif"))

fire_pred4_2027 <- rast(here(gdrive_data, "/InputData/InputData_Files/rcp4.5_AnnualFireProjections_Isaac/PredRisk_2027_2027LogGam_Class.tif"))

fire_pred4_2028 <- rast(here(gdrive_data, "/InputData/InputData_Files/rcp4.5_AnnualFireProjections_Isaac/PredRisk_2028_2028LogGam_Class.tif"))

fire_pred4_2029 <- rast(here(gdrive_data, "/InputData/InputData_Files/rcp4.5_AnnualFireProjections_Isaac/PredRisk_2029_2029LogGam_Class.tif"))

```

### Getting 4.5 layers in correct extent, resolution and projection
```{r}
#reproject to the same crs (NAD 83, CA Albers (2011))

fire_pred4_2024_albers <- terra::project(fire_pred4_2024, ca_raster)

fire_pred4_2025_albers <- terra::project(fire_pred4_2025, ca_raster)

fire_pred4_2026_albers <- terra::project(fire_pred4_2026, ca_raster)

fire_pred4_2027_albers <- terra::project(fire_pred4_2027, ca_raster)

fire_pred4_2028_albers <- terra::project(fire_pred4_2028, ca_raster)

fire_pred4_2029_albers <- terra::project(fire_pred4_2029, ca_raster)

```

#### Cropping the extents to our CA raster 

```{r}
#cropping to spatial extent of CA raster

fire_pred4_2024_mask <- mask(fire_pred4_2024_albers, ca_raster)

fire_pred4_2025_mask <- mask(fire_pred4_2025_albers, ca_raster)

fire_pred4_2026_mask <- mask(fire_pred4_2026_albers, ca_raster)

fire_pred4_2027_mask <- mask(fire_pred4_2027_albers, ca_raster)

fire_pred4_2028_mask <- mask(fire_pred4_2028_albers, ca_raster)

fire_pred4_2029_mask <- mask(fire_pred4_2029_albers, ca_raster)
```

### Reclassifying fire projection rasters for RCP 4.5
Now I need to reclassify to only keep raster values above .03 (small amount of projected fire likelihood, as a proxy for burn scars) I'm calling this high probability of fire. (this may change if there is better justification in the literature). This is our proxy for med/high intensity fires/burn scars that we selected for in our model. 

```{r}

#list out the from, to values in an object
m <- c(0, 0.03, NA, #values lower than 0.03 to 0
       0.03, 5, 1) #values greater than 0.03 to 1
       
#turn those values into a matrix
mat_fire <- matrix(m, ncol = 3, byrow = TRUE) 

fire_pred4_2024_reclassed <- classify(fire_pred4_2024_mask, mat_fire,include.lowest=TRUE)
fire_pred4_2025_reclassed <- classify(fire_pred4_2025_mask, mat_fire,include.lowest=TRUE)
fire_pred4_2026_reclassed <- classify(fire_pred4_2026_mask, mat_fire,include.lowest=TRUE)
fire_pred4_2027_reclassed <- classify(fire_pred4_2027_mask, mat_fire,include.lowest=TRUE)
fire_pred4_2028_reclassed <- classify(fire_pred4_2028_mask, mat_fire,include.lowest=TRUE)
fire_pred4_2029_reclassed <- classify(fire_pred4_2029_mask, mat_fire,include.lowest=TRUE)
# map to see what it looks like 
#tm_shape(fire_pred4_2024_reclassed) +
 # tm_raster() +
  #tmap_mode("view")
```


### Merge layers
Here I merge year layers to match our variable descriptions for burn scars, which are: 4 to 5 years ago, 2 to 3 years ago, and one year ago. 
```{r}
#4-5 year: combined 2025-2026

fire_pred4_2025_reclassed

fire_pred4_2026_reclassed

#merge those layers

fire_2025_2026 <- merge(fire_pred4_2025_reclassed, fire_pred4_2026_reclassed)

#tm_shape(fire_2025_2026) + 
  #tm_raster() + 
 # tmap_mode("view")

#2-3 year: combined 2027-2028

fire_pred4_2027_reclassed

fire_pred4_2028_reclassed

#merge those layers

fire_2027_2028 <- merge(fire_pred4_2027_reclassed, fire_pred4_2028_reclassed)

#tm_shape(fire_pred4_2029_reclassed) + 
#tm_raster() + 
#tmap_mode("view")

#save the layers

#function for saving multi-year rasters
savemultirast <- function(x){
  
  name <- paste0(deparse(substitute(x)), ".tif") #the name of the file will be the same as the input object, plus '.tif' at the end
  
  #write the raster into the correct folder in drive (the path will need to be updated for other files) and save it as a geotiff
  
  writeRaster(x, filename = here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/proj_fire4.5", name), filetype = "GTiff", overwrite = TRUE)
  
}

#save the data using the function 
savemultirast(fire_2025_2026)

savemultirast(fire_2027_2028)

savemultirast(fire_pred4_2029_reclassed)


```


### Distance to Fire RCP 4.5  Arc Layer
Due to the size of our rasters, using R to create distance layers was not feasible, and instead our group used ArcGIS to create these distance layers. We then reformatted these layers and saved them to our folder "model_ready" indicating they were ready for use in our final model code

```{r}
#Now we need to read back in the distance raster created in ArcGIS pro and mask these layers to match our CA extent again. Layers are already in the correct projection and resolution

# read in the data 
dist_proj_fire_2025_2026 <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/dist_to_proj_fire4.5/dist_proj_fire4.5_2025_2026.tif"))

dist_proj_fire_2027_2028 <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/dist_to_proj_fire4.5/dist_proj_fire4.5_2027_2028.tif"))

dist_proj_fire_2029 <-rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/dist_to_proj_fire4.5/dist_proj_fire4.5_2029.tif"))

# Now we need to mask these layers to just the CA extent 

dist_proj_fire_2025_2026_ca4.5 <- mask(dist_proj_fire_2025_2026, ca_raster)

dist_proj_fire_2027_2028_ca4.5 <- mask(dist_proj_fire_2027_2028, ca_raster)

dist_proj_fire_2029_ca4.5 <- mask(dist_proj_fire_2029, ca_raster)

```
### Save clipped fire projection DISTANCE rasters to the data drive for use in model and easy access for RCP 4.5 
```{r}
#write a function to make things easier
saverast <- function(x){
  
  name <- paste0(deparse(substitute(x)), ".tif") #the name of the file will be the same as the input object, plus '.tif' at the end
  
  #write the raster into the correct folder in drive (the path will need to be updated for other files) and save it as a geotiff
  
  writeRaster(x, filename = here(gdrive_data, "/IntermediateData/IntermediateData_Files/model_ready/dist_proj_fire/dist_proj_fire4.5", name), filetype = "GTiff", overwrite = TRUE)
  
}

saverast(dist_proj_fire_2025_2026_ca4.5)

saverast(dist_proj_fire_2027_2028_ca4.5)

saverast(dist_proj_fire_2029_ca4.5)



```


## Data Wrangling for RCP 8.5 

```{r}
#read in fire layers 

#2024 to 2029 sequence 
fire_pred8_2024 <- rast(here(gdrive_data, "/InputData/InputData_Files/01_31_23_AnnualFireProjections_Isaac/PredRisk_2024_2024LogGam_Class.tif"))

fire_pred8_2025 <- rast(here(gdrive_data, "/InputData/InputData_Files/01_31_23_AnnualFireProjections_Isaac/PredRisk_2025_2025LogGam_Class.tif"))

fire_pred8_2026 <- rast(here(gdrive_data, "/InputData/InputData_Files/01_31_23_AnnualFireProjections_Isaac/PredRisk_2026_2026LogGam_Class.tif"))

fire_pred8_2027 <- rast(here(gdrive_data, "/InputData/InputData_Files/01_31_23_AnnualFireProjections_Isaac/PredRisk_2027_2027LogGam_Class.tif"))

fire_pred8_2028 <- rast(here(gdrive_data, "/InputData/InputData_Files/01_31_23_AnnualFireProjections_Isaac/PredRisk_2028_2028LogGam_Class.tif"))

fire_pred8_2029 <- rast(here(gdrive_data, "/InputData/InputData_Files/01_31_23_AnnualFireProjections_Isaac/PredRisk_2029_2029LogGam_Class.tif"))

```

### Getting 8.5 layers in correct extent, resolution and projection
```{r}
#reproject to the same crs (NAD 83, CA Albers (2011))

fire_pred8_2024_albers <- terra::project(fire_pred8_2024, ca_raster)

fire_pred8_2025_albers <- terra::project(fire_pred8_2025, ca_raster)

fire_pred8_2026_albers <- terra::project(fire_pred8_2026, ca_raster)

fire_pred8_2027_albers <- terra::project(fire_pred8_2027, ca_raster)

fire_pred8_2028_albers <- terra::project(fire_pred8_2028, ca_raster)

fire_pred8_2029_albers <- terra::project(fire_pred8_2029, ca_raster)

```

#### Cropping the extent to our CA raster 

```{r}
#cropping to spatial extent of CA raster

fire_pred8_2024_mask <- mask(fire_pred8_2024_albers, ca_raster)

fire_pred8_2025_mask <- mask(fire_pred8_2025_albers, ca_raster)

fire_pred8_2026_mask <- mask(fire_pred8_2026_albers, ca_raster)

fire_pred8_2027_mask <- mask(fire_pred8_2027_albers, ca_raster)

fire_pred8_2028_mask <- mask(fire_pred8_2028_albers, ca_raster)

fire_pred8_2029_mask <- mask(fire_pred8_2029_albers, ca_raster)
```

### Reclassifying fire projection rasters for RCP 8.5
Now I need to reclassify to only keep raster values above .03 (small amount of projected fire likelihood, as a proxy for burn scars) I'm calling this high probability of fire. (this may change if there is better justification in the literature). This is our proxy for med/high intensity fires/burn scars that we selected for in our model. 

```{r}


#list out the from, to values in an object
m <- c(0, 0.03, NA, #values lower than 0.9 to 0
       0.03, 5, 1) #values greater than 0.9 to 1
       
#turn those values into a matrix
mat_fire <- matrix(m, ncol = 3, byrow = TRUE) 

fire_pred8_2024_reclassed <- classify(fire_pred8_2024_mask, mat_fire,include.lowest=TRUE)
fire_pred8_2025_reclassed <- classify(fire_pred8_2025_mask, mat_fire,include.lowest=TRUE)
fire_pred8_2026_reclassed <- classify(fire_pred8_2026_mask, mat_fire,include.lowest=TRUE)
fire_pred8_2027_reclassed <- classify(fire_pred8_2027_mask, mat_fire,include.lowest=TRUE)
fire_pred8_2028_reclassed <- classify(fire_pred8_2028_mask, mat_fire,include.lowest=TRUE)
fire_pred8_2029_reclassed <- classify(fire_pred8_2029_mask, mat_fire,include.lowest=TRUE)
# map to see what it looks like 
#tm_shape(fire_pred8_2024_reclassed) +
 # tm_raster() +
  #tmap_mode("view")
```


### Merge layers
Here I merge year layers to match our variable descriptions for burn scars, which are: 4 to 5 years ago, 2 to 3 years ago, and one year ago. 

```{r}
#4-5 year: combined 2025-2026

fire_pred8_2025_reclassed

fire_pred8_2026_reclassed

#merge those layers

fire8_2025_2026 <- merge(fire_pred8_2025_reclassed, fire_pred8_2026_reclassed)

#tm_shape(fire_2025_2026) + 
  #tm_raster() + 
 # tmap_mode("view")

#2-3 year: combined 2027-2028

fire_pred8_2027_reclassed

fire_pred8_2028_reclassed

#merge those layers

fire8_2027_2028 <- merge(fire_pred8_2027_reclassed, fire_pred8_2028_reclassed)

#tm_shape(fire_pred8_2029_reclassed) + 
#tm_raster() + 
#tmap_mode("view")

#save the layers

#function for saving multi-year rasters
savemultirast <- function(x){
  
  name <- paste0(deparse(substitute(x)), ".tif") #the name of the file will be the same as the input object, plus '.tif' at the end
  
  #write the raster into the correct folder in drive (the path will need to be updated for other files) and save it as a geotiff
  
  writeRaster(x, filename = here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/proj_fire8.5", name), filetype = "GTiff", overwrite = TRUE)
  
}

#save the data using the function 
savemultirast(fire8_2025_2026)

savemultirast(fire8_2027_2028)

savemultirast(fire_pred8_2029_reclassed)


```

### Distance to Fire RCP 8.5  Arc Layer
Due to the size of our rasters, using R to create distance layers was not feasible, and instead our group used ArcGIS to create these distance layers. We then reformatted these layers and saved them to our folder "model_ready" indicating they were ready for use in our final model code

```{r}
#Now we need to read back in the distance raster created in ArcGIS pro and mask these layers to match our CA extent again. Layers are already in the correct projection and resolution

# read in the data 
dist_proj_fire_2025_2026 <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/dist_to_proj_fire/dist_proj_fire_2025_2026.tif"))

dist_proj_fire_2027_2028 <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/dist_to_proj_fire/dist_proj_fire_2027_2028.tif"))

dist_proj_fire_2029 <-rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/dist_to_proj_fire/dist_proj_fire_2029.tif"))

# Now we need to mask these layers to just the CA extent 

dist_proj_fire8_2025_2026_ca <- mask(dist_proj_fire_2025_2026, ca_raster)

dist_proj_fire8_2027_2028_ca <- mask(dist_proj_fire_2027_2028, ca_raster)

dist_proj_fire8_2029_ca <- mask(dist_proj_fire_2029, ca_raster)

```
### Save clipped fire projection DISTANCE rasters to the data drive for use in model and easy access for RCP 8.5 

```{r}
#write a function to make things easier
saverast <- function(x){
  
  name <- paste0(deparse(substitute(x)), ".tif") #the name of the file will be the same as the input object, plus '.tif' at the end
  
  #write the raster into the correct folder in drive (the path will need to be updated for other files) and save it as a geotiff
  
  writeRaster(x, filename = here(gdrive_data, "/IntermediateData/IntermediateData_Files/model_ready/dist_proj_fire/dist_proj_fire8.5", name), filetype = "GTiff", overwrite = TRUE)
  
}

saverast(dist_proj_fire8_2025_2026_ca)

saverast(dist_proj_fire8_2027_2028_ca)

saverast(dist_proj_fire8_2029_ca)



```

