---
title: "demographic_exploration"
author: "Sam Rozal"
date: "2023-03-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#attach packages
library(tidyverse)
library(here)
library(sf)
library(tmap)
library(lubridate)
library(ResourceSelection)
library(rmapshaper)
library(terra)

#mia, chase, and claire's file path

#gdrive_data <- "/Volumes/GoogleDrive/Shared drives/Black_Bear_Aware/gdrive_data"

#mia's file path
gdrive_data <- "/Users/mia/Library/CloudStorage/GoogleDrive-mguarnieri@ucsb.edu/Shared drives/Black_Bear_Aware/gdrive_data"

#sam's file path
#gdrive_data <- "G:/Shared drives/Black_Bear_Aware/gdrive_data"

```

# Summary 
This rmd includes the code for a comparative statistical analysis between the demograhics of upotential underepoting regions and reported regions of high conflict. Underreported regions are defined as areas of high conflict, but no reporting. The demographic factors tested in this analysis include ethnic/racial identity, social vulnerability index, and english proficiency. 


# Exploring high conflict points where nonreporting is occurring and understanding the demographics of the areas

#Read in the model data 
```{r}
#loading in data
#model 3 raster - raster of high conflict
mod3_rast <- rast(here(gdrive_data, "/AnalysisData/model_outputs/projected_mod3_clim_map_squared.tif"))

#raster of california
ca_raster <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/CARaster_Tiger_Formatted/formattedcaraster.tif"))

#county_demographics <- vect(here(gdrive_data, "/IntermediateData/IntermediateData_Files/census_demographics/MSSA_Detail.shp"))

```

#Read in the demograhic data - Data was foreked from a github repository by njdepsky. source: https://github.com/njdepsky/CA-POP which derived high resoulion grided estimates from the 2020 census. This paper by Depsky et al. can be found here: https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0270746 

```{r}

ca_amind_alone <- rast(here(gdrive_data,"/InputData/InputData_Files/demographics_data_2020/CAPOP_2020_100m_AMIND_ALONE.tif/CAPOP_2020_100m_AMIND_ALONE.tif"))

ca_amind_combo <- rast(here(gdrive_data,"/InputData/InputData_Files/demographics_data_2020/CAPOP_2020_100m_AMIND_COMBO.tif/CAPOP_2020_100m_AMIND_COMBO.tif"))

ca_hisp <- rast(here(gdrive_data,"/InputData/InputData_Files/demographics_data_2020/CAPOP_2020_100m_HISP.tif/CAPOP_2020_100m_HISP.tif"))

ca_black <- rast(here(gdrive_data,"/InputData/InputData_Files/demographics_data_2020/CAPOP_2020_100m_NHBLACK.tif/CAPOP_2020_100m_NHBLACK.tif"))

ca_minors <- rast(here(gdrive_data,"/InputData/InputData_Files/demographics_data_2020/CAPOP_2020_100m_MINORS.tif/CAPOP_2020_100m_MINORS.tif"))

ca_nhamind <- rast(here(gdrive_data, "/InputData/InputData_Files/demographics_data_2020/CAPOP_2020_100m_NHAMIND.tif/CAPOP_2020_100m_NHAMIND.tif"))

ca_nhasian <- rast(here(gdrive_data,"/InputData/InputData_Files/demographics_data_2020/CAPOP_2020_100m_NHASIAN.tif/CAPOP_2020_100m_NHASIAN.tif"))

ca_nhhipi <- rast(here(gdrive_data,"/InputData/InputData_Files/demographics_data_2020/CAPOP_2020_100m_NHHIPI.tif/CAPOP_2020_100m_NHHIPI.tif"))

ca_nhothermulti <- rast(here(gdrive_data,"/InputData/InputData_Files/demographics_data_2020/CAPOP_2020_100m_NHOTHERMULTI.tif/CAPOP_2020_100m_OTHERMULTI.tif"))

ca_nhwhite <- rast(here(gdrive_data, "/InputData/InputData_Files/demographics_data_2020/CAPOP_2020_100m_NHWHITE.tif/CAPOP_2020_100m_NHWHITE.tif"))

ca_total <- rast(here(gdrive_data,"/InputData/InputData_Files/demographics_data_2020/CAPOP_2020_100m_TOTAL.tif/CAPOP_2020_100m_TOTAL.tif"))

```


```{r}
ca_amind_alone_proj <- project(ca_amind_alone,"epsg:6414")

ca_amind_combo_proj <- project(ca_amind_combo, "epsg:6414")

ca_hisp_proj <- project(ca_hisp, "epsg:6414")

ca_minors_proj <- project(ca_minors, "epsg:6414")

ca_nhamind_proj <- project(ca_nhamind,"epsg:6414")

ca_nhasain_proj <- project(ca_nhasian,"epsg:6414")

ca_nhhipi_proj <- project(ca_nhhipi,"epsg:6414")

ca_nhothermulti_proj <- project(ca_nhothermulti, "epsg:6414")

ca_nhwhite_proj <- project(ca_nhwhite,"epsg:6414")

ca_total_proj <- project(ca_total, "epsg:6414")

ca_black_proj <- project(ca_black, "epsg:6414")
```

```{r}
demographic_stack <- c(ca_amind_alone_proj, ca_amind_combo_proj, ca_hisp_proj, ca_minors_proj, ca_nhamind_proj, ca_nhasain_proj, ca_nhhipi_proj, ca_nhothermulti_proj, ca_nhwhite_proj, ca_total_proj, ca_black_proj)

```

```{r}
#read in model data 

high_conflict <- rast(here(gdrive_data, "/AnalysisData/model_outputs/mod3sq_current_highconf.tif"))

buffer_conflict <- vect(here(gdrive_data, "/IntermediateData/IntermediateData_Files/model_ready/conflict_buffered_refined/conflict_buffered_refined.shp"))

```

```{r}
high_conflict_reporting <- mask(high_conflict, buffer_conflict)

high_conflict_reporting_proj <- project(high_conflict_reporting,ca_total_proj)
```

```{r}
high_conflict_no_reporting <- mask(high_conflict, buffer_conflict, inverse=TRUE)

high_conflict_no_reporting_proj <- project(high_conflict_no_reporting, ca_total_proj)

```

```{r}
demographic_mask_reporting <- mask(demographic_stack, high_conflict_reporting_proj)

demographic_mask_no_reporting <- mask(demographic_stack, high_conflict_no_reporting_proj)
```

```{r}
reporting_data_frame <- as.data.frame(demographic_mask_reporting)

no_reporting_data_frame <- as.data.frame(demographic_mask_no_reporting)

```

```{r}

reporting_df_filter<- reporting_data_frame %>% select(CAPOP_2020_100m_HISP,
                                                      CAPOP_2020_100m_NHBLACK,
                                                      CAPOP_2020_100m_AMIND_COMBO,
                                                      CAPOP_2020_100m_NHASIAN,
                                                      CAPOP_2020_100m_NHHIPI,
                                                      CAPOP_2020_100m_NHWHITE,
                                                      CAPOP_2020_100m_OTHERMULTI)

no_reporting_df_filter <- no_reporting_data_frame %>% select(CAPOP_2020_100m_HISP,
                                                      CAPOP_2020_100m_NHBLACK,
                                                      CAPOP_2020_100m_AMIND_COMBO,
                                                      CAPOP_2020_100m_NHASIAN,
                                                      CAPOP_2020_100m_NHHIPI,
                                                      CAPOP_2020_100m_NHWHITE,
                                                      CAPOP_2020_100m_OTHERMULTI)

```


```{r}
hisp_mean_reporting <- mean(reporting_df_filter$CAPOP_2020_100m_HISP)
hisp_mean_no_reporting <- mean(no_reporting_df_filter$CAPOP_2020_100m_HISP)
```

```{r}
t_test_hisp <- t.test(reporting_df_filter$CAPOP_2020_100m_HISP,no_reporting_df_filter$CAPOP_2020_100m_HISP)
#pval = 0.04961954

t_test_nhblack <- t.test(reporting_df_filter$CAPOP_2020_100m_NHBLACK,no_reporting_df_filter$CAPOP_2020_100m_NHBLACK)
#pval = 0.0006943373

t_test_amind_combo <- t.test(reporting_df_filter$CAPOP_2020_100m_AMIND_COMBO,no_reporting_df_filter$CAPOP_2020_100m_AMIND_COMBO)
#pval = 0.0002210685

t_test_nhasian <- t.test(reporting_df_filter$CAPOP_2020_100m_NHASIAN,no_reporting_df_filter$CAPOP_2020_100m_NHASIAN)
#pval = 0.07275465
#not significant

t_test_nhhipi <- t.test(reporting_df_filter$CAPOP_2020_100m_NHHIPI,no_reporting_df_filter$CAPOP_2020_100m_NHHIPI)
#pval = 0.7993745
#not significant

t_testnhwhite <- t.test(reporting_df_filter$CAPOP_2020_100m_NHWHITE,no_reporting_df_filter$CAPOP_2020_100m_NHWHITE)
#pval = 5.339204e-19
```

```{r}
hisp_reporting_mean <- mean(reporting_df_filter$CAPOP_2020_100m_HISP)
hisp_no_reporting_mean <- mean(no_reporting_df_filter$CAPOP_2020_100m_HISP)
#hisp_reporting_mean = 1.936476
#hisp_no_reporting_mean = 1.63793

```

```{r}
nhblack_reporting_mean <- mean(reporting_df_filter$CAPOP_2020_100m_NHBLACK)
nhblack_no_reporting_mean <- mean(no_reporting_df_filter$CAPOP_2020_100m_NHBLACK)

#nhblack_reporting_mean = 0.06228506
#nhblack_no_reporting_mean =  0.1616677
```

```{r}
amind_combo_reporting_mean <- mean(reporting_df_filter$CAPOP_2020_100m_AMIND_COMBO)
amind_combo_no_reporting_mean <- mean(no_reporting_df_filter$CAPOP_2020_100m_AMIND_COMBO)

#amind_combo_reporting_mean = 0.3629799
#amind_combo_no_reporting_mean = 0.260228
```

```{r}
nhasian_reporting_mean <- mean(reporting_df_filter$CAPOP_2020_100m_NHASIAN)
nhasian_no_reporting_mean <- mean(no_reporting_df_filter$CAPOP_2020_100m_NHASIAN)

#nhasian_reporting_mean = 0.187453
#nhasian_no_reporting_mean = 0.2852263
```

```{r}
nhhipi_reporting_mean <- mean(reporting_df_filter$CAPOP_2020_100m_NHHIPI)
nhhipi_no_reporting_mean <- mean(no_reporting_df_filter$CAPOP_2020_100m_NHHIPI)

#nhhipi_reporting_mean  = 0.01253714
#nhhipi_no_reporting_mean = 0.01362763

```

```{r}
nhwhite_reporting_mean <- mean(reporting_df_filter$CAPOP_2020_100m_NHWHITE)
nhwhite_no_reporting_mean <- mean(no_reporting_df_filter$CAPOP_2020_100m_NHWHITE)

#nhwhite_reporting_mean = 5.844748 
#nhwhite_no_reporting_mean = 4.106572

```
# Poverty Reporting 
```{r}
poverty <- rast((here(gdrive_data, "/IntermediateData/IntermediateData_Files/zonal_statistics/svi_2018_tract_overall_nad83.tif")))

plot(poverty)

```


```{r}
proj_poverty <- project(poverty, ca_raster)
proj_mask <- mask(proj_poverty, ca_raster)
plot(proj_mask)
```

```{r}
high_conflict_reporting_pov <- mask(high_conflict, buffer_conflict)

high_conflict_reporting_proj_pov <- project(high_conflict_reporting_pov,ca_raster)
```

```{r}
high_conflict_no_reporting_pov <- mask(high_conflict, buffer_conflict, inverse=TRUE)

high_conflict_no_reporting_proj_pov <- project(high_conflict_no_reporting_pov, ca_raster)

```

```{r}
poverty_mask_reporting <- mask(proj_mask, high_conflict_reporting_proj_pov)

poverty_mask_no_reporting <- mask(proj_mask, high_conflict_no_reporting_proj_pov)
```

```{r}
poverty_reporting_data_frame <- as.data.frame(poverty_mask_reporting)

poverty_no_reporting_data_frame <- as.data.frame(poverty_mask_no_reporting)

```

```{r}
mean_poverty_reporting <- mean(poverty_reporting_data_frame$svi_2018_tract_overall_nad83)
mean_poverty_no_reporting <- mean(poverty_no_reporting_data_frame$svi_2018_tract_overall_nad83)
#mean_poverty_reporting = 0.531186
#mean_poverty_no_reporting = 0.6585639
```
```{r}
t_test_poverty <- t.test(poverty_reporting_data_frame$svi_2018_tract_overall_nad83,poverty_no_reporting_data_frame$svi_2018_tract_overall_nad83)
```

# Linguistic Data Analysis

Create reporting and non-reporting rasters
```{r}
#high conflict binary raster (modeled conflict probability ≥ 0.7)
high_conflict <- rast(here(gdrive_data, "/AnalysisData/model_outputs/mod3sq_current_highconf.tif"))

#conflict reports buffered by 5km
buffered_confpoints <- vect(here(gdrive_data, "/IntermediateData/IntermediateData_Files/model_ready/conflict_buffered_refined/conflict_buffered_refined.shp"))

#reporting areas - within buffers
high_conflict_reporting <- mask(high_conflict, buffered_confpoints)

#non-reporting areas - outside of buffers
high_conflict_no_reporting <- mask(high_conflict, buffered_confpoints, inverse=TRUE)

```

Read in linguistic isolation data
```{r}
#load in the data and select for linguistic isolation (Percent limited English speaking households)
ces <- read_sf(here(gdrive_data, "/InputData/InputData_Files/calenviroscreen40shpf2021shp/CES4 Final Shapefile.shp")) %>% 
  select(Ling_Isol) %>% 
  naniar::replace_with_na(replace = list(Ling_Isol = -999))

#turn into a spatvector
lingisol <- vect(ces)

#rasterize using the conflict raster
lingisol_proj <- rasterize(lingisol, high_conflict, field = "Ling_Isol")
```

Mask the linguistic isolation data by these rasters and create data frames
```{r}
#isolation in reporting zones
isol_reporting <- mask(lingisol_proj, high_conflict_reporting)

isol_reporting_df <- as.data.frame(isol_reporting)

#isolation in non-reporting zones
isol_nonreporting <- mask(lingisol_proj, high_conflict_no_reporting)

isol_nonreporting_df <- as.data.frame(isol_nonreporting)
```

Take the mean of each category and run a t-test on it
```{r}
#mean isolation in reporting zones
mean_isol_reporting <- mean(isol_reporting_df$Ling_Isol) #2.007024

mean_isol_nonreporting <- mean(isol_nonreporting_df$Ling_Isol) #2.104348

#t test of these means
t_test_lingisol <- t.test(isol_reporting_df$Ling_Isol, isol_nonreporting_df$Ling_Isol) #not statistically significant; p val 0.511796
```

# Poverty analysis (more recent)

```{r}
#load in the data and select for linguistic isolation (Percent limited English speaking households)
ces <- read_sf(here(gdrive_data, "/InputData/InputData_Files/calenviroscreen40shpf2021shp/CES4 Final Shapefile.shp")) %>% 
  select(Ling_Isol) %>% 
  naniar::replace_with_na(replace = list(Ling_Isol = -999))
```