---
title: "Data Visualization"
author: "Mia Guarnieri"
date: '2022-11-02'
output: html_document
---

# Summary

This Rmd contains the code used to make  data visualizations for this project, including hotspot maps, a table of model coefficients, figures displaying transformed coefficients and their confidence intervals, and a marginal effects plot for one specified coefficient.

# Load in necessary packages (and file path, if needed)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#required packages

library(tidyverse)
library(terra)
library(tmap)
library(here)
library(sf)
library(ResourceSelection)
library(ggplot2)
library(ggtext)
library(kableExtra)
library(magick)


#setting up a file path (or file paths for multiple collaborators) - useful if your data is being read in from a Google Drive, your desktop, or somewhere other than your designated working directory

gdrive_data <- "/Users/mia/Library/CloudStorage/GoogleDrive-mguarnieri@ucsb.edu/Shared drives/Black_Bear_Aware/gdrive_data"

#gdrive_data <- "/Volumes/GoogleDrive/Shared drives/Black_Bear_Aware/gdrive_data"

#sam's file path
#gdrive_data <- "G:/Shared drives/Black_Bear_Aware/gdrive_data"

```

#Conflict data vis - THIS SECTION COMING OUT FOR PUBLIC-FACING FILES BUT KEPT IN ORIGINAL REPO

```{r}
conflict <- read_sf(here(gdrive_data, "/IntermediateData/IntermediateData_Files/conflict_formatted/conflict_points/conflict_points.shp")) %>% 
  separate(incdnt_d, into = c("inc_date", "inc_time"), sep = " ") %>% 
  separate(inc_date, into = c("month", "day", "year"), sep = "/") %>% 
  mutate(year = as.character.Date(year),
         "type" = cnfrmd_,
         "county" = cnty_nm,
         "conflict" = 1) %>% 
  filter(year != 2022)


ggplot(conflict) +
  geom_histogram(aes(x = year), stat = "count", fill = "#89a676") +
  theme_minimal() +
  labs(x = "Year",
       y = "Number of Reported Conflicts")

```


# Creating model hotspot maps

## Present day

Read in the model and all necessary data files

```{r}
#read in the model
model<- load(here("models", "sq_mod3climate.rds")) #you have to name the object "model or it will not read in right

#read in all necessary data layers

#elevation
dem <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/model_ready/elevation/dem_formatted.tif"))

#aspect
aspect <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/model_ready/aspect/aspect.tif"))

#terrain ruggedness
TRI <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/model_ready/terrain_ruggedness/terrain_ruggedness.tif"))

#human population density
popdens <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/model_ready/population_density/pop_den_2021_mask.tif"))

#land cover
lc <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/model_ready/land_cover/lc19_reclassed.tif"))

#road density
road_dens <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/model_ready/road_density/road_density_ca.tif"))

#distance to roads
road_dist <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/dist_to_road/model_ready/roads15_ca_distance.tif"))

#distance to streams
streams_dist <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/dist_to_stream/model_ready/stream17_ca_distance.tif"))

#distance to urban areas
urban_dist <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/dist_to_urban/model_ready/ua19_ca_distance.tif"))

#distance to recreational areas
rec_dist <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/dist_to_rec/model_ready/rec_ca_distance.tif"))

#distance to forest cover
forest_dist <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/dist_to_forest/model_ready/fc19_ca_distance.tif"))

#forest density
forest_dens <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/model_ready/forest_density/forest_dens_2016_mask.tif"))

#fire layers (present-day/historical values; 3, one for each time step)

#1 year ago
dist_fire_1yr <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/model_ready/dist_fire/fire20_ca_distance.tif"))

#2-3 years ago
dist_fire_23yrs <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/model_ready/dist_fire/fire18_19_ca_distance.tif"))

#4-5 years ago
dist_fire_45yrs <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/model_ready/dist_fire/fire16_17_ca_distance.tif"))

#drought (present-day/historical values)
drought <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/drought/drought_rasters/UDM_2021/mean_drought_21.tif"))


#create a "stack" of these rasters so they can all be used in a function as one object
stackca_clim_rast <- c(dem, aspect, TRI, popdens, lc, road_dens, road_dist, streams_dist, urban_dist, rec_dist, forest_dist, forest_dens, dist_fire_1yr, dist_fire_23yrs, dist_fire_45yrs, drought)

#read in the data used to create the model
climate_model_data <- read_csv(here(gdrive_data, "/AnalysisData/model_df_climate/1_31_2023/climate_model_data_allyears.csv"))
```

Wrangle the data files to be in the correct format for hotspot generation and create the hotspot map

```{r}
#change the names of the raster stack to match the names of our model variables EXACTLY (must be in the correct order as well, so that names correspond with correct data raster)
names(stackca_clim_rast) <- c("dem", "aspect", "TRI", "popdens", "lc", "road_dens", "road_dist", "streams_dist", "urban_dist", "rec_dist", "forest_dist", "forest_dens", "dist_fire_1yr", "dist_fire_23yrs", "dist_fire_45yrs", "drought")

#any categorical variables used (land cover) need to be identified before generating the hotspot map

#turn land cover categories into factors
lc_factors <- as.factor(climate_model_data$lc)

#turn the levels of that factor into into an object
f_lc <- levels(lc)

#use the predict function in the terra package to generate a map of predicted conflict risk within the area of your stacked spatraster object (that contains your environmental and climate variables); make sure to specify the levels of categorical variables (land cover) using the `factors` argument and remove NA values using `na.rm = TRUE`
map <- terra::predict(object = stackca_clim_rast, model = sq_mod3climate, fun = predict, factors = f_lc, na.rm = TRUE, type = "response")

#save the map for future analysis
writeRaster(map, filename = here(gdrive_data, "/AnalysisData/model_outputs/mod3clim_map_squared.tif"), filetype = "GTiff", overwrite= TRUE)

#plot the map to test it visually
plot(map, main="Model Predictions")
```

## Future

Read in the model and all necessary data files - we only used projected (2030) data for fire and drought and used most recently available data for other variables, but any of these rasters could be replaced with projected data if desired.

```{r}
#read in the model
model<- load(here("models", "sq_mod3climate.rds")) #you have to name the object "model or it will not read in right

#read in all necessary data layers

#elevation
dem <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/model_ready/elevation/dem_formatted.tif"))

#aspect
aspect <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/model_ready/aspect/aspect.tif"))

#terrain ruggedness
TRI <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/model_ready/terrain_ruggedness/terrain_ruggedness.tif"))

#human pop density
popdens <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/model_ready/population_density/pop_den_2021_mask.tif"))

#land cover
lc <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/model_ready/land_cover/lc19_reclassed.tif"))

#road density
road_dens <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/model_ready/road_density/road_density_ca.tif"))

#distance to roads
road_dist <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/dist_to_road/model_ready/roads15_ca_distance.tif"))

#distance to streams
streams_dist <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/dist_to_stream/model_ready/stream17_ca_distance.tif"))

#distance to urban areas
urban_dist <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/dist_to_urban/model_ready/ua19_ca_distance.tif"))

#distance to recreational areas
rec_dist <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/dist_to_rec/model_ready/rec_ca_distance.tif"))

#distance to forest cover
forest_dist <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/distance_rasters/dist_to_forest/model_ready/fc19_ca_distance.tif"))

#forest density
forest_dens <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/model_ready/forest_density/forest_dens_2016_mask.tif"))


#fire layers (projected values; 3, one for each time step)

#1 year ago
dist_fire_1yr <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/model_ready/dist_proj_fire/dist_proj_fire_2029_ca.tif"))

#2-3 years ago
dist_fire_23yrs <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/model_ready/dist_proj_fire/dist_proj_fire_2027_2028_ca.tif"))

#4-5 years ago
dist_fire_45yrs <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/model_ready/dist_proj_fire/dist_proj_fire_2025_2026_ca.tif"))

#drought (projected)
drought <- rast(here(gdrive_data, "/IntermediateData/IntermediateData_Files/model_ready/drought/pdsi_masked.tif"))

#create a "stack" of these rasters so they can all be used in a function as one object
stackca_clim_rast <- c(dem, aspect, TRI, popdens, lc, road_dens, road_dist, streams_dist, urban_dist, rec_dist, forest_dist, forest_dens, dist_fire_1yr, dist_fire_23yrs, dist_fire_45yrs, drought)

#read in the data used to create the model
climate_model_data <- read_csv(here(gdrive_data, "/AnalysisData/model_df_climate/1_31_2023/climate_model_data_allyears.csv"))
```

Wrangle the data files to be in the correct format for hotspot generation and create the hotspot map

```{r}
#change the names of the raster stack to match the names of our model variables EXACTLY (must be in the correct order as well, so that names correspond with correct data raster)
names(stackca_clim_rast) <- c("dem", "aspect", "TRI", "popdens", "lc", "road_dens", "road_dist", "streams_dist", "urban_dist", "rec_dist", "forest_dist", "forest_dens", "dist_fire_1yr", "dist_fire_23yrs", "dist_fire_45yrs", "drought")

#any categorical variables used (land cover) need to be identified before generating the hotspot map

#turn land cover categories into factors
lc_factors <- as.factor(climate_model_data$lc)

#turn the levels of that factor into into an object
f_lc <- levels(lc)

#use the predict function in the terra package to generate a map of predicted conflict risk within the area of your stacked spatraster object (that contains your environmental and climate variables); make sure to specify the levels of categorical variables (land cover) using the `factors` argument and remove NA values using `na.rm = TRUE`
map <- terra::predict(object = stackca_clim_rast, model = sq_mod3climate, fun = predict, factors = f_lc, na.rm = TRUE, type = "response")

#save the map for future analysis
writeRaster(map, filename = here(gdrive_data, "/AnalysisData/model_outputs/projected_mod3_clim_map_squared.tif"), filetype = "GTiff", overwrite= TRUE)

#plot the map to test it visually
plot(map, main="Model Predictions")
```

## Differenced map (Future - Present)

```{r}
#read in the saved maps for current and future predicted conflict (the outputs from the terra `predict` function)
current <- rast(here(gdrive_data, "/AnalysisData/model_outputs/mod3clim_map_squared.tif"))

future <- rast(here(gdrive_data, "/AnalysisData/model_outputs/projected_mod3_clim_map_squared.tif"))

#subtract current conflict risk from future conflict risk
diff <- future - current

#plot all three rasters to compare visually
plot(diff)

plot(current)

plot(future)

#save the map for future analysis
writeRaster(diff, filename = here(gdrive_data, "/AnalysisData/model_outputs/differenced_conflict_risk.tif"), filetype = "GTiff", overwrite= TRUE)
```


# Model coefficient tables and plots

## Overall model coefficients table

```{r}
#load in the model
model<- load(here("models", "sq_mod3climate.rds")) #you have to name the object "model or it will not read in right

#create a model summary object
mod_sum <- summary(sq_mod3climate)

#extract the coefficients from the model summary object
coeffs <- mod_sum$coefficients

#using kable, create a formatted table of model coefficients
nice_table <- coeffs %>% 
  kable() %>% 
  kable_classic(full_width = TRUE, html_font = "Times New Roman")

#check the table
nice_table

#save the table as a photo object (.jpeg)
save_kable(nice_table, file = here("/Users/mia/Desktop/Black Bear Aware/coeffs_nicetable.jpeg"), zoom = 2.5)
```

## Individual plots for model coefficient groups

To create these plots, use the coefficients table from the code above (coeffs)

```{r}
#convert the row names (variables) to a column for plotting
coeffs_tidy <- cbind(rownames(coeffs), data.frame(coeffs, row.names=NULL)) %>% 
  janitor::clean_names() %>% 
  mutate(coeff_trans = exp(estimate),
         ster_trans = exp(std_error),
         prob_change = ifelse(coeff_trans < 1,
                              (1 - coeff_trans)*100,
                              (coeff_trans - 1)*100))

#remove perennial ice and snow to make plot look more reasonable and convert log odds to probability
coeffs_tidy_nosnow <- coeffs_tidy %>% 
  filter(rownames_coeffs != "lcperrenial_ice_snow") %>% 
  mutate(coeff_trans = exp(estimate)) 

#test plot with ice and snow removed for visual check
ggplot(coeffs_tidy_nosnow, aes(x = coeff_trans, y = rownames_coeffs)) + 
  geom_point() +
  geom_pointrange(aes(xmin = estimate-std_error, xmax = estimate+std_error)) +
  scale_x_continuous(breaks = seq(-3, 4, 0.5))

#make plots for each variable group

#first, separate into independent dataframes by variable type

#perennial ice and snow land cover class; not plotted but standard error and confidence interval calculated (values included in figure caption of other land cover classes)
ice_snow <- coeffs_tidy[10,] %>% #select ice and snow row from coeffs dataframe
  mutate(coeff_trans = exp(estimate), #convert from log odds to odds
         prob_change = (coeff_trans - 1) * 100, #find the change in likelihood as a percentage
         ster_min = exp(estimate - (2 * std_error)), #create std. error min and max values
         ster_max = exp(estimate + (2 * std_error)),
         cimax = (1 - ster_min)*100, #create confidence interval min and max values
         cimin = (1 - ster_max)*100)

#elevation; not plotted so no confidence interval and standard error calculated
elev <- coeffs_tidy_nosnow[c(2), ] %>% #select elevation from coeffs dataframe
  mutate(coeff_trans = exp(estimate*1000), #convert m to km (*1000) and convert from log odds to odds
         prob_change = (coeff_trans - 1)*100) %>% #find the change in likelihood as a percentage
  mutate(rownames_coeffs = c("Elevation")) #rename to variable

#terrain ruggedness; not plotted so no confidence interval and standard error calculated
tri <- coeffs_tidy_nosnow[c(18), ] %>% #select from dataframe
  mutate(coeff_trans = exp(estimate), #convert to odds
         prob_change = (1 - coeff_trans)*100) %>% #find change in likelihood
  mutate(rownames_coeffs = c("Terrain Ruggedness")) #rename

#land cover variables
bluevar_lconly <- coeffs_tidy_nosnow[c(3:12), ] %>% #select land cover classes
  mutate(coeff_trans = exp(estimate), #convert to odds
         prob_change = (coeff_trans - 1) * 100, #find change in likelihood
         ster_min = exp(estimate - (2 * std_error)), #find standard error
         ster_max = exp(estimate + (2 * std_error)),
         cimax = (ster_max - 1) * 100, #find conficence interval
         cimin = ifelse(ster_min > 1,
                        (ster_min - 1) * 100,
                        (1 - ster_min) * 100)) %>% 
  mutate(rownames_coeffs = c("High-Intensity Development", "Low-Intensity Development", "Medium-Intensity Development", "Developed Open Space", "Forest", "Grassland/Herbaceous", "Open Water", "Planted/Cultivated", "Shrub/Scrub", "Wetlands"))

#variables biologically relevant to black bears; not plotted to standard error and confidence interval not calculated here
greenvar <- coeffs_tidy_nosnow[c(13, 17), ] %>% #select variables
  mutate(coeff_trans = exp(estimate*1000), #convert m to km (*1000) and convert from log odds to odds
         prob_change = (1 - coeff_trans)*100) %>% #find change in likelihood
  mutate(rownames_coeffs = c("Distance to Forest", "Distance to Streams")) #rename

#variables that show human influence

#separate the distance variables (these need to be converted from m to km)
gray_dist <- coeffs_tidy_nosnow[c(16, 19), ] %>% #select variables
  mutate(coeff_trans = exp(estimate*1000), #convert m to km (*1000) and convert from log odds to odds
         ster_trans = exp(std_error * 1000), #standard error (converted)
         prob_change = (1 - coeff_trans)*100, #find change in likelihood
         ster_min = exp((estimate - (2 * std_error))*1000), #standard error min and max values
         ster_max = exp((estimate + (2 * std_error))*1000),
         cimax = (1 - ster_min)*100, #confidence intervals min and max values
         cimin = (1 - ster_max)*100) %>% 
  mutate(rownames_coeffs = c( "Distance to Recreational Areas", "Distance to Urban Areas")) #rename

#population density
gray_pop <- coeffs_tidy_nosnow[c(14), ] %>% #select variable
  mutate(coeff_trans = exp(estimate), #convert from log odds to odds
         ster_trans = exp(std_error), #transform standard error
         prob_change = ifelse(coeff_trans < 1, #find change in likelihood; 1-x if x < 1 and x-1 if x > 1
                              (1 - coeff_trans)*100,
                              (coeff_trans - 1)*100),
         ster_min = exp(estimate - (2 * std_error)), #standard error min and max values
         ster_max = exp(estimate + (2 * std_error)),
         cimax = (ster_min - 1)*100, #confidence intervals min and max values
         cimin = (ster_max - 1)*100) %>% 
  mutate(rownames_coeffs = c("Population Density")) #rename

#bind distance and non-distance human influence variables together for plotting
grayvar<- rbind(gray_dist, gray_pop)

#fire variables
orangevar <- coeffs_tidy_nosnow[c(20:22), ] %>% #select variables
  mutate(coeff_mult = estimate * 1000, #multiply by 1000 to convert m to km
         ster_mult = std_error * 1000,
         coeff_trans = exp(coeff_mult), #transform from log odds to odds
         prob_change = (1 - coeff_trans) * 100, #find change in likelihood
         ster_min = exp((estimate - (2 * std_error))*1000), #standard error min and max
         ster_max = exp((estimate + (2 * std_error))*1000),
         cimax = (1 - ster_min)*100, #confidence intervals min and max
         cimin = (1 - ster_max)*100) %>% 
  mutate(rownames_coeffs = c("Distance to Fire 1 Year Ago", "Distance to Fire 2-3 Years Ago", "Distance to Fire 4-5 Years Ago")) #rename

#drought; not plotted to standard error min and max, CI min and max not calculated here
brownvar <- coeffs_tidy_nosnow[c(23), ] %>% #select variable
  mutate (coeff_trans = exp(estimate), #transform from log odds to odds
         ster_trans = exp(std_error), #transform standard error
         prob_change = (1 - coeff_trans)*100)


#now that the data is wrangled, plot desired variable groups

#plot for fire variables
ggplot(orangevar, mapping = aes(x = prob_change, y = reorder(rownames_coeffs, prob_change))) + 
  geom_point(color = "#cc5500", size = 3) +
  geom_linerange(aes(xmin = cimin, xmax = cimax), color = "#cc5500") +
  labs(y = "Coefficient", x = "Estimate") +
  theme_minimal() +
  theme(plot.caption = ggtext::element_markdown(hjust = 0),
        text = element_text(family = "Times New Roman")) 

#plot for land cover variables
ggplot(bluevar_lconly, mapping = aes(x = prob_change, y = reorder(rownames_coeffs, prob_change))) + 
  geom_point(color = "#528faa", size = 3) +
  geom_linerange(aes(xmin = cimin, xmax = cimax), color = "#528faa") +
  labs(y = "Coefficient", x = "Estimate") +
  theme_minimal() +
  theme(plot.caption = ggtext::element_markdown(hjust = 0),
        text = element_text(family = "Times New Roman"))

#plot for human influence variables
ggplot(grayvar, mapping = aes(x = prob_change, y = reorder(rownames_coeffs, prob_change))) + 
  geom_point(color = "#777777", size = 3) +
  geom_linerange(aes(xmin = cimin, xmax = cimax), color = "#777777") +
  labs(y = "Coefficient", x = "Estimate") +
  theme_minimal() +
  theme(plot.caption = ggtext::element_markdown(hjust = 0),
        text = element_text(family = "Times New Roman"))

```

## Marginal effects plot

```{r}
#load the model
model<- load(here("models", "sq_mod3climate.rds")) #you have to name the object "model or it will not read in right

#create a marginal effects plot using the `mep` function within the ResourceSelection package
mep(sq_mod3climate, which = "popdens", col.points = "#c3c2b1", pch = 1, n = 100, family = "Times New Roman")

```
